### 2015-11-08 paper/paper.tex
Start writing

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :

\documentclass{llncs}

% \usepackage{showframe}

\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{cleveref}                  % no need to type Figure etc

% lncs style
\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Fig.}{Figs.}
\crefname{table}{Table}{Tables}
\Crefname{table}{Table}{Tables}
\crefname{proposition}{Proposition}{Propositions}
\Crefname{proposition}{Proposition}{Propositions}

\title{Portfolios of Subgraph Isomorphism Algorithms}

\author{
    Lars Kotthoff\inst{1}
    \and Ciaran McCreesh\thanks{This work was supported by the Engineering
        and Physical Sciences Research Council [grant number EP/K503058/1]}\inst{2}
    \and Patrick Prosser\inst{2}
    \and Christine Solnon\inst{3}}

\institute{
    University of British Columbia, Vancouver, Canada
    \and University of Glasgow, Glasgow, Scotland
    \and INSA-Lyon, LIRIS, UMR5205, F-69621, France}

\begin{document}

\maketitle

\begin{abstract}
There are several competitive algorithms for the subgraph isomorphism problem, and one of these
algorithms additionally has a choice of parameter. As with many other hard problems, the single best
choice of algorithm overall is rarely the best algorithm on a per-instance basis. We show that using
witchcraft to pick from a portfolio of algorithms on an instance by instance basis provides super
dooper amazingly quite OKish results, most of the time, and opens up some flexibility in algorithm
design.
\end{abstract}

\section{Introduction}

The \emph{(non-induced) subgraph isomorphism problem} is to find an injective mapping from a small
\emph{pattern} graph to a large \emph{target} graph which preserves adjacency. This problem is
\NP-complete, but applications in chemistry and computer vision and model checking and some other
things we should cite mean various exact algorithms exist:

\begin{itemize}
    \item VF2 \cite{Cordella:2004} is a backtracking search algorithm which is especially fast on
        trivially satisfiable instances (which are common in certain application areas).
    \item LAD \cite{Solnon:2010} and SND \cite{Audemard:2014} do very strong reasoning, at the
        expense of making few guesses per second when working with larger graphs.
    \item Most recently, the Glasgow algorithm \cite{McCreesh:2015} does an expensive preprocessing
        pass once, followed by weaker reasoning during search.
\end{itemize}

\noindent The experiments by the Glasgow authors indicated that overall, the Glasgow algorithm is
the single best algorithm when evaluated across a wide range of problem instances. However, they
also observed that on an instance by instance basis, it is often not the single best.  Additionally,
the Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about
non-adjacent vertices.  The choice of paths of length 3 was used as a reasonable compromise---longer
paths lead to prohibitively expensive preprocessing on larger, denser instances. Again, this is
often not the best choice on an instance by instance basis: sometimes path-based reasoning gives no
benefit at all, sometimes considering only paths of length 2 suffices, occasionally paths of
length 4 are helpful, and even looking at paths of length 3 is relatively expensive on some graphs.

?? Also, we can use supplemental graphs in LAD. Christine: can you explain what you implemented?

We could try to come up with rules to pick an algorithm on an instance by instance basis using our
knowledge of how the algorithms behave: for example, maybe we should use the Glasgow algorithm with
paths of length 2 with large target graphs (but how large?), and only use length 3 on smaller
graphs. Also, we could guess that LAD might do better than Glasgow on smaller, heavily-structured
graphs. However, this is not very systematic.

?? Lars: a paragraph of fluff about instance selection please.

We show that picking algorithms and parameter choices on an instance by instance basis, by
considering features of the pattern and target graphs, gives better overall performance than any
single algorithm. This is the first time this has been done for subgraph isomorphism, and we must
consider an unusual factor: our features must look at pairs of inputs, and the relationships between
them.

?? Lars: do you know of any other problems where you have two inputs rather than one? Or is this
novel?

?? Can we get better results if we explicitly consider, say, how close the target density and
pattern density are?

\section{Experiments}

Our runtimes were measured on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds (a little over a day).

?? Lars: your stuff here.

\subsection{Problem instances}

All instances are available in a simple text format
\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}.

Real-world and application-derived graphs:

\begin{itemize}
    \item The LV graphs (family 2). Also, the same targets, but with larger pattern graphs (family
        11, Multiple edges have been combined into a single edge.)
    \item CVIU11 images (family 8)
    \item CVIU11 meshes (family 10)
    \item PR15 (family 9)
\end{itemize}

\noindent Randomly generated instances with particular properties:

\begin{itemize}
    \item Bounded-degree random graphs (family 3, 4) \cite{GraphDatabase}
    \item Random 4d mesh graphs (family 5, 6) \cite{GraphDatabase}
    \item Random graphs (family 7) \cite{GraphDatabase}
    \item Randomly generated scale-free graphs. (family 1) \cite{Solnon:2010}
    \item Phase transition (family 12)
\end{itemize}

\noindent Note that this is not like SAT, where we can reasonably split things up based upon which
family they're from.

\subsection{Machine learning voodoo stuff}

\subsection{Results}

?? Can we do better than picking the single best on a family by family basis? Bear in mind that we
don't have the family as a usable feature.

\section{Conclusion and Future Work}

This stuff works respectably well, considering we're dealing with two inputs rather than one.

We have only looked at sequential algorithms: the Glasgow algorithm has a parallel implementation,
and a similar technique could easily be used to parallelise LAD.  We have also not considered the
induced variant of the problem, nor instances involving labels or directed edges.  We could also
look at other encodings. For example, a direct SAT requires $O(v^4)$ clauses, but may be good in
certain situations.

\bibliographystyle{splncs}
\bibliography{paper}

\end{document}

### 2015-11-08 portfolio/as-graphs.Rnw
add portfolio stuff

\documentclass{article}

\usepackage{booktabs}
\usepackage{hyperref}

\title{Algorithm Selection for Graph Algorithms}

\author{}

\begin{document}

\maketitle

<<echo=FALSE, message=FALSE, warning=FALSE>>=
opts_chunk$set(comment=NA, fig.width='\\textwidth', fig.height=5, fig.pos='!ht')
library(llama)
library(BBmisc)
library(xtable)
library(plyr)
library(aslib)

library(parallelMap)
parallelStartSocket(4)
parallelLibrary("llama", "mlr")
@

\section*{Data Analysis}

<<echo=FALSE, results='asis'>>=
source("convertToAslib.R")

asscenario = parseASScenario("graphs-2015")
ldf = convertToLlamaCVFolds(asscenario)
# add family information
times = read.table("gpgnode-sip-runtimes.data", header = TRUE, sep = " ", stringsAsFactors = FALSE)
timess = times
times$family = factor(times$family)
ldf$data$family = times[order(times$instance, as.character(ldf$data$instance_id)), 'family']
ldf$extra = "family"

tab = t(sapply(ldf$data[,ldf$performance], mean))
print(xtable(tab, caption = "Mean performance of each algorithm."), booktabs = TRUE, include.rownames = FALSE)

tabs = by(ldf$data, ldf$data$family, function(x) as.data.frame.matrix(t(sapply(x[,ldf$performance], mean))))
tab = rbind.fill(tabs)
tab$family = rownames(tab)
tab = tab[,order(names(tab))]
print(xtable(tab, digits = 3, caption = "Mean performance by family."), booktabs = TRUE, include.rownames = FALSE)

tab = t(sapply(ldf$data[,ldf$performance], median))
print(xtable(tab, caption = "Median performance of each algorithm."), booktabs = TRUE, include.rownames = FALSE)

tabs = by(ldf$data, ldf$data$family, function(x) as.data.frame.matrix(t(sapply(x[,ldf$performance], median))))
tab = rbind.fill(tabs)
tab$family = rownames(tab)
tab = tab[,order(names(tab))]
print(xtable(tab, digits = 3, caption = "Median performance by family."), booktabs = TRUE, include.rownames = FALSE)

tab = t(table(unlist(ldf$best)))
print(xtable(tab, caption = "Number of times each algorithm is best. More than one algorithm may be best for one instance if there are ties."), booktabs = TRUE, include.rownames = FALSE)

tabs = tapply(ldf$best, ldf$data$family, function(x) as.data.frame.matrix(t(table(unlist(x)))))
tab = rbind.fill(tabs)
tab[is.na(tab)] = 0
tab$family = rownames(tab)
tab = tab[,order(names(tab))]
print(xtable(tab, digits = 0, caption = "Number of times each algorithm is best by family."), booktabs = TRUE, include.rownames = FALSE)
@

\clearpage

\section*{Exploratory Data Analysis}

\subsection*{Algorithms}

\subsubsection*{Overview of performance values}

The following statistics were calculated from the performance values of each alg
orithm:

\begin{itemize}
  \item obs (number of observations within performance values)
  \item nas (number of NAs, i.e.\ missing values, within performance values)
  \item min (minimum), mean (arithmetic mean), max (maximum), sd (standard deviation)
  \item qu\_1st (1st quartile = lower quartile = 25\%-quantile)
  \item med (median = 50\%-quantile)
  \item qu\_3rd (3rd quartile = upper quartile = 75\%-quantile)
  \item coeff\_var (coefficient of variation = standard deviation / arithmetic mean)
\end{itemize}

<<echo=FALSE, results='asis'>>=
source("eda_config.R")
config = readEDAConfig(asscenario, ".")

res = summarizeAlgoPerf(asscenario)
print(xtable(res, digits = 6, display = c("s", "d", "d", rep("g", 8))), booktabs = TRUE)
@

\subsubsection*{Summary of the runstatus per algorithm}

The following table summarizes the runstatus of each algorithm over all instances (in \%).

<<echo=FALSE, results='asis'>>=
res = summarizeAlgoRunstatus(asscenario)
print(xtable(res, digits = 3), booktabs = TRUE)
@

\subsubsection*{Dominated Algorithms}

<<echo=FALSE, results='asis'>>=
res = findDominatedAlgos(asscenario, reduce = TRUE, type = "character")
if (nrow(res) != 0) {
  print(xtable(res), include.rownames = TRUE, "html")
} else {
  cat("None of the algorithms was superior to any of the other.")
}
@

An algorithm (A) is considered to be superior to an other algorithm (B), if it
has at least an equal performance on all instances (compared to B) and if
it is better on at least one of them. A missing value is automatically a worse
performance. However, instances which could not be solved by either one of the
algorithms, were not considered for the dominance relation.

\subsubsection*{Boxplot of performance values}

Important note w.r.t.\ some of the following plots:
If appropriate, we imputed performance values for failed runs.
We used $max + 0.3 * (max - min)$, in case of minimization problems,
or $min - 0.3 * (max - min)$, in case of maximization problems.

In addition, a small noise is added to the imputed values (except for the
cluster matrix, based on correlations, which is shown at the end of this page).

<<echo=FALSE, fig.cap="Performance values with imputation.", message=FALSE>>=
plotAlgoPerfBoxplots(asscenario, impute.zero.vals = config$algo.perf.impute.zero.vals, log = config$algo.perf.log)
@

<<echo=FALSE, fig.cap="Cumulative density plot of performance values with imputation.", message=FALSE>>=
plotAlgoPerfCDFs(asscenario, impute.zero.vals = config$algo.perf.impute.zero.vals, log = config$algo.perf.log)
@

<<echo=FALSE, fig.cap="Performance values without imputation.", message=FALSE>>=
plotAlgoPerfScatterMatrix(asscenario, impute.zero.vals = config$algo.perf.impute.zero.vals, log = config$algo.perf.log)
@

The figure showing the correlations of the ranks of the performance values shows
the Spearman correlation coefficient. Missing values were imputed prior to
computing the correlation coefficients. The algorithms are ordered in a way that
similar (highly correlated) algorithms are close to each other. Per default the
clustering is based on hierarchical clustering, using Ward's method.

<<echo=FALSE, fig.cap="Algorithm correlation clusters.", message=FALSE>>=
plotAlgoCorMatrix(asscenario)
@

\clearpage

\subsection*{Features}

\subsubsection*{Overview of Feature Values}

The following measures were taken for the results of each feature:

\begin{itemize}
  \item obs (number of observations within the feature set)
  \item NAs (number of NAs, i.e. missing values, within the feature set)
  \item min (minimum), mean (arithmetic mean), max (maximum), std\_dev (standard deviation)
  \item 1st\_qu (1st quartile = lower quartile = 25\%-quantile)
  \item median (median = 50\%-quantile)
  \item 3rd\_qu (3rd quartile = upper quartile = 75\%-quantile)
  \item co\_var (coefficient of variation = standard deviation / arithmetic mean)
\end{itemize}

<<echo=FALSE, results='asis'>>=
res = summarizeFeatureValues(asscenario)
print(xtable(res, display = c("s", "d", "d", rep("g", 8))), booktabs = TRUE)
@

\subsubsection*{Summary of feature steps}

The following table summarizes the feature steps over all instances.

\begin{itemize}
  \item size (percentage of features affected by this step)
  \item ok, \ldots, other (percentage of run status)
  \item cost\_min, cost\_mean, cost\_max (range of costs for this step)
  \item cost\_na (percentage of missing values occurring in costs)
\end{itemize}

<<echo=FALSE, results='asis'>>=
res = summarizeFeatureSteps(asscenario)
print(xtable(res, display = c("s", "d", rep("g", ncol(res) - 1L))), booktabs = TRUE)
@

%\subsubsection*{Duplicated Features}
%
%<<echo=FALSE, results='asis'>>=
%res = checkDuplicatedInstances(asscenario)
%if (length(res) > 0L) {
%  numbers = sapply(res, length)
%  res = sapply(seq_along(res), function(i) {
%    if (numbers[i] > 5)
%      res[[i]] = c(res[[i]][1:5], "...")
%    collapse(res[[i]], sep = " +++++++ ")
%  })
%  res = data.frame(block = seq_along(numbers), numbers = numbers, instances = res)
%  print(xtable(res), include.rownames = FALSE, booktabs = TRUE)
%} else {
%  cat("There were no duplicated features.")
%}
%@

\clearpage

\section*{Algorithm Selection Results}

<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
resvbs = data.frame(model = "virtual best",
    mean.performance = mean(parscores(ldf, vbs, factor = 1)),
    median.performance = median(parscores(ldf, vbs, factor = 1)))
ressb = data.frame(model = "single best",
    mean.performance = mean(parscores(ldf, singleBest, factor = 1)),
    median.performance = median(parscores(ldf, singleBest, factor = 1)))

modelc = classify(makeLearner("classif.randomForest"), ldf)
resc = data.frame(model = "classification",
    mean.performance = mean(parscores(ldf, modelc, factor = 1)),
    median.performance = median(parscores(ldf, modelc, factor = 1)))
times = sapply(modelc$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classification."), booktabs = TRUE)

modelr = regression(makeLearner("regr.randomForest"), ldf)
resr = data.frame(model = "regression",
    mean.performance = mean(parscores(ldf, modelr, factor = 1)),
    median.performance = median(parscores(ldf, modelr, factor = 1)))
times = sapply(modelr$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regression."), booktabs = TRUE)

modelcp = classifyPairs(makeLearner("classif.randomForest"), ldf)
rescp = data.frame(model = "classifyPairs",
    mean.performance = mean(parscores(ldf, modelcp, factor = 1)),
    median.performance = median(parscores(ldf, modelcp, factor = 1)))
times = sapply(modelcp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classificationPairs."), booktabs = TRUE)

modelrp = regressionPairs(makeLearner("regr.randomForest"), ldf)
resrp = data.frame(model = "regressionPairs",
    mean.performance = mean(parscores(ldf, modelrp, factor = 1)),
    median.performance = median(parscores(ldf, modelrp, factor = 1)))
times = sapply(modelrp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regressionPairs."), booktabs = TRUE)

tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results."), booktabs = TRUE, include.rownames = FALSE)
@

<<echo=FALSE,fig.cap='Single best algorithm vs.\ portfolio performance (best model).'>>=
best = which.min(c(resc$mean.performance, resr$mean.performance, rescp$mean.performance, resrp$mean.performance))
mod = list(modelc, modelr, modelcp, modelrp)[[best]]
perfScatterPlot(parscores, mod, singleBest, ldf, addCostsx = TRUE, addCostsy = FALSE, factor = 1, pargs=aes(colour = family)) + scale_x_log10() + scale_y_log10() + xlab("portfolio") + ylab("single best")
@

\clearpage

\subsection*{Algorithm Selection Results with Presolver}

Run vf2 for 0.05 second as a presolver.

<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
# presolver
ldf2 = ldf
presolvetime = 0.05
ldf2$data = subset(ldf$data, ldf$data$vf2 > presolvetime)
ldf2$best = subset(ldf$best, ldf$data$vf2 > presolvetime)
ldf2$train = NULL
ldf2$test = NULL
ldf2 = cvFolds(ldf2)
timesss = subset(timess, timess$instance %in% ldf2$data$instance_id)
ldf2$data$family = factor(timesss[order(timesss$instance, as.character(ldf2$data$instance_id)), 'family'])
ldf2$extra = "family"

tab = t(sapply(ldf2$data[,ldf2$performance], mean))
print(xtable(tab, caption = "Mean performance of each algorithm."), booktabs = TRUE, include.rownames = FALSE)

tabs = by(ldf2$data, ldf2$data$family, function(x) as.data.frame.matrix(t(sapply(x[,ldf2$performance], mean))))
tab = rbind.fill(tabs)
tab$family = rownames(tab)
tab = tab[,order(names(tab))]
print(xtable(tab, digits = 3, caption = "Mean performance by family."), booktabs = TRUE, include.rownames = FALSE)

tab = t(sapply(ldf2$data[,ldf2$performance], median))
print(xtable(tab, caption = "Median performance of each algorithm."), booktabs = TRUE, include.rownames = FALSE)

tabs = by(ldf2$data, ldf2$data$family, function(x) as.data.frame.matrix(t(sapply(x[,ldf2$performance], median))))
tab = rbind.fill(tabs)
tab$family = rownames(tab)
tab = tab[,order(names(tab))]
print(xtable(tab, digits = 3, caption = "Median performance by family."), booktabs = TRUE, include.rownames = FALSE)

tab = t(table(unlist(ldf2$best)))
print(xtable(tab, caption = "Number of times each algorithm is best. More than one algorithm may be best for one instance if there are ties."), booktabs = TRUE, include.rownames = FALSE)

tabs = tapply(ldf2$best, ldf2$data$family, function(x) as.data.frame.matrix(t(table(unlist(x)))))
tab = rbind.fill(tabs)
tab[is.na(tab)] = 0
tab$family = rownames(tab)
tab = tab[,order(names(tab))]
print(xtable(tab, digits = 0, caption = "Number of times each algorithm is best by family."), booktabs = TRUE, include.rownames = FALSE)
@

\clearpage

<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
resvbs = data.frame(model = "virtual best",
    mean.performance = mean(parscores(ldf2, vbs, factor = 1)),
    median.performance = median(parscores(ldf2, vbs, factor = 1)))
ressb = data.frame(model = "single best",
    mean.performance = mean(parscores(ldf2, singleBest, factor = 1)),
    median.performance = median(parscores(ldf2, singleBest, factor = 1)))

modelc = classify(makeLearner("classif.randomForest"), ldf2)
resc = data.frame(model = "classification",
    mean.performance = mean(parscores(ldf2, modelc, factor = 1)),
    median.performance = median(parscores(ldf2, modelc, factor = 1)))
times = sapply(modelc$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classification."), booktabs = TRUE)

modelr = regression(makeLearner("regr.randomForest"), ldf2)
resr = data.frame(model = "regression",
    mean.performance = mean(parscores(ldf2, modelr, factor = 1)),
    median.performance = median(parscores(ldf2, modelr, factor = 1)))
times = sapply(modelr$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regression."), booktabs = TRUE)

modelcp = classifyPairs(makeLearner("classif.randomForest"), ldf2)
rescp = data.frame(model = "classifyPairs",
    mean.performance = mean(parscores(ldf2, modelcp, factor = 1)),
    median.performance = median(parscores(ldf2, modelcp, factor = 1)))
times = sapply(modelcp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classificationPairs."), booktabs = TRUE)

modelrp = regressionPairs(makeLearner("regr.randomForest"), ldf2)
resrp = data.frame(model = "regressionPairs",
    mean.performance = mean(parscores(ldf2, modelrp, factor = 1)),
    median.performance = median(parscores(ldf2, modelrp, factor = 1)))
times = sapply(modelrp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regressionPairs."), booktabs = TRUE)

tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolver."), booktabs = TRUE, include.rownames = FALSE)
@

<<echo=FALSE,fig.cap='Single best algorithm vs.\ portfolio performance with presolver (best model).'>>=
best = which.min(c(resc$mean.performance, resr$mean.performance, rescp$mean.performance, resrp$mean.performance))
mod = list(modelc, modelr, modelcp, modelrp)[[best]]
perfScatterPlot(parscores, mod, singleBest, ldf2, factor = 1, pargs=aes(colour = family)) + scale_x_log10() + scale_y_log10() + xlab("portfolio") + ylab("single best")
@

<<echo=FALSE, message=FALSE, warning=FALSE>>=
parallelStop()
@

\end{document}

### 2015-11-08 portfolio/convertToAslib.R

@@ -1,0 +1,112 @@
+library(aslib)
+library(BBmisc)
+library(plyr)
+library(reshape2)
+
+features.cheap = read.table("gpgnode-sip-cheap-features.data", header = TRUE, sep = " ", stringsAsFactors = FALSE)
+features.cheap$family = NULL
+names(features.cheap)[-1] = paste0("cheap.", names(features.cheap))[-1]
+features.distance = read.table("gpgnode-sip-distance-features.data", header = TRUE, sep = " ", stringsAsFactors = FALSE)
+features.distance$family = NULL
+names(features.distance)[-1] = paste0("distance.", names(features.distance))[-1]
+
+times = read.table("gpgnode-sip-runtimes.data", header = TRUE, sep = " ", stringsAsFactors = FALSE)
+times$family = NULL
+times$vbs = NULL
+times$sat = NULL
+
+timeout = 1e8/1000
+algorithms = names(times)[-1]
+
+feats.cheap.pattern = names(features.cheap)[grep("pattern.", names(features.cheap), fixed = TRUE)][-1]
+costs.cheap.pattern = "cheap.pattern.time"
+feats.cheap.target = names(features.cheap)[grep("target.", names(features.cheap), fixed = TRUE)][-1]
+costs.cheap.target = "cheap.target.time"
+
+feats.distance.pattern = names(features.distance)[grep("pattern.", names(features.distance), fixed = TRUE)][-1]
+costs.distance.pattern = "distance.pattern.time"
+feats.distance.target = names(features.distance)[grep("target.", names(features.distance), fixed = TRUE)][-1]
+costs.distance.target = "distance.target.time"
+
+# convert ms to s to avoid integer overflows later
+for (alg in algorithms) {
+    times[,alg] = times[,alg] / 1000
+    times[,alg][times[,alg] >= timeout] = NA
+}
+features.cheap[,costs.cheap.pattern] = features.cheap[,costs.cheap.pattern] / 1000
+features.cheap[,costs.cheap.target] = features.cheap[,costs.cheap.target] / 1000
+features.distance[,costs.distance.pattern] = features.distance[,costs.distance.pattern] / 1000
+features.distance[,costs.distance.target] = features.distance[,costs.distance.target] / 1000
+
+# convert to aslib
+desc = makeS3Obj("ASScenarioDesc",
+  scenario_id = "graphs-2015",
[…]

### 2015-11-11 portfolio/convertToAslib.R
add lad features

@@ -9,6 +9,8 @@ names(features.cheap)[-1] = paste0("cheap.", names(features.cheap))[-1]
 features.distance = read.table("gpgnode-sip-distance-features.data", header = TRUE, sep = " ", stringsAsFactors = FALSE)
 features.distance$family = NULL
 names(features.distance)[-1] = paste0("distance.", names(features.distance))[-1]
+features.lad = read.table("gpgnode-sip-lad-features.data", header = TRUE, sep = " ", stringsAsFactors = FALSE)
+features.lad$family = NULL
 
 times = read.table("gpgnode-sip-runtimes.data", header = TRUE, sep = " ", stringsAsFactors = FALSE)
 times$family = NULL
@@ -28,6 +30,10 @@ costs.distance.pattern = "distance.pattern.time"
 feats.distance.target = names(features.distance)[grep("target.", names(features.distance), fixed = TRUE)][-1]
 costs.distance.target = "distance.target.time"
 
+feats.lad = names(features.lad)[grep("lad.", names(features.lad), fixed = TRUE)][-c(length(names(features.lad))-1, length(names(features.lad)))]
+costs.lad = "lad.time"
+presolved.lad = "lad.detected.inconsistent"
+
 # convert ms to s to avoid integer overflows later
 for (alg in algorithms) {
     times[,alg] = times[,alg] / 1000
@@ -37,11 +43,12 @@ features.cheap[,costs.cheap.pattern] = features.cheap[,costs.cheap.pattern] / 10
 features.cheap[,costs.cheap.target] = features.cheap[,costs.cheap.target] / 1000
 features.distance[,costs.distance.pattern] = features.distance[,costs.distance.pattern] / 1000
 features.distance[,costs.distance.target] = features.distance[,costs.distance.target] / 1000
+features.lad[,costs.lad] = features.lad[,costs.lad] / 1000
 
 # convert to aslib
 desc = makeS3Obj("ASScenarioDesc",
   scenario_id = "graphs-2015",
-  features_deterministic = c(feats.cheap.pattern, feats.cheap.target, feats.distance.pattern, feats.distance.target),
+  features_deterministic = c(feats.cheap.pattern, feats.cheap.target, feats.distance.pattern, feats.distance.target, feats.lad),
   features_stochastic = character(0),
   algorithms_deterministic = algorithms,
   algorithms_stochastic = character(0),
[…]

### 2015-11-11 portfolio/convertToAslib.R
fix duplicate names

@@ -30,9 +30,10 @@ costs.distance.pattern = "distance.pattern.time"
 feats.distance.target = names(features.distance)[grep("target.", names(features.distance), fixed = TRUE)][-1]
 costs.distance.target = "distance.target.time"
 
-feats.lad = names(features.lad)[grep("lad.", names(features.lad), fixed = TRUE)][-c(length(names(features.lad))-1, length(names(features.lad)))]
+feats.lad = names(features.lad)[grep("lad.", names(features.lad), fixed = TRUE)]
 costs.lad = "lad.time"
 presolved.lad = "lad.detected.inconsistent"
+feats.lad = feats.lad[!(feats.lad %in% c(costs.lad, presolved.lad))]
 
 # convert ms to s to avoid integer overflows later
 for (alg in algorithms) {
@@ -63,13 +64,13 @@ desc = makeS3Obj("ASScenarioDesc",
   features_cutoff_memory = NA,
 
   number_of_feature_steps = 5,
-  default_steps = c("cheap_pattern", "cheap_target", "distance_pattern", "distance_target", "lad"),
+  default_steps = c("cheap_pattern", "cheap_target", "distance_pattern", "distance_target", "lad_features"),
   feature_steps =
       list(cheap_pattern = list(provides = feats.cheap.pattern),
            cheap_target = list(provides = feats.cheap.target),
            distance_pattern = list(provides = feats.distance.pattern),
            distance_target = list(provides = feats.distance.target),
-           lad = list(provides = feats.lad))
+           lad_features = list(provides = feats.lad))
 )
 
 mfeats = merge(merge(features.cheap, features.distance, by = "instance"), features.lad, by = "instance")
@@ -77,15 +78,15 @@ mfeats = merge(merge(features.cheap, features.distance, by = "instance"), featur
 feature.values = cbind(instance_id = mfeats$instance, repetition = 1, mfeats[, c(feats.cheap.pattern, feats.cheap.target, feats.distance.pattern, feats.distance.target, feats.lad)])
 feature.values$instance_id = as.character(feature.values$instance_id)
 
-feature.runstatus = data.frame(instance_id = mfeats$instance, repetition = 1, cheap_pattern = "ok", cheap_target = "ok", distance_pattern = "ok", distance_target = "ok", lad = ifelse(mfeats[, presolved.lad] == "true", "presolved", "ok"))
[…]

### 2015-11-12 portfolio/eda_config.R
add missing file

@@ -1,0 +1,39 @@
+library(aslib)
+
+makeEDAConfig = function(
+  algo.perf.impute.zero.vals = FALSE,
+  algo.perf.log = FALSE
+) {
+
+  makeS3Obj("ASScenarioHTMLConfig",
+    algo.perf.impute.zero.vals = algo.perf.impute.zero.vals,
+    algo.perf.log = algo.perf.log
+  )
+}
+
+print.ASScenarioHTMLConfig = function(x, ...) {
+  ns = names(x)
+  for (i in seq_along(x)) {
+   catf("%-30s : %s", ns[i], x[[i]])
+  }
+}
+
+# Sources config file and returns a config s3 object
+readEDAConfig = function(asscenario, confpath) {
+  checkArg(asscenario, "ASScenario")
+  id = asscenario$desc$scenario_id
+  conffile = file.path(confpath, paste0(id, ".R"))
+  if (file.exists(conffile)) {
+    # source config into envir, then construct
+    conf = new.env()
+    x = try(sys.source(conffile, envir = conf))
+    if (is.error(x)) {
+      stopf("There was an error in sourcing your configuration file '%s': %s!",
+        conffile, as.character(x))
+    }
+    conf = as.list(conf)
+    return(do.call(makeEDAConfig, conf))
+  } else {
+    stopf("Config file for scenario does not exist: %s", id)
+  }
+}

### 2015-11-12 portfolio/as-graphs.Rnw
analysis for presolved instances

@@
 <<echo=FALSE, results='asis'>>=
-source("convertToAslib.R")
+#source("convertToAslib.R")
 
@@
 
+\subsection*{Algorithm Selection Results with instances solved during
+computation of LAD features removed}
+
+<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
+ldf3 = ldf
+presolved = asscenario$feature.runstatus$instance_id[asscenario$feature.runstatus$lad_features == "presolved"]
+ldf3$data = subset(ldf$data, ldf$data$instance_id %in% presolved)
+ldf3$best = subset(ldf$best, ldf$data$instance_id %in% presolved)
+ldf3$train = NULL
+ldf3$test = NULL
+ldf3 = cvFolds(ldf3)
+timess2 = subset(timess, timess$instance %in% ldf3$data$instance_id)
+ldf3$data$family = factor(timess2[order(timess2$instance, as.character(ldf3$data$instance_id)), 'family'])
+ldf3$extra = "family"
+
+tab = t(sapply(ldf3$data[,ldf3$performance], mean))
+print(xtable(tab, caption = "Mean performance of each algorithm."), booktabs = TRUE, include.rownames = FALSE)
+
+tabs = by(ldf3$data, ldf3$data$family, function(x) as.data.frame.matrix(t(sapply(x[,ldf3$performance], mean))))
+tab = rbind.fill(tabs)
+tab$family = rownames(tab)
+tab = tab[,order(names(tab))]
+print(xtable(tab, digits = 3, caption = "Mean performance by family."), booktabs = TRUE, include.rownames = FALSE)
+
+tab = t(sapply(ldf3$data[,ldf3$performance], median))
+print(xtable(tab, caption = "Median performance of each algorithm."), booktabs = TRUE, include.rownames = FALSE)
+
+tabs = by(ldf3$data, ldf3$data$family, function(x) as.data.frame.matrix(t(sapply(x[,ldf3$performance], median))))
+tab = rbind.fill(tabs)
+tab$family = rownames(tab)
+tab = tab[,order(names(tab))]
+print(xtable(tab, digits = 3, caption = "Median performance by family."), booktabs = TRUE, include.rownames = FALSE)
+
+tab = t(table(unlist(ldf3$best)))
+print(xtable(tab, caption = "Number of times each algorithm is best. More than one algorithm may be best for one instance if there are ties."), booktabs = TRUE, include.rownames = FALSE)
+
+tabs = tapply(ldf3$best, ldf3$data$family, function(x) as.data.frame.matrix(t(table(unlist(x)))))
+tab = rbind.fill(tabs)
+tab[is.na(tab)] = 0
+tab$family = rownames(tab)
+tab = tab[,order(names(tab))]
+print(xtable(tab, digits = 0, caption = "Number of times each algorithm is best by family."), booktabs = TRUE, include.rownames = FALSE)
+@
+
+\clearpage
+
+<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
+resvbs = data.frame(model = "virtual best",
+    mean.performance = mean(parscores(ldf3, vbs, factor = 1)),
+    median.performance = median(parscores(ldf3, vbs, factor = 1)))
+ressb = data.frame(model = "single best",
+    mean.performance = mean(parscores(ldf3, singleBest, factor = 1)),
+    median.performance = median(parscores(ldf3, singleBest, factor = 1)))
+
+modelc = classify(makeLearner("classif.randomForest"), ldf3)
+resc = data.frame(model = "classification",
+    mean.performance = mean(parscores(ldf3, modelc, factor = 1)),
+    median.performance = median(parscores(ldf3, modelc, factor = 1)))
+times = sapply(modelc$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
+print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classification."), booktabs = TRUE)
+
+modelr = regression(makeLearner("regr.randomForest"), ldf3)
+resr = data.frame(model = "regression",
+    mean.performance = mean(parscores(ldf3, modelr, factor = 1)),
+    median.performance = median(parscores(ldf3, modelr, factor = 1)))
+times = sapply(modelr$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
+print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regression."), booktabs = TRUE)
+
+modelcp = classifyPairs(makeLearner("classif.randomForest"), ldf3)
+rescp = data.frame(model = "classifyPairs",
+    mean.performance = mean(parscores(ldf3, modelcp, factor = 1)),
+    median.performance = median(parscores(ldf3, modelcp, factor = 1)))
+times = sapply(modelcp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
+print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classificationPairs."), booktabs = TRUE)
+
+modelrp = regressionPairs(makeLearner("regr.randomForest"), ldf3)
+resrp = data.frame(model = "regressionPairs",
+    mean.performance = mean(parscores(ldf3, modelrp, factor = 1)),
+    median.performance = median(parscores(ldf3, modelrp, factor = 1)))
+times = sapply(modelrp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
+print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regressionPairs."), booktabs = TRUE)
+
+tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
+print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolver."), booktabs = TRUE, include.rownames = FALSE)
+@
+
+<<echo=FALSE,fig.cap='Single best algorithm vs.\ portfolio performance with presolved instances removed (best model).'>>=
+best = which.min(c(resc$mean.performance, resr$mean.performance, rescp$mean.performance, resrp$mean.performance))
+mod = list(modelc, modelr, modelcp, modelrp)[[best]]
+perfScatterPlot(parscores, mod, singleBest, ldf3, factor = 1, pargs=aes(colour = family)) + scale_x_log10() + scale_y_log10() + xlab("portfolio") + ylab("single best")
+@
+
+\clearpage
+
 \subsection*{Algorithm Selection Results with Presolver}
 
-Run vf2 for 0.05 second as a presolver.
+Run vf2 for 0.05 second as a presolver, after removing presolved instances.
 
@@
 # presolver
-ldf2 = ldf
+ldf2 = ldf3
 presolvetime = 0.05
-ldf2$data = subset(ldf$data, ldf$data$vf2 > presolvetime)
-ldf2$best = subset(ldf$best, ldf$data$vf2 > presolvetime)
+ldf2$data = subset(ldf3$data, ldf3$data$vf2 > presolvetime)
+ldf2$best = subset(ldf3$best, ldf3$data$vf2 > presolvetime)
 ldf2$train = NULL

### 2015-11-12 portfolio/as-graphs.Rnw
fix logic bug

@@
 presolved = asscenario$feature.runstatus$instance_id[asscenario$feature.runstatus$lad_features == "presolved"]
-ldf3$data = subset(ldf$data, ldf$data$instance_id %in% presolved)
-ldf3$best = subset(ldf$best, ldf$data$instance_id %in% presolved)
+ldf3$data = subset(ldf$data, !(ldf$data$instance_id %in% presolved))
+ldf3$best = subset(ldf$best, !(ldf$data$instance_id %in% presolved))
 ldf3$train = NULL

### 2015-11-13 lad_stats/LADpreprocessing.c
Updated LAD stats

@@ -136,26 +136,24 @@ int main(int argc, char* argv[]){
 	Tgraph *Gt = createGraph(fileNameGt);       // Target graph
     
     struct rusage ru;     // reusable structure to get CPU time usage
+    
+    float minReduction = 100.0;
+    float maxReduction = 0.0;
     if ((Gp->nbVertices > Gt->nbVertices) || (Gp->maxDegree > Gt->maxDegree)){
         getrusage(RUSAGE_SELF, &ru);
         printf("Inconsistency detected in %d.%06d seconds\n",
                (int) ru.ru_utime.tv_sec, (int) ru.ru_utime.tv_usec);
     }
-    else if (Gt->minDegree == Gt->maxDegree){
-        if (Gp->maxDegree <= Gt->maxDegree){
-            getrusage(RUSAGE_SELF, &ru);
-            printf("0 removed values (percentage = 0) in %d.%06d seconds\n",(int) ru.ru_utime.tv_sec, (int) ru.ru_utime.tv_usec);
-        }
-        else{
-            getrusage(RUSAGE_SELF, &ru);
-            printf("Inconsistency detected in %d.%06d seconds\n",
-                   (int) ru.ru_utime.tv_sec, (int) ru.ru_utime.tv_usec);
-        }
+    else if (Gp->maxDegree <= Gt->minDegree){
+        getrusage(RUSAGE_SELF, &ru);
+        printf("0 removed values (percentage = 0) in %d.%06d seconds\n",(int) ru.ru_utime.tv_sec, (int) ru.ru_utime.tv_usec);
     }
     else{
         int cpt = 0;
         int mu[Gt->maxDegree+1];
         int** mv = (int**)malloc(Gt->nbVertices*sizeof(int*));
+        bool isUsed[Gt->nbVertices];
+        memset(isUsed,false,(Gt->nbVertices)*sizeof(bool));
         for (int v=0; v<Gt->nbVertices; v++){
             mv[v] = (int*)calloc(Gt->maxDegree+1,sizeof(int));
             memset(mv[v],0,(Gt->maxDegree+1)*sizeof(int));
@@ -163,25 +161,34 @@ int main(int argc, char* argv[]){
         }
         for (int u=0; u<Gp->nbVertices; u++){
             int domainSize = 0;
+            int val;
             memset(mu,0,(Gt->maxDegree+1)*sizeof(int));
             for (int i=0; i<Gp->nbAdj[u]; i++) mu[Gp->nbAdj[Gp->adj[u][i]]]++;
[…]

### 2015-11-30 paper/paper.tex
update abstract

@@
 \begin{abstract}
-There are several competitive algorithms for the subgraph isomorphism problem, and one of these
-algorithms additionally has a choice of parameter. As with many other hard problems, the single best
-choice of algorithm overall is rarely the best algorithm on a per-instance basis. We show that using
-witchcraft to pick from a portfolio of algorithms on an instance by instance basis provides super
-dooper amazingly quite OKish results, most of the time, and opens up some flexibility in algorithm
-design.
+The subgraph isomorphism problem is a computationally challenging problem with
+lots of important practical applications, for example in computer vision,
+biochemistry, and model checking. There are a number of state-of-the-art
+algorithms for solving it, each of which has its own performance
+characteristics. As with many other hard problems, the single best choice of
+algorithm overall is rarely the best algorithm for each instance. We develop a
+portfolio approach that leverages novel features to characterise subgraph
+isomorphism problems to dynamically decide which algorithm to use on a
+per-instance basis. We demonstrate significant performance improvements on a
+large set of hard benchmark problems. In addition, we show how algorithm
+selection models can be leveraged to gain new insights into what affects the
+performance of an algorithm.
 \end{abstract}

### 2015-11-30 portfolio/as-graphs.Rnw
update experimental data

@@
 library(parallelMap)
-parallelStartSocket(4)
+parallelStartSocket(5)
 parallelLibrary("llama", "mlr")
@@
 
-modelc = classify(makeLearner("classif.randomForest"), ldf)
+rflearner = makeLearner("classif.randomForest")
+
+modelc = classify(rflearner, ldf)
 resc = data.frame(model = "classification",
@@
 
-modelcp = classifyPairs(makeLearner("classif.randomForest"), ldf)
+modelcp = classifyPairs(rflearner, ldf)
 rescp = data.frame(model = "classifyPairs",
@@
 <<echo=FALSE,fig.cap='Single best algorithm vs.\ portfolio performance (best model).'>>=
-best = which.min(c(resc$mean.performance, resr$mean.performance, rescp$mean.performance, resrp$mean.performance))
-mod = list(modelc, modelr, modelcp, modelrp)[[best]]
+best = which.min(c(resc$mean.performance))
+mod = list(modelc)[[best]]
 perfScatterPlot(parscores, mod, singleBest, ldf, addCostsx = TRUE, addCostsy = FALSE, factor = 1, pargs=aes(colour = family)) + scale_x_log10() + scale_y_log10() + xlab("portfolio") + ylab("single best")
@@
 
-modelc = classify(makeLearner("classif.randomForest"), ldf3)
+modelc = classify(rflearner, ldf3)
 resc = data.frame(model = "classification",
@@
 
-modelcp = classifyPairs(makeLearner("classif.randomForest"), ldf3)
+modelcp = classifyPairs(rflearner, ldf3)
 rescp = data.frame(model = "classifyPairs",
@@
 tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
-print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolver."), booktabs = TRUE, include.rownames = FALSE)
+print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolved instances removed."), booktabs = TRUE, include.rownames = FALSE)
 @
@@
 
-modelc = classify(makeLearner("classif.randomForest"), ldf2)
+modelc = classify(rflearner, ldf2)
 resc = data.frame(model = "classification",
@@
 
-modelcp = classifyPairs(makeLearner("classif.randomForest"), ldf2)
+modelcp = classifyPairs(rflearner, ldf2)
 rescp = data.frame(model = "classifyPairs",

### 2015-11-30 paper/paper.tex
some background, experimental setup, results

@@
 \usepackage{cleveref}                  % no need to type Figure etc
+
+\usepackage{todonotes}
+\usepackage{booktabs}
 
@@
 
-We could try to come up with rules to pick an algorithm on an instance by instance basis using our
-knowledge of how the algorithms behave: for example, maybe we should use the Glasgow algorithm with
-paths of length 2 with large target graphs (but how large?), and only use length 3 on smaller
-graphs. Also, we could guess that LAD might do better than Glasgow on smaller, heavily-structured
-graphs. However, this is not very systematic.
-
-?? Lars: a paragraph of fluff about instance selection please.
-
-We show that picking algorithms and parameter choices on an instance by instance basis, by
-considering features of the pattern and target graphs, gives better overall performance than any
-single algorithm. This is the first time this has been done for subgraph isomorphism, and we must
-consider an unusual factor: our features must look at pairs of inputs, and the relationships between
-them.
-
-?? Lars: do you know of any other problems where you have two inputs rather than one? Or is this
-novel?
+We show that by choosing algorithms on an instance by instance basis, by
+considering features of the pattern and target graphs, we achieve better overall
+performance than any single algorithm.
 
@@
 pattern density are?
+
+\section{Background}
+
+\todo[inline]{Some background on subgraph isomorphism algorithms, maybe move
+some stuff from the intro here.}
+
+The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to
+select from an algorithm
+portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected
+to perform best on a given problem instance. Algorithm selection systems usually
+build machine learning models of the algorithms or the portfolio they are
+contained in to forecast which algorithm to use in a particular context. Using
+the model predictions, one or more algorithms from the portfolio are selected to
+be run sequentially or in parallel.
+
+Here, we consider the case where exactly one algorithm is selected for
+solving the problem. One of the most prominent and successful systems that
+employs this approach is SATzilla~\cite{xu_satzilla_2008}, which defined the
+state of the art in SAT solving for a number of years. Other application areas
+include constraint solving~\cite{omahony_using_2008},
+the travelling salesperson problem~\cite{kotthoff_improving_2015},
+and AI planning~\cite{seipp_learning_2012}.
+
+The interested reader is referred to a recent
+survey~\cite{kotthoff_algorithm_2014} for additional information on algorithm
+selection.
 
@@
 timeout of $10^8$ milliseconds (a little over a day).
-
-?? Lars: your stuff here.
 
@@
 
-\noindent Note that this is not like SAT, where we can reasonably split things up based upon which
-family they're from.
+We consider a total of 5725 instances.
 
-\subsection{Machine learning voodoo stuff}
+\subsubsection{Features}
 
-\subsection{Results}
+\todo[inline]{Describe features here.}
 
-?? Can we do better than picking the single best on a family by family basis? Bear in mind that we
-don't have the family as a usable feature.
+This gives us a total of 30 features.
+
+The data we use in this paper is available as Aslib scenario GRAPHS-2015.
+
+\subsection{Algorithm Selection approach}
+
+We apply two static steps to each instance before solving it. First, the
+computation of the LAD features can detect unsatisfiability of an instance. If
+an instance is identified as unsatisfiable, we do not process it further.
+Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
+without the overhead of running algorithm selection and avoids potentially
+making incorrect solver choices.
+
+Computing the LAD features detects unsatisfiability on 1919 instances. Running
+VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
+instances that we consider for algorithm selection.
+
+To decide which solver to run on the remaining instances, we use
+LLAMA~\cite{kotthoff_llama_2013}. In preliminary experiments, we identified
+random forest regression to predict the performance difference between pairs of
+algorithms, with the best algorithm identified as the one with the best
+cumulative performance difference over all pairs, as the best approach. We make
+no claims that this is the overall best approach.
+
+\section{Results}
+
+Table~\ref{tab:res} shows the performance of our algorithm selection approach,
+compared to two baselines. The virtual best solver is the oracle predictor that,
+for each instance, chooses the best solver from our portfolio. This is the upper
+bound of what an algorithm selection approach can achieve. The single best
+solver is the one solver from the portfolio that has the overall best
+performance across the entire set of instances. We consider it a lower bound on
+performance.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lr}
+  \toprule
+& mean runtime\\
+  \midrule
+virtual best & 5822.809\\
+  single best & 7782.569\\
+  LLAMA & 6595.275\\
+   \bottomrule
+\end{tabular}
+\vspace{1ex}
+\caption{Algorithm selection performance on reduced set of 2336
+instances.}\label{tab:res}
+\end{table}
+
+We are able to close more than 60\% of the performance gap between the single
+best and the virtual best solver, our lower and upper bounds. For comparison, we
+show the results achieved on the entire set of instances without
+unsatisfiability detection and presolving in Table~\ref{tab:resfull}. We achieve
+similar results with more than 58\% of the gap between the single best and the
+virtual best closed.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lr}
+  \toprule
+& mean runtime\\
+  \midrule
+virtual best & 2375.913\\
+  single best & 3177.569\\
+  LLAMA & 2706.330\\
+   \bottomrule
+\end{tabular}
+\vspace{1ex}
+\caption{Algorithm selection performance on the full set of 5725
+instances.}\label{tab:resfull}
+\end{table}
+
+Figure~\ref{fig:scatter} shows detailed results for each of the 2336 instances
+in the reduced set. Our algorithm selection approach is by no means able to
+identify the best solver on a majority of instances, although we are able to
+achieve performance improvements overall. There seems to be no particular
+pattern as to where algorithm selection performs well and where it fails.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figures/perfScatter}
+\caption{Single best algorithm vs.\ portfolio performance on all instances by
+family. Points above the diagonal indicate that the portfolio achieved better
+performance, while the single best solver was better on instances below the
+diagonal.}
+\label{fig:scatter}
+\end{figure}
 

### 2015-12-05 paper/paper.tex
Chrisine's changes

--- 2015-11-30 paper/paper.tex
@@
 \usepackage{cleveref}                  % no need to type Figure etc
+\usepackage{xcolor,colortbl}
 
@@
 \begin{abstract}
-The subgraph isomorphism problem is a computationally challenging problem with
-lots of important practical applications, for example in computer vision,
-biochemistry, and model checking. There are a number of state-of-the-art
-algorithms for solving it, each of which has its own performance
-characteristics. As with many other hard problems, the single best choice of
-algorithm overall is rarely the best algorithm for each instance. We develop a
-portfolio approach that leverages novel features to characterise subgraph
-isomorphism problems to dynamically decide which algorithm to use on a
-per-instance basis. We demonstrate significant performance improvements on a
-large set of hard benchmark problems. In addition, we show how algorithm
-selection models can be leveraged to gain new insights into what affects the
-performance of an algorithm.
-\end{abstract}
+Subgraph isomorphism is a computationally challenging problem with lots of important practical applications, for example in computer vision, biochemistry, and model checking.
+There are a number of state-of-the-art algorithms for solving it, each of which has its own performance characteristics.
+As with many other hard problems, the single best choice of algorithm overall is rarely the best algorithm for each instance.
+We develop a portfolio approach that leverages novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm to use on a per-instance basis.
+We demonstrate significant performance improvements on a large set of hard benchmark problems.
+In addition, we show how algorithm selection models can be leveraged to gain new insights into what affects the performance of an algorithm.\end{abstract}
 
@@
 
-The \emph{(non-induced) subgraph isomorphism problem} is to find an injective mapping from a small
-\emph{pattern} graph to a large \emph{target} graph which preserves adjacency. This problem is
-\NP-complete, but applications in chemistry and computer vision and model checking and some other
-things we should cite mean various exact algorithms exist:
+The subgraph isomorphism problem is to find an injective mapping from vertices of a small \emph{pattern} graph to vertices of a large \emph{target} graph which preserves adjacency.
+This \NP-complete problem has lots of important practical applications, for example in computer vision, biochemistry, and model checking.
+There exist various exact algorithms, which have been compared on a large benchmark in  \cite{McCreesh:2015}.
+These experiments indicated that the single best algorithm depends on the CPU time limit considered: for very small (resp.
+larger) CPU time limits, VF2 \cite{Cordella:2004} (resp.
+the Glasgow algorithm \cite{McCreesh:2015}) has better success rates.
+Furthermore, they also showed us that on an instance by instance basis, other algorithms may become better.
 
+In this paper, we report experimental results on a larger set of 8 algorithms, including new variants, and a larger benchmark.
+We show that 5 of these algorithms are single best algorithms, depending on the CPU time limit considered, and that combining preprocessings with an algorithm selection on an instance by instance basis allows us to achieve better overall performance than any single algorithm.
+
+\section{Algorithms for subgraph isomorphism}
+
+We consider the following exact algorithms:
+ VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially fast on trivially satisfiable instances;
+ LAD \cite{Solnon:2010}, which  does very strong reasoning based on local all different redundant constraints;
+ and the Glasgow algorithm \cite{McCreesh:2015}, which does an expensive preprocessing pass once, followed by weaker reasoning during search.
+ The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about non-adjacent vertices.
+In experiments reported in \cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive preprocessing on larger, denser instances.
+This is
+often not the best choice on an instance by instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths of length 3 is relatively expensive on some graphs.
+Therefore, in this new study, we consider 4 different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.
+
+We introduce two new variants of LAD.
+The first variant is a weaker LAD filtering (called LAD preprocessing) which is applied once, without performing a backtracking search, and detects inconsistencies on many instances: For each pattern node $u$, we check that there exists at least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
+LAD preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: When it does not detect inconsistency, the instance may either be consistent or not.
+The time complexity of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target nodes, respectively, and $p$ is the number of target edges.
+
+The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in  \cite{Solnon:2010}.
+It combines LAD filtering with new invariant properties borrowed from SND \cite{Audemard:2014}.
+The idea is to label each edge $(u,v)$ with the number of paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$, and to add the constraint that the label of a pattern node (resp.
+edge) must be smaller than or equal to the label of its associated target node (resp.
+edge).
+
+\section{Benchmark}
+
+We consider 5725 instances which are available in a simple text format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12 classes.
 \begin{itemize}
-    \item VF2 \cite{Cordella:2004} is a backtracking search algorithm which is especially fast on
-        trivially satisfiable instances (which are common in certain application areas).
-    \item LAD \cite{Solnon:2010} and SND \cite{Audemard:2014} do very strong reasoning, at the
-        expense of making few guesses per second when working with larger graphs.
-    \item Most recently, the Glasgow algorithm \cite{McCreesh:2015} does an expensive preprocessing
-        pass once, followed by weaker reasoning during search.
+\item Class 1 contains randomly generated scale-free graphs \cite{}.
+\item Classes 2 and 3 contain various kinds of graphs coming from \cite{}: class 2 contains small instances generated from the 50 first graphs of the benchmark, whereas class 3 contains larger instances where pattern graphs are coming from class 2, and target graphs are coming from the next 50 graphs of the benchmark.
+\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}: bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
+\item Classes 9 and 10 contain instances coming from images \cite{}.
+\item Class 11 contains instances coming from meshes modeling 3D objects \cite{}.
+\item Class 12 contains random graph instances belonging to the phase transition.
 \end{itemize}
+Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.
 
-\noindent The experiments by the Glasgow authors indicated that overall, the Glasgow algorithm is
-the single best algorithm when evaluated across a wide range of problem instances. However, they
-also observed that on an instance by instance basis, it is often not the single best.  Additionally,
-the Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about
-non-adjacent vertices.  The choice of paths of length 3 was used as a reasonable compromise---longer
-paths lead to prohibitively expensive preprocessing on larger, denser instances. Again, this is
-often not the best choice on an instance by instance basis: sometimes path-based reasoning gives no
-benefit at all, sometimes considering only paths of length 2 suffices, occasionally paths of
-length 4 are helpful, and even looking at paths of length 3 is relatively expensive on some graphs.
+\section{Experimental comparison}
 
-?? Also, we can use supplemental graphs in LAD. Christine: can you explain what you implemented?
+Our runtimes were measured on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running Scientific Linux 6.5.
+We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015}, the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2 \cite{Cordella:2004}.
+Software was compiled using GCC 4.9.
+Each problem instance was run with a timeout of $10^8$ milliseconds (a little over a day).
 
-We could try to come up with rules to pick an algorithm on an instance by instance basis using our
-knowledge of how the algorithms behave: for example, maybe we should use the Glasgow algorithm with
-paths of length 2 with large target graphs (but how large?), and only use length 3 on smaller
-graphs. Also, we could guess that LAD might do better than Glasgow on smaller, heavily-structured
-graphs. However, this is not very systematic.
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
+\hline
+$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
+&&pre&def&new&1&2&3&4&&\\\hline
+$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
+$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
+$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
+$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
+$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
+$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
+$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
+$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
+$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
+\end{tabular}
+\end{center}
+\caption{Number of solved instances at different CPU time limits: Each line displays a time limit $t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and 4, and the VBS.
+The best single algorithm is highlighted in blue, and the difference between VBS and single best if displayed in the last column ($\delta$).\label{expTime}}
+\end{table}
 
-?? Lars: a paragraph of fluff about instance selection please.
+Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging from 1 to $10^8$ milliseconds.
+It shows us that the best single solver depends on the time limit considered.
+Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very quickly, in a few milliseconds.
+However, they are not able to solve harder instances: LAD preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2 performs a basic backtracking search and is very fast on easy instances, because it does not compute expensive invariants.
+Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2 and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow 1, and 100 (resp.
+1000 and 10000) ms for Glasgow 2 (resp.
+3 and 4).
+The Virtual Best Solver (VBS), which considers the best algorithm for each instance separately, obtains much better results, showing us that the algorithms have complementary performance.
+The difference between VBS and single best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU time limits.
+In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the limit is 100 ms.
+In many applications, it is important to have the fastest possible algorithm.
+For example, in pattern recognition applications, we often have to solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern image in a large database of target images, for example), so that having an algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference.
+Therefore, it is important to select the best algorithm for each instance.
 
-We show that picking algorithms and parameter choices on an instance by instance basis, by
-considering features of the pattern and target graphs, gives better overall performance than any
-single algorithm. This is the first time this has been done for subgraph isomorphism, and we must
-consider an unusual factor: our features must look at pairs of inputs, and the relationships between
-them.
+Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class instances are comin from, unlike SAT.
+For all classes, there are always at least 2 algorithms which are the best for at least one instance of the class.
+In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for Class 2).
 
-?? Lars: do you know of any other problems where you have two inputs rather than one? Or is this
-novel?
+\begin{table}
+\begin{center}
+\begin{tabular}{|c||r||r|r|r||r|r|r|r|}
+\hline
+Class & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c|}{Glasgow}\\
+&&pre&def&new&1&2&3&4\\\hline
+1\\\hline
+2\\\hline
+3\\\hline
+4\\\hline
+5\\\hline
+6\\\hline
+7\\\hline
+8\\\hline
+9\\\hline
+10\\\hline
+11\\\hline
+12\\\hline
 
-?? Can we get better results if we explicitly consider, say, how close the target density and
-pattern density are?
+\end{tabular}
+\end{center}
+\caption{Number of times each algorithm is best, for each class.\label{expClass}}
+\end{table}
 
-\section{Experiments}
+\section{Algorithm selection approach}
 
-Our runtimes were measured on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
-Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
-the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
-\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
-timeout of $10^8$ milliseconds (a little over a day).
+Describe features + selection approach.
 
-?? Lars: your stuff here.
+Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the upper bound of what an algorithm selection approach can achieve), and the single best solver, at different CPU time limits (the single best corresponds to results highlighted in blue in Table 1).
+Comment Table 3: Mean runtime but also number of solved instances at different CPU time limits.
+Logically, LLAMA may solve less instances than single best for small time limits because it spends time to compute features and choose a solver, but it should be better for larger time limits...
 
-\subsection{Problem instances}
+\begin{table}
+\begin{tabular}{|l|r|rrrrrrrrr|}
+&mean & \multicolumn{9}{c|}{Number of solved instances}\\
+&runtime & $10^0$ &  $10^1$ &  $10^2$ &  $10^3$ &  $10^4$ &  $10^5$ &  $10^6$ &  $10^7$ &  $10^8$\\\hline
+VBS & 2375.9 & 1132 & 3388 & 4636 & 5170 & 5332 & 5434 & 5500 & 5567 & 5608\\\hline
+Single best & 3177.6 & 1126 & 2287 & 3943 & 4908 & 5253 & 5377 & 5452 & 5518 & 5561\\\hline
+LLAMA & 2706.3\\\hline
+\end{tabular}
+\caption{}
+\end{table}
 
-All instances are available in a simple text format
-\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}.
-
-Real-world and application-derived graphs:
-
-\begin{itemize}
-    \item The LV graphs (family 2). Also, the same targets, but with larger pattern graphs (family
-        11, Multiple edges have been combined into a single edge.)
-    \item CVIU11 images (family 8)
-    \item CVIU11 meshes (family 10)
-    \item PR15 (family 9)
-\end{itemize}
-
-\noindent Randomly generated instances with particular properties:
-
-\begin{itemize}
-    \item Bounded-degree random graphs (family 3, 4) \cite{GraphDatabase}
-    \item Random 4d mesh graphs (family 5, 6) \cite{GraphDatabase}
-    \item Random graphs (family 7) \cite{GraphDatabase}
-    \item Randomly generated scale-free graphs. (family 1) \cite{Solnon:2010}
-    \item Phase transition (family 12)
-\end{itemize}
-
-\noindent Note that this is not like SAT, where we can reasonably split things up based upon which
-family they're from.
-
-\subsection{Machine learning voodoo stuff}
-
-\subsection{Results}
-
-?? Can we do better than picking the single best on a family by family basis? Bear in mind that we
-don't have the family as a usable feature.
+Add Figure 1 which compares single best at $10^8$ ms (Glasgow 2) with LLAMA ?
 
 \section{Conclusion and Future Work}
-
-This stuff works respectably well, considering we're dealing with two inputs rather than one.
-
-We have only looked at sequential algorithms: the Glasgow algorithm has a parallel implementation,
-and a similar technique could easily be used to parallelise LAD.  We have also not considered the
-induced variant of the problem, nor instances involving labels or directed edges.  We could also
-look at other encodings. For example, a direct SAT requires $O(v^4)$ clauses, but may be good in
-certain situations.
 

### 2015-12-06 paper/paper.tex

@@
 \usepackage{xcolor,colortbl}
+
+\usepackage{todonotes}
+\usepackage{booktabs}
 
@@
 \begin{abstract}
-Subgraph isomorphism is a computationally challenging problem with lots of important practical applications, for example in computer vision, biochemistry, and model checking.
-There are a number of state-of-the-art algorithms for solving it, each of which has its own performance characteristics.
-As with many other hard problems, the single best choice of algorithm overall is rarely the best algorithm for each instance.
-We develop a portfolio approach that leverages novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm to use on a per-instance basis.
-We demonstrate significant performance improvements on a large set of hard benchmark problems.
-In addition, we show how algorithm selection models can be leveraged to gain new insights into what affects the performance of an algorithm.\end{abstract}
+Subgraph isomorphism is a computationally challenging problem with lots of important practical
+applications, for example in computer vision, biochemistry, and model checking. There are a number
+of state-of-the-art algorithms for solving it, each of which has its own performance
+characteristics. As with many other hard problems, the single best choice of algorithm overall is
+rarely the best algorithm for each instance. We develop a portfolio approach that leverages novel
+features to characterise subgraph isomorphism problems to dynamically decide which algorithm to use
+on a per-instance basis. We demonstrate significant performance improvements on a large set of hard
+benchmark problems. In addition, we show how algorithm selection models can be leveraged to gain new
+insights into what affects the performance of an algorithm.
+\end{abstract}
 
@@
 
-The subgraph isomorphism problem is to find an injective mapping from vertices of a small \emph{pattern} graph to vertices of a large \emph{target} graph which preserves adjacency.
-This \NP-complete problem has lots of important practical applications, for example in computer vision, biochemistry, and model checking.
-There exist various exact algorithms, which have been compared on a large benchmark in  \cite{McCreesh:2015}.
-These experiments indicated that the single best algorithm depends on the CPU time limit considered: for very small (resp.
-larger) CPU time limits, VF2 \cite{Cordella:2004} (resp.
-the Glasgow algorithm \cite{McCreesh:2015}) has better success rates.
-Furthermore, they also showed us that on an instance by instance basis, other algorithms may become better.
+The subgraph isomorphism problem is to find an injective mapping from vertices of a small
+\emph{pattern} graph to vertices of a large \emph{target} graph which preserves adjacency. This
+\NP-complete problem has lots of important practical applications, for example in computer vision,
+biochemistry, and model checking. There exist various exact algorithms, which have been compared on
+a large benchmark in  \cite{McCreesh:2015}. These experiments indicated that the single best
+algorithm depends on the CPU time limit considered: for very small (resp. larger) CPU time limits,
+VF2 \cite{Cordella:2004} (resp. the Glasgow algorithm \cite{McCreesh:2015}) has better success
+rates. Furthermore, they also showed us that on an instance by instance basis, other algorithms may
+become better.
 
-In this paper, we report experimental results on a larger set of 8 algorithms, including new variants, and a larger benchmark.
-We show that 5 of these algorithms are single best algorithms, depending on the CPU time limit considered, and that combining preprocessings with an algorithm selection on an instance by instance basis allows us to achieve better overall performance than any single algorithm.
+In this paper, we report experimental results on a larger set of 8 algorithms, including new
+variants, and a larger benchmark. We show that 5 of these algorithms are single best algorithms,
+depending on the CPU time limit considered, and that combining preprocessing with an algorithm
+selection on an instance by instance basis allows us to achieve better overall performance than any
+single algorithm.
+
+\section{Background}
+
+\todo[inline]{Some background on subgraph isomorphism algorithms, maybe move
+some stuff from the intro here.}
+
+The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
+algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
+best on a given problem instance. Algorithm selection systems usually build machine learning models
+of the algorithms or the portfolio they are contained in to forecast which algorithm to use in a
+particular context. Using the model predictions, one or more algorithms from the portfolio are
+selected to be run sequentially or in parallel.
+
+Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
+the most prominent and successful systems that employs this approach is
+SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
+years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
+salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
+
+The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
+information on algorithm selection.
 
@@
 
-We consider the following exact algorithms:
- VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially fast on trivially satisfiable instances;
- LAD \cite{Solnon:2010}, which  does very strong reasoning based on local all different redundant constraints;
- and the Glasgow algorithm \cite{McCreesh:2015}, which does an expensive preprocessing pass once, followed by weaker reasoning during search.
- The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about non-adjacent vertices.
-In experiments reported in \cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive preprocessing on larger, denser instances.
-This is
-often not the best choice on an instance by instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths of length 3 is relatively expensive on some graphs.
-Therefore, in this new study, we consider 4 different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.
+We consider the following exact algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking
+search algorithm, especially fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which
+does very strong reasoning based on local all different redundant constraints; and the Glasgow
+algorithm \cite{McCreesh:2015}, which does an expensive preprocessing pass once, followed by weaker
+reasoning during search.
 
-We introduce two new variants of LAD.
-The first variant is a weaker LAD filtering (called LAD preprocessing) which is applied once, without performing a backtracking search, and detects inconsistencies on many instances: For each pattern node $u$, we check that there exists at least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
-LAD preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: When it does not detect inconsistency, the instance may either be consistent or not.
-The time complexity of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target nodes, respectively, and $p$ is the number of target edges.
+ The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning
+ about non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice of paths
+ of length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive
+ preprocessing on larger, denser instances. This is often not the best choice on an instance by
+ instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only
+ paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths
+ of length 3 is relatively expensive on some graphs. Therefore, in this new study, we consider 4
+ different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.
 
-The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in  \cite{Solnon:2010}.
-It combines LAD filtering with new invariant properties borrowed from SND \cite{Audemard:2014}.
-The idea is to label each edge $(u,v)$ with the number of paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$, and to add the constraint that the label of a pattern node (resp.
-edge) must be smaller than or equal to the label of its associated target node (resp.
-edge).
+We introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
+preprocessing) which is applied once, without performing a backtracking search, and detects
+inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
+one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$
+of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
+preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: When
+it does not detect inconsistency, the instance may either be consistent or not. The time complexity
+of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
+nodes, respectively, and $p$ is the number of target edges.
+
+The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in
+\cite{Solnon:2010}. It combines LAD filtering with new invariant properties borrowed from SND
+\cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
+between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
+and to add the constraint that the label of a pattern node (resp. edge) must be smaller than or
+equal to the label of its associated target node (resp. edge).
 
@@
 
-We consider 5725 instances which are available in a simple text format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12 classes.
+We consider 5725 instances which are available in a simple text
+format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
+classes.
+
 \begin{itemize}
@@
 \end{itemize}
+
+\noindent Randomly generated instances with particular properties:
+
+\begin{itemize}
+    \item Bounded-degree random graphs (family 3, 4) \cite{GraphDatabase}
+    \item Random 4d mesh graphs (family 5, 6) \cite{GraphDatabase}
+    \item Random graphs (family 7) \cite{GraphDatabase}
+    \item Randomly generated scale-free graphs. (family 1) \cite{Solnon:2010}
+    \item Phase transition (family 12)
+\end{itemize}
+
+We consider a total of 5725 instances.
+
+\subsubsection{Features}
+
+\todo[inline]{Describe features here.}
+
+This gives us a total of 30 features.
+
+The data we use in this paper is available as Aslib scenario GRAPHS-2015.
+
+\subsection{Algorithm Selection approach}
+
+We apply two static steps to each instance before solving it. First, the
+computation of the LAD features can detect unsatisfiability of an instance. If
+an instance is identified as unsatisfiable, we do not process it further.
+Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
+without the overhead of running algorithm selection and avoids potentially
+making incorrect solver choices.
+
+Computing the LAD features detects unsatisfiability on 1919 instances. Running
+VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
+instances that we consider for algorithm selection.
+
+To decide which solver to run on the remaining instances, we use
+LLAMA~\cite{kotthoff_llama_2013}. In preliminary experiments, we identified
+random forest regression to predict the performance difference between pairs of
+algorithms, with the best algorithm identified as the one with the best
+cumulative performance difference over all pairs, as the best approach. We make
+no claims that this is the overall best approach.
+
+\section{Results}
+
+Table~\ref{tab:res} shows the performance of our algorithm selection approach,
+compared to two baselines. The virtual best solver is the oracle predictor that,
+for each instance, chooses the best solver from our portfolio. This is the upper
+bound of what an algorithm selection approach can achieve. The single best
+solver is the one solver from the portfolio that has the overall best
+performance across the entire set of instances. We consider it a lower bound on
+performance.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lr}
+  \toprule
+& mean runtime\\
+  \midrule
+virtual best & 5822.809\\
+  single best & 7782.569\\
+  LLAMA & 6595.275\\
+   \bottomrule
+\end{tabular}
+\vspace{1ex}
+\caption{Algorithm selection performance on reduced set of 2336
+instances.}\label{tab:res}
+\end{table}
+
+We are able to close more than 60\% of the performance gap between the single
+best and the virtual best solver, our lower and upper bounds. For comparison, we
+show the results achieved on the entire set of instances without
+unsatisfiability detection and presolving in Table~\ref{tab:resfull}. We achieve
+similar results with more than 58\% of the gap between the single best and the
+virtual best closed.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lr}
+  \toprule
+& mean runtime\\
+  \midrule
+virtual best & 2375.913\\
+  single best & 3177.569\\
+  LLAMA & 2706.330\\
+   \bottomrule
+\end{tabular}
+\vspace{1ex}
+\caption{Algorithm selection performance on the full set of 5725
+instances.}\label{tab:resfull}
+\end{table}
+
+Figure~\ref{fig:scatter} shows detailed results for each of the 2336 instances
+in the reduced set. Our algorithm selection approach is by no means able to
+identify the best solver on a majority of instances, although we are able to
+achieve performance improvements overall. There seems to be no particular
+pattern as to where algorithm selection performs well and where it fails.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figures/perfScatter}
+\caption{Single best algorithm vs.\ portfolio performance on all instances by
+family. Points above the diagonal indicate that the portfolio achieved better
+performance, while the single best solver was better on instances below the
+diagonal.}
+\label{fig:scatter}
+\end{figure}
+
 Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.
@@
 
-Our runtimes were measured on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running Scientific Linux 6.5.
-We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015}, the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2 \cite{Cordella:2004}.
-Software was compiled using GCC 4.9.
-Each problem instance was run with a timeout of $10^8$ milliseconds (a little over a day).
+Our runtimes were measured on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
+Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
+the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
+\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
+timeout of $10^8$ milliseconds (a little over a day).
 
@@
 \end{center}
-\caption{Number of solved instances at different CPU time limits: Each line displays a time limit $t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and 4, and the VBS.
-The best single algorithm is highlighted in blue, and the difference between VBS and single best if displayed in the last column ($\delta$).\label{expTime}}
+\caption{Number of solved instances at different CPU time limits: Each line displays a time limit
+$t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD
+preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and
+4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
+single best if displayed in the last column ($\delta$).\label{expTime}}
 \end{table}
 
-Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging from 1 to $10^8$ milliseconds.
-It shows us that the best single solver depends on the time limit considered.
-Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very quickly, in a few milliseconds.
-However, they are not able to solve harder instances: LAD preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2 performs a basic backtracking search and is very fast on easy instances, because it does not compute expensive invariants.
-Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2 and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow 1, and 100 (resp.
-1000 and 10000) ms for Glasgow 2 (resp.
-3 and 4).
-The Virtual Best Solver (VBS), which considers the best algorithm for each instance separately, obtains much better results, showing us that the algorithms have complementary performance.
-The difference between VBS and single best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU time limits.
-In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the limit is 100 ms.
-In many applications, it is important to have the fastest possible algorithm.
-For example, in pattern recognition applications, we often have to solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern image in a large database of target images, for example), so that having an algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference.
-Therefore, it is important to select the best algorithm for each instance.
+Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging
+from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
+considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very
+quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
+preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2
+performs a basic backtracking search and is very fast on easy instances, because it does not compute
+expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
+and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow
+1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS),
+which considers the best algorithm for each instance separately, obtains much better results,
+showing us that the algorithms have complementary performance. The difference between VBS and single
+best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU
+time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the
+limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For
+example, in pattern recognition applications, we often have to solve subgraph isomorphism problems
+for a very large number of graphs (in order to find a pattern image in a large database of target
+images, for example), so that having an algorithm that is able to solve an instance in 100 ms
+instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm
+for each instance.
 
-Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class instances are comin from, unlike SAT.
-For all classes, there are always at least 2 algorithms which are the best for at least one instance of the class.
-In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for Class 2).
+Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class
+instances are comin from, unlike SAT. For all classes, there are always at least 2 algorithms which
+are the best for at least one instance of the class. In particular, for classes 2 and 3, each
+algorithm is the best for at least one instance (except Glasgow 4 for Class 2).
 
@@
 
-Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the upper bound of what an algorithm selection approach can achieve), and the single best solver, at different CPU time limits (the single best corresponds to results highlighted in blue in Table 1).
+Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the
+upper bound of what an algorithm selection approach can achieve), and the single best solver, at
+different CPU time limits (the single best corresponds to results highlighted in blue in Table 1).
 Comment Table 3: Mean runtime but also number of solved instances at different CPU time limits.
-Logically, LLAMA may solve less instances than single best for small time limits because it spends time to compute features and choose a solver, but it should be better for larger time limits...
+Logically, LLAMA may solve less instances than single best for small time limits because it spends
+time to compute features and choose a solver, but it should be better for larger time limits...
 

### 2015-12-06 paper/paper.tex
Rework and deduplicate a bit

@@
 
-\section{Background}
-
-\todo[inline]{Some background on subgraph isomorphism algorithms, maybe move
-some stuff from the intro here.}
+\section{Per-Instance Algorithm Selection}
 
@@
 salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
-
 The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
@@
 
-\section{Algorithms for subgraph isomorphism}
+\subsection{Subgraph isomorphism algorithms}
 
-We consider the following exact algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking
-search algorithm, especially fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which
-does very strong reasoning based on local all different redundant constraints; and the Glasgow
-algorithm \cite{McCreesh:2015}, which does an expensive preprocessing pass once, followed by weaker
-reasoning during search.
+We consider the following exact subgraph isomorphism
+algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially
+fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which does very strong reasoning
+based on local all different redundant constraints; and the Glasgow algorithm \cite{McCreesh:2015},
+which does an expensive preprocessing pass once, followed by weaker reasoning during search.
 
- The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning
- about non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice of paths
- of length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive
+The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about
+non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice of paths of
+length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive
 preprocessing on larger, denser instances. This is often not the best choice on an instance by
 instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only
- paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths
- of length 3 is relatively expensive on some graphs. Therefore, in this new study, we consider 4
+paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths of
+length 3 is relatively expensive on some graphs. Therefore, in this new study, we consider 4
 different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.
 
-We introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
+We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
 preprocessing) which is applied once, without performing a backtracking search, and detects
@@
 of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
-preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: When
+preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: when
 it does not detect inconsistency, the instance may either be consistent or not. The time complexity
@@
 
-\section{Benchmark}
+\subsection{Problem instances}
 
-We consider 5725 instances which are available in a simple text
+We consider 5725 instances, which are available in a simple text
 format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
@@
 \item Class 1 contains randomly generated scale-free graphs \cite{}.
-\item Classes 2 and 3 contain various kinds of graphs coming from \cite{}: class 2 contains small instances generated from the 50 first graphs of the benchmark, whereas class 3 contains larger instances where pattern graphs are coming from class 2, and target graphs are coming from the next 50 graphs of the benchmark.
-\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}: bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
+\item Classes 2 and 3 contain various kinds of graphs coming from \cite{}: class 2 contains small
+    instances generated from the 50 first graphs of the benchmark, whereas class 3 contains larger
+    instances where pattern graphs are coming from class 2, and target graphs are coming from the
+    next 50 graphs of the benchmark.
+\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}:
+    bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs
+    with uniform edge probabilities for class 7.
 \item Classes 9 and 10 contain instances coming from images \cite{}.
@@
 
-\noindent Randomly generated instances with particular properties:
+Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.
 
-\begin{itemize}
-    \item Bounded-degree random graphs (family 3, 4) \cite{GraphDatabase}
-    \item Random 4d mesh graphs (family 5, 6) \cite{GraphDatabase}
-    \item Random graphs (family 7) \cite{GraphDatabase}
-    \item Randomly generated scale-free graphs. (family 1) \cite{Solnon:2010}
-    \item Phase transition (family 12)
-\end{itemize}
+\subsection{Features}
 
-We consider a total of 5725 instances.
+For both the pattern and the target graph, we consider the number of vertices, edges and loops, the
+mean and maximum degree, the density, whether or not every vertex has same degree, whether or not
+the graph is connected, the mean and maximum distances between all pairs of vertices, and the
+proportion of vertex pairs which are distance at least 2, 3 and 4 apart. We also consider the number
+of variable-value assignments removed by LAD preprocessing in absolute terms, as a percentage, and
+the minimum and maximum on a per-variable basis.  This gives us a total of 30 features.
 
-\subsubsection{Features}
+\subsection{Experimental Setup}
 
-\todo[inline]{Describe features here.}
+We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
+Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
+the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
+\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
+timeout of $10^8$ milliseconds (a little over a day).
 
-This gives us a total of 30 features.
+\subsection{Algorithm Selection approach}
+
+We apply two static steps to each instance before solving it. First, the computation of the LAD
+features can detect unsatisfiability of an instance. If an instance is identified as unsatisfiable,
+we do not process it further.  Second, we run the VF2 solver for 0.05 seconds. This solves easy
+instances without the overhead of running algorithm selection and avoids potentially making
+incorrect solver choices.
+
+Computing the LAD features detects unsatisfiability on 1919 instances. Running VF2 for 0.05 seconds
+removes a further 1470. We are left with 2336 hard instances that we consider for algorithm
+selection.
+
+To decide which solver to run on the remaining instances, we use LLAMA~\cite{kotthoff_llama_2013}.
+In preliminary experiments, we identified random forest regression to predict the performance
+difference between pairs of algorithms, with the best algorithm identified as the one with the best
+cumulative performance difference over all pairs, as the best approach. We make no claims that this
+is the overall best approach.
 
@@
 
-\subsection{Algorithm Selection approach}
+\section{Experimental Results}
 
-We apply two static steps to each instance before solving it. First, the
-computation of the LAD features can detect unsatisfiability of an instance. If
-an instance is identified as unsatisfiable, we do not process it further.
-Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
-without the overhead of running algorithm selection and avoids potentially
-making incorrect solver choices.
-
-Computing the LAD features detects unsatisfiability on 1919 instances. Running
-VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
-instances that we consider for algorithm selection.
-
-To decide which solver to run on the remaining instances, we use
-LLAMA~\cite{kotthoff_llama_2013}. In preliminary experiments, we identified
-random forest regression to predict the performance difference between pairs of
-algorithms, with the best algorithm identified as the one with the best
-cumulative performance difference over all pairs, as the best approach. We make
-no claims that this is the overall best approach.
-
-\section{Results}
-
-Table~\ref{tab:res} shows the performance of our algorithm selection approach,
-compared to two baselines. The virtual best solver is the oracle predictor that,
-for each instance, chooses the best solver from our portfolio. This is the upper
-bound of what an algorithm selection approach can achieve. The single best
-solver is the one solver from the portfolio that has the overall best
-performance across the entire set of instances. We consider it a lower bound on
-performance.
+Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
+baselines. The virtual best solver is the oracle predictor that, for each instance, chooses the best
+solver from our portfolio. This is the upper bound of what an algorithm selection approach can
+achieve. The single best solver is the one solver from the portfolio that has the overall best
+performance across the entire set of instances. We consider it a lower bound on performance.
 
@@
 
-Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.
-
-\section{Experimental comparison}
-
-Our runtimes were measured on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
-Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
-the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
-\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
-timeout of $10^8$ milliseconds (a little over a day).
-
 \begin{table}[t]
@@
 Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class
-instances are comin from, unlike SAT. For all classes, there are always at least 2 algorithms which
+instances are coming from, unlike SAT. For all classes, there are always at least 2 algorithms which
 are the best for at least one instance of the class. In particular, for classes 2 and 3, each
@@
 
-\section{Algorithm selection approach}
-
-Describe features + selection approach.
-
 Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the
@@
 
+Mention parallel, directed, labelled in here briefly.
+
 \bibliographystyle{splncs}

### 2015-12-08 portfolio/as-graphs.Rnw
more graphs

@@
 library(aslib)
+library(tidyr)
 
@@
 <<echo=FALSE,fig.cap='Single best algorithm vs.\ portfolio performance (best model).'>>=
-best = which.min(c(resc$mean.performance))
-mod = list(modelc)[[best]]
+best = which.min(c(resc$mean.performance, resr$mean.performance, rescp$mean.performance, resrp$mean.performance))
+mod = list(modelc, modelr, modelcp, modelrp)[[best]]
 perfScatterPlot(parscores, mod, singleBest, ldf, addCostsx = TRUE, addCostsy = FALSE, factor = 1, pargs=aes(colour = family)) + scale_x_log10() + scale_y_log10() + xlab("portfolio") + ylab("single best")
+@
+
+<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio (best model).'>>=
+vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf), function(ss) { ss[1] })
+vbs.agg$virtual.best = parscores(ldf, vbs, factor = 1)
+vbs.agg$iteration = NULL
+vbs.agg$score = NULL
+
+pmod = mod$predictions
+pmod$portfolio = parscores(ldf, mod, factor = 1)
+pmod$iteration = NULL
+pmod$algorithm = NULL
+pmod$score = NULL
+
+perfs = subset(ldf$data, TRUE, c("instance_id", ldf$performance))
+perfs = merge(perfs, vbs.agg, by = "instance_id")
+perfs = merge(perfs, pmod, by = "instance_id")
+wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+ggplot(wide, aes(x = time, col = solver)) + stat_ecdf() + scale_x_log10() +
+    ylab("fraction of instances solved")
 @
@@
 perfScatterPlot(parscores, mod, singleBest, ldf3, factor = 1, pargs=aes(colour = family)) + scale_x_log10() + scale_y_log10() + xlab("portfolio") + ylab("single best")
+@
+
+<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio (best model).'>>=
+vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf3), function(ss) { ss[1] })
+vbs.agg$virtual.best = parscores(ldf3, vbs, factor = 1)
+vbs.agg$iteration = NULL
+vbs.agg$score = NULL
+
+pmod = mod$predictions
+pmod$portfolio = parscores(ldf3, mod, factor = 1)
+pmod$iteration = NULL
+pmod$algorithm = NULL
+pmod$score = NULL
+
+perfs = subset(ldf3$data, TRUE, c("instance_id", ldf3$performance))
+perfs = merge(perfs, vbs.agg, by = "instance_id")
+perfs = merge(perfs, pmod, by = "instance_id")
+wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+ggplot(wide, aes(x = time, col = solver)) + stat_ecdf() + scale_x_log10() +
+    ylab("fraction of instances solved")
 @
@@
 
+<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio (best model).'>>=
+vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf2), function(ss) { ss[1] })
+vbs.agg$virtual.best = parscores(ldf2, vbs, factor = 1)
+vbs.agg$iteration = NULL
+vbs.agg$score = NULL
+
+pmod = mod$predictions
+pmod$portfolio = parscores(ldf2, mod, factor = 1)
+pmod$iteration = NULL
+pmod$algorithm = NULL
+pmod$score = NULL
+
+perfs = subset(ldf2$data, TRUE, c("instance_id", ldf2$performance))
+perfs = merge(perfs, vbs.agg, by = "instance_id")
+perfs = merge(perfs, pmod, by = "instance_id")
+wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+ggplot(wide, aes(x = time, col = solver)) + stat_ecdf() + scale_x_log10() +
+    ylab("fraction of instances solved")
+@
+
 <<echo=FALSE, message=FALSE, warning=FALSE>>=

### 2015-12-08 portfolio/eda_config.R

@@ -1,8 +1,8 @@
 library(aslib)
 
 makeEDAConfig = function(
-  algo.perf.impute.zero.vals = FALSE,
-  algo.perf.log = FALSE
+  algo.perf.impute.zero.vals = TRUE,
+  algo.perf.log = TRUE
 ) {
 
   makeS3Obj("ASScenarioHTMLConfig",

### 2015-12-08 paper/paper.tex
updated figures

@@
 virtual best & 5822.809\\
-  single best & 7782.569\\
-  LLAMA & 6595.275\\
+  single best & 7788.719\\
+  LLAMA & 6509.947\\
    \bottomrule
@@
   single best & 3177.569\\
-  LLAMA & 2706.330\\
+  LLAMA & 2706.749\\
    \bottomrule
@@
 \label{fig:scatter}
+\end{figure}
+
+Figure~\ref{fig:portfolio-ecdf} shows the cumulative density function for the
+individual solvers, the virtual best solver, and the LLAMA portfolio. The
+virtual best solver clearly dominates with a significant margin to the next
+best. VF2 is far below all other solvers. The portfolio does not perform well
+for instances that can be solved quickly because of the overhead incurred
+through feature computation. As the instances become more difficult to solve,
+its performance improves. For instances that take more than 1,000 seconds to
+solve, the portfolio eclipses any individual solver and closes most of the gap
+to the virtual best solver.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
+\caption{Empirical cumulative density function of solvers, virtual best solvers,
+and the LLAMA portfolio.}
+\label{fig:portfolio-ecdf}
 \end{figure}

### 2015-12-08 paper/paper.tex
restructure and add ECDF graph

@@
 
+\usepackage{times}
 \usepackage{complexity}
@@
 
-\section{Per-Instance Algorithm Selection}
+\section{Background}
+
+\subsection{Per-Instance Algorithm Selection}
 
@@
 
-\subsection{Subgraph isomorphism algorithms}
+\subsection{Subgraph Isomorphism}
 
-We consider the following exact subgraph isomorphism
-algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially
-fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which does very strong reasoning
-based on local all different redundant constraints; and the Glasgow algorithm \cite{McCreesh:2015},
-which does an expensive preprocessing pass once, followed by weaker reasoning during search.
+\todo[inline]{Add some background on subgraph isomorphism.}
 
-The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about
-non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice of paths of
-length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive
-preprocessing on larger, denser instances. This is often not the best choice on an instance by
-instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only
-paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths of
-length 3 is relatively expensive on some graphs. Therefore, in this new study, we consider 4
-different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.
+\section{Experimental Setup}
 
-We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
-preprocessing) which is applied once, without performing a backtracking search, and detects
-inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
-one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$
-of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
-preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: when
-it does not detect inconsistency, the instance may either be consistent or not. The time complexity
-of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
-nodes, respectively, and $p$ is the number of target edges.
-
-The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in
-\cite{Solnon:2010}. It combines LAD filtering with new invariant properties borrowed from SND
-\cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
-between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
-and to add the constraint that the label of a pattern node (resp. edge) must be smaller than or
-equal to the label of its associated target node (resp. edge).
-
-\subsection{Problem instances}
+\subsection{Problem Instances}
 
@@
 
+\subsection{Subgraph Isomorphism Algorithms}
+
+We consider the following exact subgraph isomorphism
+algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially
+fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which does very strong reasoning
+based on local all different redundant constraints; and the Glasgow algorithm \cite{McCreesh:2015},
+which does an expensive preprocessing pass once, followed by weaker reasoning during search.
+
+We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
+preprocessing) which is applied once, without performing a backtracking search, and detects
+inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
+one target node $v$ such that for each neighbour $u'$ of $u$ there exists a
+different neighbour $v'$
+of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
+preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: when
+it does not detect inconsistency, the instance may either be consistent or not. The time complexity
+of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
+nodes, respectively, and $p$ is the number of target edges.
+
+The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in
+\cite{Solnon:2010}. It combines LAD filtering with new invariant properties borrowed from SND
+\cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
+between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
+and to add the constraint that the label of a pattern node (resp.\ edge) must be smaller than or
+equal to the label of its associated target node (resp.\ edge).
+
+The Glasgow algorithm has a parameter, which controls the lengths of paths used
+when reasoning about non-adjacent vertices.  In experiments reported in
+\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable
+compromise -- longer paths lead to prohibitively expensive preprocessing on
+larger, denser instances. This is often not the best choice on an instance by
+instance basis: sometimes path-based reasoning gives no benefit at all,
+sometimes considering only paths of length 2 suffices, occasionally paths of
+length 4 are helpful, and even looking at paths of length 3 is relatively
+expensive on some graphs.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figures/ecdf-glasgow}
+\caption{Empirical cumulative distribution function for the Glasgow algorithm
+with path lengths 1, 2, 3, and 4.}
+\label{fig:ecdf-glasgow}
+\end{figure}
+
+Figure~\ref{fig:ecdf-glasgow} shows the cumulative distribution function for the
+Glasgow algorithm with four different path lengths. As the instances become
+harder to solve, longer paths become better. This is what we expect, as more
+reasoning is expensive, but potentially increases the reduction of the search
+space. In particular, the relative performance of the algorithms is monotone in
+the time it takes to solve the instance -- path length 1 is best at first, but
+gets worse as the time to solve the instance increases until path length 2 is
+better. After that point, length 1 is never better than length 2 again. The same
+is true for the other path lengths.
+
+The figure illustrates the potential for portfolios and algorithm selection we
+have. There is clearly no path length that dominates throughout. Therefore, we
+consider 4 different instances of the Glasgow algorithm here, with path lengths
+varying from 1 to 4.
+
 \subsection{Features}
@@
 
-\subsection{Experimental Setup}
-
-We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
-Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
-the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
-\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
-timeout of $10^8$ milliseconds (a little over a day).
-
-\subsection{Algorithm Selection approach}
+\section{Algorithm Selection Approach}
 
@@
 The data we use in this paper is available as Aslib scenario GRAPHS-2015.
+
+\subsection{Experimental Environment}
+
+We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
+Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
+the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
+\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
+timeout of $10^8$ milliseconds (a little over a day).
 
@@
 
-Figure~\ref{fig:portfolio-ecdf} shows the cumulative density function for the
-individual solvers, the virtual best solver, and the LLAMA portfolio. The
+Figure~\ref{fig:portfolio-ecdf} shows the cumulative distribution function for
+the individual solvers, the virtual best solver, and the LLAMA portfolio. The
 virtual best solver clearly dominates with a significant margin to the next
@@
 \includegraphics[width=\textwidth]{figures/portfolio-ecdf}
-\caption{Empirical cumulative density function of solvers, virtual best solvers,
-and the LLAMA portfolio.}
+\caption{Empirical cumulative distribution function of solvers, virtual best
+solvers, and the LLAMA portfolio.}
 \label{fig:portfolio-ecdf}

### 2015-12-08 paper/paper.tex
extend algorithm selection part

@@
 
-For both the pattern and the target graph, we consider the number of vertices, edges and loops, the
-mean and maximum degree, the density, whether or not every vertex has same degree, whether or not
-the graph is connected, the mean and maximum distances between all pairs of vertices, and the
-proportion of vertex pairs which are distance at least 2, 3 and 4 apart. We also consider the number
-of variable-value assignments removed by LAD preprocessing in absolute terms, as a percentage, and
-the minimum and maximum on a per-variable basis.  This gives us a total of 30 features.
+For both the pattern and the target graph, we consider the number of vertices,
+edges and loops, the mean and maximum degree, the density, whether or not every
+vertex has same degree, whether or not the graph is connected, the mean and
+maximum distances between all pairs of vertices, and the proportion of vertex
+pairs which are distance at least 2, 3 and 4 apart. We also consider the number
+of variable-value assignments removed by LAD preprocessing in absolute terms, as
+a percentage, and the minimum and maximum on a per-variable basis. This gives
+us a total of 30 features.
 
-\section{Algorithm Selection Approach}
-
-We apply two static steps to each instance before solving it. First, the computation of the LAD
-features can detect unsatisfiability of an instance. If an instance is identified as unsatisfiable,
-we do not process it further.  Second, we run the VF2 solver for 0.05 seconds. This solves easy
-instances without the overhead of running algorithm selection and avoids potentially making
-incorrect solver choices.
-
-Computing the LAD features detects unsatisfiability on 1919 instances. Running VF2 for 0.05 seconds
-removes a further 1470. We are left with 2336 hard instances that we consider for algorithm
-selection.
-
-To decide which solver to run on the remaining instances, we use LLAMA~\cite{kotthoff_llama_2013}.
-In preliminary experiments, we identified random forest regression to predict the performance
-difference between pairs of algorithms, with the best algorithm identified as the one with the best
-cumulative performance difference over all pairs, as the best approach. We make no claims that this
-is the overall best approach.
-
-The data we use in this paper is available as Aslib scenario GRAPHS-2015.
+In addition to computing feature values, the LAD preprocessing is able to detect
+whether an instance is unsatisfiable. In this case, it is not necessary to run
+an algorithm on the instance.
 
@@
 timeout of $10^8$ milliseconds (a little over a day).
+
+\section{Algorithm Selection Approach}
+
+We apply two static steps to each instance before solving it. First, the
+computation of the LAD features can detect unsatisfiability of an instance. If
+an instance is identified as unsatisfiable, we do not process it further.
+Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
+without the overhead of running algorithm selection and avoids potentially
+making incorrect solver choices.
+
+Computing the LAD features detects unsatisfiability on 1919 instances. Running
+VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
+instances that we consider for algorithm selection.
+
+To decide which solver to run on the remaining instances, we use
+LLAMA~\cite{kotthoff_llama_2013}. LLAMA supports the most common algorithm
+selection approaches used in the literature. We performed a set of preliminary
+experiments to determine the approach that works best in this case.
+
+LLAMA's pairwise regression approach with random forest regression gave the best
+performance. The idea is very similar to the pairwise classification used
+in~\cite{xu_satzilla_2008}. For each pair of algorithms in our portfolio, we
+train a model that predicts the performance difference between them. That is, if
+the first algorithm is better than the second, the difference is positive,
+otherwise negative. The algorithm with the highest cumulative performance
+difference, i.e.\ the most positive difference over all other algorithms, is
+chosen to be run.
+
+As this approach gives very good performance already, we did not tune the
+parameters of the random forest machine learning algorithm. It is possible that
+overall performance can be improved by doing so and we make no claims that the
+particular algorithm selection approach we use in this paper cannot be improved.
+
+The data we use in this paper is available as ASlib~\cite{aslib} scenario
+GRAPHS-2015.
 

### 2015-12-08 paper/paper.tex
paragraph about the other subgraph isomorphism portfolio paper

@@
 
+\subsection{Subgraph Isomorphism}
+
+\todo[inline]{Add some background on subgraph isomorphism.}
+
 \subsection{Per-Instance Algorithm Selection}
@@
 
-\subsection{Subgraph Isomorphism}
+\medskip
 
-\todo[inline]{Add some background on subgraph isomorphism.}
+Battiti and Mascia~\cite{battiti_algorithm_2007} were the first to propose
+algorithm portfolios for subgraph isomorphism problems. However, they only
+consider a pure parallel portfolio without a selection mechanism consisting of
+only two solvers. In addition, our instance set is much larger and considers
+both hard and unsatisfiable instances.
 

### 2015-12-09 paper/paper.tex
Merge Christine's changes

@@
 In this paper, we report experimental results on a larger set of 8 algorithms, including new
-variants, and a larger benchmark. We show that 5 of these algorithms are single best algorithms,
+variants, and a larger benchmark. We show that 4 of these algorithms are single best algorithms,
 depending on the CPU time limit considered, and that combining preprocessing with an algorithm
@@
 
-\section{Background}
+\section{Definitions and notations}
 
-\subsection{Subgraph Isomorphism}
+A \emph{graph} $G=(N,E)$ consists of a \emph{node set} $N$ and an \emph{edge set} $E \subseteq N \times N$, where an edge $(u,u')$ is a couple of nodes.
+The number of neighbors of a node $u$ is the degree of $u$, $d^\circ(u)=\#\{ (u,u')\in E\}$.
+In this paper, we implicitely consider non directed graphs, such that $(u,u')\in E\Leftrightarrow (u',u)\in E$.
+The extension to directed graphs is rather straightforward, and all algorithms compared in this paper can handle directed graphs as well.
 
-\todo[inline]{Add some background on subgraph isomorphism.}
+A \emph{subgraph isomorphism problem} between a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$ consists in deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
+More precisely, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a different target node to each pattern node, and that preserves pattern edges, i.e., $\forall (u,u') \in E_p, (f(u),f(u')) \in E_t$.
 
-\subsection{Per-Instance Algorithm Selection}
+Note that the subgraph is not necessarily induced so that two pattern nodes that are not linked by an edge may be matched to two target nodes which are linked by an edge.
+This problem is also called subgraph monomorphism or subgraph matching in the literature.
+%{\em Induced subgraph isomorphism} is a particular case of subgraph isomorphism such that the subgraph isomorphism function must also preserve target edges between matched nodes, i.e., 
+%\[\forall (u,u') \in N_p^2, (u,u')\in E_p \Leftrightarrow(f(u),f(u')) \in E_t\]
+
+In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of subgraph isomorphism problem. 
+We also define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\# E_t$, and $d_p$ and $d_t$ the maximal degrees of the graphs $G_p$ and $G_t$.
+
+\section{Subgraph isomorphism algorithms}
+
+Subgraph isomorphism problems may be solved by a systematic exploration of the search space composed of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one incrementally extends a partial matching by matching a non matched pattern node to a non matched target node until either some edges are not matched by the current matching (the search must backtrack to a previous choice point and go on with another extension) or all pattern nodes have been matched (a solution has been found).
+To reduce the search space, this exhaustive exploration is combined with filtering techniques that aim at removing candidate couples of non matched pattern-target nodes $(u,v)\in N_p\times N_t$.
+Different filterings may be considered; some are stronger than others (they remove more candidate couples), but also have higher time complexities.
+
+\paragraph{Filtering for subgraph isomorphism.}
+
+Simple filterings basically propagate difference constraints (which ensure that the matching is injective) and edge constraints (which ensure that the matching preserves pattern edges): each time a pattern node $u\in N_p$ is matched with a target node $v\in N_t$, one removes every candidate couple $(u',v')\in N_p\times N_t$ such that either $v'=v$ (difference constraint) or $(u,u')$ is a pattern edge whereas $(v,v')$ is not a target edge (edge constraint).
+This simple filtering (called {\em Forward-Checking}) is very fast to achieve: in ${\cal O}(n_p)$ for difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints.
+It is used, for example, in McGregor's algorithm \cite{mcgregor79} and in VF2 \cite{Cordella:2004}.
+
+R\'egin has introduced in \cite{regin} a stronger filtering for difference constraints, which ensures that all pattern nodes can be matched with different target nodes, all together.
+This filtering (called {\em AllDifferent Generalized Arc Consistency}) removes more candidate couples than when each difference contraint is propagated separately which Forward-Checking.
+However, it is also more time consuming as it is done in ${\cal O}(n_p^2\cdot n_t^2)$.
+
+Various filterings have been introduced for edge constraints.
+Ullman \cite{ullman} has introduced a filtering which basically ensures that for each pattern edge $(u,u')\in E_p$ and each candidate couple $(u,v)\in N_p\times N_t$, there exists a candidate couple $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a target edge.
+Candidate couples $(u,v)$ that do not satisfy this property are iteratively removed until a fixed point is reached.
+This filtering (called {\em Arc Consistency}) removes more candidate couples than Forward-Checking, but it is also more time consuming as it is done in ${\cal O}(e_p\cdot n_t^2)$ when using the algorithm AC4  \cite{MH86}.
+
+A stronger filtering may be obtained by propagating edge constraints in a more global way, as proposed in \cite{LV02}.
+The idea is to check for each candidate couple $(u,v)\in N_p\times N_t$ that the number of pattern nodes adjacent to $u$ is smaller than or equal to the number of target nodes that are both adjacent to $v$ and that may be matched with nodes adjacent to $u$.
+This is done in ${\cal O}(n_p^2\cdot n_t^2)$.
+This idea has been generalized in \cite{Solnon:2010} where, for each candidate couple $(u,v)\in N_p\times N_t$, a redundant Local AllDifferent (LAD) constraint ensures that each neighbour of $u$ may be matched with a different neighbour of $v$.
+This is done in ${\cal O}(n_p\cdot n_t\cdot d_p^2\cdot d_t^2)$.
+
+\paragraph{Propagation of invariant properties.}
+
+Some filterings exploit invariant properties, {\em i.e.}, properties associated with nodes such that nodes may be matched only if they have compatible properties.
+A classical property is the degree: a pattern node $u\in N_p$ may be matched with a target node $v\in N_t$ only if $d^\circ(u)\leq d^\circ(v)$.
+This property is usually used at the beginning of the search to reduce the set of candidate couples to $\{(u,v)\in N_p\times N_t\;|\;d^\circ(u)\leq d^\circ(v)\}$.
+Other examples of invariant properties are the number of cycles of length $k$ passing through the node, and the number of cliques of size $k$ containing the node, which must be smaller for a pattern node than for its matched target node.
+Invariant properties may also be associated with couples of nodes.
+For example, the number of paths of length $k$ between two pattern nodes is smaller than or equal to the number of paths of length $k$ between the target nodes they may be matched with.
+
+These invariant properties are used, for example:
+\begin{itemize}
+\item  In \cite{battiti-mascia07}, to remove candidate couples $(u,v)\in N_p\times N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of paths starting from  target node $v$~;
+\item In \cite{Audemard:2014} to generalize the LAD constraint proposed in \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be matched with all different compatible target nodes, where compatibility is defined with respect to invariant properties~;
+\item In \cite{McCreesh:2015} to filter the set of candidate couples before starting the search.
+\end{itemize}
+In \cite{Audemard:2014} , the length of paths is not limited, and it is iteratively incremented until no more couple is removed.
+In \cite{battiti-mascia07} and \cite{McCreesh:2015},  algorithms are parameterized by the maximum path length considered when counting paths: Larger values for this parameter remove more candidate couples, but are also more time consuming.
+Experiments reported in \cite{battiti-mascia07} show us that the best setting depends on the instance considered, and that a portfolio running several randomized
+versions in time-sharing decreases the total CPU-time needed to find a solution for feasible instances.
+In experiments reported in \cite{McCreesh:2015}, this parameter has been set to 3, as this setting seemed to be a reasonable compromise.
+
+\paragraph{Algorithms selected in our portfolio.}
+We have selected a set of algorithms with complementary performance:
+\begin{itemize}
+\item VF2 \cite{Cordella:2004}, which performs a weak filtering so that it is especially fast on trivially satisfiable instances;
+\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques (AllDifferent Generalized Arc Consistency and LAD);
+\item Glasgow \cite{McCreesh:2015}, which does an expensive preprocessing based on path length invariant properties, followed by weaker filtering (forward-checking) during search.
+We consider 4 different instanciations of Glasgow, with path lengths varying from 1 to 4: The longer the path, the stronger the reduction, and the more expensive to achieve.
+\end{itemize}
+
+We also introduce two new variants of LAD.
+The first variant is a weaker LAD filtering (called incompleteLAD) which is applied once, without performing a backtracking search, and very quickly detects inconsistencies on many instances: For each pattern node $u$, we check that there exists at least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
+IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when it does not detect inconsistency, the instance may either be consistent or not.
+As a counterpart of this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.
+
+The second variant of LAD is called pathLAD and it combines the LAD constraints introduced in \cite{Solnon:2010} with the exploitation of path length properties, as proposed in \cite{Audemard:2014}.
+The idea is to label each edge $(u,v)$ with the number of paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$, and to add the constraint that the label of a pattern node (resp.
+edge) must be smaller than or equal to the label of its associated target node (resp.
+edge).
+
+\section{Experimental Comparison of Algorithms of the Portfolio}\label{expComp}
+
+\paragraph{Benchmark.} We consider 5725 instances, which are available in a simple text format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12 classes.
+
+\begin{itemize}
+\item Class 1 contains randomly generated scale-free graphs \cite{constraints10}.
+\item Classes 2 and 3 contain various kinds of graphs coming from \cite{LV02}: class 2 contains small
+    instances generated from the 50 first graphs, whereas class 3 contains larger
+    instances where pattern graphs are coming from class 2, and target graphs are coming from the
+    next 50 graphs.
+\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}:
+    bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs
+    with uniform edge probabilities for class 7.
+\item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
+\item Class 11 contains instances coming from meshes modeling 3D objects \cite{cviu11}.
+\item Class 12 contains random graph instances belonging to the phase transition.
+\end{itemize}
+
+Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.
+
+\paragraph{Experimental setup.} We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
+Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
+the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
+\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
+timeout of $10^8$ milliseconds (a little over a day).
+
+\begin{figure}[t]
+\includegraphics[width=\textwidth]{courbe.pdf}
+\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms of the portfolio, and a Virtual Best Solver (VBS).\label{expTime}}
+\end{figure}
+
+\paragraph{Results.}Figure \ref{expTime} displays the evolution of the number of instances solved with respect to CPU time. It shows us that the best solver depends on the time limit
+considered.
+IncompleteLAD is able to solve easy inconsistent instances very quickly, in a few milliseconds.
+Hence, for time limits lower than 5 ms, the best solver is incompleteLAD.
+However, it is not able to solve harder inconsistent instances, neither can it solve consistent instances.
+
+PathLAD and Glasgow(1) outperform incompleteLAD when considering longer time limits: PathLAD is the best solver for time limits larger than 5ms and lower than 40 ms, and Glasgow(1) is the best solver for time limits larger than 40ms and lower than 3000 ms.
+
+Finally, for CPU time limits larger than 3000 ms, the best solver becomes Glasgow(2).
+As we increase the CPU time limit, variants of Glasgow with longer paths (Glasgow(3) and Glasgow(4)) become better.
+This is what we expect, as more reasoning is expensive, but potentially increases the reduction of the search space.
+In particular, the relative performance of Glasgow variants is monotone in the time it takes to solve the instance – Glasgow(1) is best at first, but gets worse as the time to solve the instance increases until Glasgow(2) is better.
+After that point, Glasgow(1) is never better than Glasgow(2) again.
+The same is true for path lengths 3 and 4.
+
+The figure illustrates the potential for portfolios and algorithm selection we have. There is clearly no single solver that dominates throughout. 
+Furthermore, the Virtual Best Solver (VBS), which considers the best algorithm for each instance separately, obtains much better results, showing us that the algorithms have complementary performance.
+The difference between VBS and single best is more particularly important for rather small CPU time limits (smaller than 1000 ms).
+In many applications, it is important to have the fastest possible algorithm.
+For example, in pattern recognition applications such as those reported in \cite{pr15,cviu11}, we often have to solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern image in a large database of target images, for example), so that having an algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference.
+Therefore, it is important to select the best algorithm for each instance, even if the instance is an easy one.
+
+Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class instances are coming from, unlike SAT.
+For all classes, there are always at least 2 algorithms which are the best for at least one instance of the class.
+In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for Class 3).
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{|c||r||r|r|r||r|r|r|r|}
+\hline
+Class & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c|}{Glasgow}\\
+&&incomplete&default&path&1&2&3&4\\\hline\hline
+1 &        0 &       20 &        0 &        0 &       80 &        0 &        0 &        0 \\\hline
+2 &      201 &       92 &      189 &      270 &      520 &      180 &       53 &       15 \\\hline
+3 &      112 &     1608 &      617 &      959 &      396 &      195 &       21 &        0 \\\hline
+4 &      270 &        0 &        0 &        0 &        5 &        0 &        0 &        0 \\\hline
+5 &      266 &        0 &        1 &        3 &       31 &        0 &        0 &        0 \\\hline
+6 &       71 &        0 &        0 &        0 &        7 &       14 &        1 &        0 \\\hline
+7 &      270 &        0 &        0 &        0 &        5 &        0 &        0 &        0 \\\hline
+8 &        0 &        0 &        0 &        1 &      195 &       69 &        6 &        0 \\\hline
+9 &       77 &        3 &        0 &       19 &      103 &        1 &        0 &        0 \\\hline
+10 &       13 &        0 &        2 &        2 &        7 &        0 &        0 &        0 \\\hline
+11 &        2 &      142 &       71 &       17 &       23 &        0 &        0 &        0 \\\hline
+12 &        0 &        0 &        1 &        2 &      158 &        6 &        1 &        0 \\\hline\hline
+Total & 1282 & 1865 & 881 & 1273 & 1530 & 465 & 82& 15\\\hline
+\end{tabular}
+\end{center}
+\caption{Number of times each algorithm is best, for each class.\label{expClass}}
+\end{table}
+
+\section{Per-Instance Algorithm Selection for subgraph isomorphism}
 
@@
 
-Battiti and Mascia~\cite{battiti_algorithm_2007} were the first to propose
+Battiti and Mascia~\cite{battiti-mascia07} were the first to propose
 algorithm portfolios for subgraph isomorphism problems. However, they only
@@
 
-\section{Experimental Setup}
+Our approach is composed of 3 steps described below: First, we run two presolvers in a static way to quickly solve easy instances; Second, we extract features from instances which are not solved by the first step; Finally, we select an algorithm and run it.
 
-\subsection{Problem Instances}
+\paragraph{Static steps.}
 
-We consider 5725 instances, which are available in a simple text
-format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
-classes.
+Experimental results reported in Section \ref{expComp} have shown us that incompleteLAD is very fast (7 ms on average) and is able to very quickly solve 1919 instances.
+Therefore, we propose to first run incompleteLAD: If the instance is solved, then we do not process it further.
+Furthermore, running IncompleteLAD allows us to collect features on the number of candidate couples which are removed by the filtering.
 
-\begin{itemize}
-\item Class 1 contains randomly generated scale-free graphs \cite{}.
-\item Classes 2 and 3 contain various kinds of graphs coming from \cite{}: class 2 contains small
-    instances generated from the 50 first graphs of the benchmark, whereas class 3 contains larger
-    instances where pattern graphs are coming from class 2, and target graphs are coming from the
-    next 50 graphs of the benchmark.
-\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}:
-    bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs
-    with uniform edge probabilities for class 7.
-\item Classes 9 and 10 contain instances coming from images \cite{}.
-\item Class 11 contains instances coming from meshes modeling 3D objects \cite{}.
-\item Class 12 contains random graph instances belonging to the phase transition.
-\end{itemize}
+VF2 is also able to solve very quickly many easy instances: Among the 3806 instances which are not solved by incompleteLAD, 1470 are solved by VF2 in less than 50 milliseconds.
+Therefore, after running incompleteLAD, we run the VF2 solver for 50 ms.
+This solves easy instances without the overhead of running algorithm selection and avoids potentially making incorrect solver choices.
+After running incompleteLAD and VF2 for 50 ms, we are left with 2336 hard instances that we consider for algorithm selection.
 
-Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.
-
-\subsection{Subgraph Isomorphism Algorithms}
-
-We consider the following exact subgraph isomorphism
-algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially
-fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which does very strong reasoning
-based on local all different redundant constraints; and the Glasgow algorithm \cite{McCreesh:2015},
-which does an expensive preprocessing pass once, followed by weaker reasoning during search.
-
-We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
-preprocessing) which is applied once, without performing a backtracking search, and detects
-inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
-one target node $v$ such that for each neighbour $u'$ of $u$ there exists a
-different neighbour $v'$
-of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
-preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: when
-it does not detect inconsistency, the instance may either be consistent or not. The time complexity
-of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
-nodes, respectively, and $p$ is the number of target edges.
-
-The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in
-\cite{Solnon:2010}. It combines LAD filtering with new invariant properties borrowed from SND
-\cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
-between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
-and to add the constraint that the label of a pattern node (resp.\ edge) must be smaller than or
-equal to the label of its associated target node (resp.\ edge).
-
-The Glasgow algorithm has a parameter, which controls the lengths of paths used
-when reasoning about non-adjacent vertices.  In experiments reported in
-\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable
-compromise -- longer paths lead to prohibitively expensive preprocessing on
-larger, denser instances. This is often not the best choice on an instance by
-instance basis: sometimes path-based reasoning gives no benefit at all,
-sometimes considering only paths of length 2 suffices, occasionally paths of
-length 4 are helpful, and even looking at paths of length 3 is relatively
-expensive on some graphs.
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figures/ecdf-glasgow}
-\caption{Empirical cumulative distribution function for the Glasgow algorithm
-with path lengths 1, 2, 3, and 4.}
-\label{fig:ecdf-glasgow}
-\end{figure}
-
-Figure~\ref{fig:ecdf-glasgow} shows the cumulative distribution function for the
-Glasgow algorithm with four different path lengths. As the instances become
-harder to solve, longer paths become better. This is what we expect, as more
-reasoning is expensive, but potentially increases the reduction of the search
-space. In particular, the relative performance of the algorithms is monotone in
-the time it takes to solve the instance -- path length 1 is best at first, but
-gets worse as the time to solve the instance increases until path length 2 is
-better. After that point, length 1 is never better than length 2 again. The same
-is true for the other path lengths.
-
-The figure illustrates the potential for portfolios and algorithm selection we
-have. There is clearly no path length that dominates throughout. Therefore, we
-consider 4 different instances of the Glasgow algorithm here, with path lengths
-varying from 1 to 4.
-
-\subsection{Features}
+\paragraph{Features.}
 
@@
 pairs which are distance at least 2, 3 and 4 apart. We also consider the number
-of variable-value assignments removed by LAD preprocessing in absolute terms, as
+of candidate couples removed by incompleteLAD in absolute terms, as
 a percentage, and the minimum and maximum on a per-variable basis. This gives
@@
 
-In addition to computing feature values, the LAD preprocessing is able to detect
-whether an instance is unsatisfiable. In this case, it is not necessary to run
-an algorithm on the instance.
-
-\subsection{Experimental Environment}
-
-We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
-Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
-the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
-\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
-timeout of $10^8$ milliseconds (a little over a day).
-
-\section{Algorithm Selection Approach}
-
-We apply two static steps to each instance before solving it. First, the
-computation of the LAD features can detect unsatisfiability of an instance. If
-an instance is identified as unsatisfiable, we do not process it further.
-Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
-without the overhead of running algorithm selection and avoids potentially
-making incorrect solver choices.
-
-Computing the LAD features detects unsatisfiability on 1919 instances. Running
-VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
-instances that we consider for algorithm selection.
+\paragraph{Algorithm selection.}
 
@@
 
-\section{Experimental Results}
+\section{Experimental Evaluation of Algorithm Selection}
 
@@
 achieve. The single best solver is the one solver from the portfolio that has the overall best
-performance across the entire set of instances. We consider it a lower bound on performance.
+performance across the entire set of instances, at the CPU time limit of $10^8$ ms, {\em i.e.}, Glasgow(2). We consider it a lower bound on performance.
 
@@
 
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
-\caption{Empirical cumulative distribution function of solvers, virtual best
-solvers, and the LLAMA portfolio.}
-\label{fig:portfolio-ecdf}
-\end{figure}
-
-\begin{table}[t]
-\begin{center}
-\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
-\hline
-$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
-&&pre&def&new&1&2&3&4&&\\\hline
-$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
-$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
-$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
-$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
-$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
-$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
-$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
-$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
-$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
-\end{tabular}
-\end{center}
-\caption{Number of solved instances at different CPU time limits: Each line displays a time limit
-$t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD
-preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and
-4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
-single best if displayed in the last column ($\delta$).\label{expTime}}
-\end{table}
-
-Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging
-from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
-considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very
-quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
-preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2
-performs a basic backtracking search and is very fast on easy instances, because it does not compute
-expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
-and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow
-1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS),
-which considers the best algorithm for each instance separately, obtains much better results,
-showing us that the algorithms have complementary performance. The difference between VBS and single
-best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU
-time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the
-limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For
-example, in pattern recognition applications, we often have to solve subgraph isomorphism problems
-for a very large number of graphs (in order to find a pattern image in a large database of target
-images, for example), so that having an algorithm that is able to solve an instance in 100 ms
-instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm
-for each instance.
-
-Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class
-instances are coming from, unlike SAT. For all classes, there are always at least 2 algorithms which
-are the best for at least one instance of the class. In particular, for classes 2 and 3, each
-algorithm is the best for at least one instance (except Glasgow 4 for Class 2).
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|c||r||r|r|r||r|r|r|r|}
-\hline
-Class & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c|}{Glasgow}\\
-&&pre&def&new&1&2&3&4\\\hline
-1\\\hline
-2\\\hline
-3\\\hline
-4\\\hline
-5\\\hline
-6\\\hline
-7\\\hline
-8\\\hline
-9\\\hline
-10\\\hline
-11\\\hline
-12\\\hline
-
-\end{tabular}
-\end{center}
-\caption{Number of times each algorithm is best, for each class.\label{expClass}}
-\end{table}
-
-Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the
-upper bound of what an algorithm selection approach can achieve), and the single best solver, at
-different CPU time limits (the single best corresponds to results highlighted in blue in Table 1).
-Comment Table 3: Mean runtime but also number of solved instances at different CPU time limits.
-Logically, LLAMA may solve less instances than single best for small time limits because it spends
-time to compute features and choose a solver, but it should be better for larger time limits...
-
-\begin{table}
-\begin{tabular}{|l|r|rrrrrrrrr|}
-&mean & \multicolumn{9}{c|}{Number of solved instances}\\
-&runtime & $10^0$ &  $10^1$ &  $10^2$ &  $10^3$ &  $10^4$ &  $10^5$ &  $10^6$ &  $10^7$ &  $10^8$\\\hline
-VBS & 2375.9 & 1132 & 3388 & 4636 & 5170 & 5332 & 5434 & 5500 & 5567 & 5608\\\hline
-Single best & 3177.6 & 1126 & 2287 & 3943 & 4908 & 5253 & 5377 & 5452 & 5518 & 5561\\\hline
-LLAMA & 2706.3\\\hline
-\end{tabular}
-\caption{}
-\end{table}
-
-Add Figure 1 which compares single best at $10^8$ ms (Glasgow 2) with LLAMA ?
-
 \section{Conclusion and Future Work}

### 2015-12-09 paper/paper.tex

@@
 
+The Glasgow algorithm has a parameter, which controls the lengths of paths used
+when reasoning about non-adjacent vertices.  In experiments reported in
+\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable
+compromise -- longer paths lead to prohibitively expensive preprocessing on
+larger, denser instances. This is often not the best choice on an instance by
+instance basis: sometimes path-based reasoning gives no benefit at all,
+sometimes considering only paths of length 2 suffices, occasionally paths of
+length 4 are helpful, and even looking at paths of length 3 is relatively
+expensive on some graphs.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figures/ecdf-glasgow}
+\caption{Empirical cumulative distribution function for the Glasgow algorithm
+with path lengths 1, 2, 3, and 4.}
+\label{fig:ecdf-glasgow}
+\end{figure}
+
+Figure~\ref{fig:ecdf-glasgow} shows the cumulative distribution function for the
+Glasgow algorithm with four different path lengths. As the instances become
+harder to solve, longer paths become better. This is what we expect, as more
+reasoning is expensive, but potentially increases the reduction of the search
+space. In particular, the relative performance of the algorithms is monotone in
+the time it takes to solve the instance -- path length 1 is best at first, but
+gets worse as the time to solve the instance increases until path length 2 is
+better. After that point, length 1 is never better than length 2 again. The same
+is true for the other path lengths.
+
+The figure illustrates the potential for portfolios and algorithm selection we
+have. There is clearly no path length that dominates throughout. Therefore, we
+consider 4 different instances of the Glasgow algorithm here, with path lengths
+varying from 1 to 4.
+
 We also introduce two new variants of LAD.
@@
 
-\paragraph{Benchmark.} We consider 5725 instances, which are available in a simple text format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12 classes.
+\paragraph{Benchmark.}
+
+We consider 5725 instances, which are available in a simple text
+format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These
+instances are grouped into 12 classes.
 
@@
 
-\section{Per-Instance Algorithm Selection for subgraph isomorphism}
+\subsection{Per-Instance Algorithm Selection}
 
@@
 
+\section{Algorithm Selection Approach}
+
 Our approach is composed of 3 steps described below: First, we run two presolvers in a static way to quickly solve easy instances; Second, we extract features from instances which are not solved by the first step; Finally, we select an algorithm and run it.
@@
 
-\paragraph{Features.}
+\subsection{Features}
 
@@
 
-\paragraph{Algorithm selection.}
+In addition to computing feature values, the LAD preprocessing is able to detect
+whether an instance is unsatisfiable. In this case, it is not necessary to run
+an algorithm on the instance.
+
+We apply two static steps to each instance before solving it. First, the
+computation of the LAD features can detect unsatisfiability of an instance. If
+an instance is identified as unsatisfiable, we do not process it further.
+Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
+without the overhead of running algorithm selection and avoids potentially
+making incorrect solver choices.
+
+Computing the LAD features detects unsatisfiability on 1919 instances. Running
+VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
+instances that we consider for algorithm selection.
 
@@
 
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
+\caption{Empirical cumulative distribution function of solvers, virtual best
+solvers, and the LLAMA portfolio.}
+\label{fig:portfolio-ecdf}
+\end{figure}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
+\hline
+$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
+&&pre&def&new&1&2&3&4&&\\\hline
+$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
+$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
+$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
+$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
+$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
+$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
+$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
+$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
+$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
+\end{tabular}
+\end{center}
+\caption{Number of solved instances at different CPU time limits: Each line displays a time limit
+$t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD
+preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and
+4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
+single best if displayed in the last column ($\delta$).\label{expTime}}
+\end{table}
+
+Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging
+from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
+considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very
+quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
+preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2
+performs a basic backtracking search and is very fast on easy instances, because it does not compute
+expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
+and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow
+1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS),
+which considers the best algorithm for each instance separately, obtains much better results,
+showing us that the algorithms have complementary performance. The difference between VBS and single
+best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU
+time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the
+limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For
+example, in pattern recognition applications, we often have to solve subgraph isomorphism problems
+for a very large number of graphs (in order to find a pattern image in a large database of target
+images, for example), so that having an algorithm that is able to solve an instance in 100 ms
+instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm
+for each instance.
+
+Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the
+upper bound of what an algorithm selection approach can achieve), and the single best solver, at
+different CPU time limits (the single best corresponds to results highlighted in blue in Table 1).
+Comment Table 3: Mean runtime but also number of solved instances at different CPU time limits.
+Logically, LLAMA may solve less instances than single best for small time limits because it spends
+time to compute features and choose a solver, but it should be better for larger time limits...
+
+\begin{table}
+\begin{tabular}{|l|r|rrrrrrrrr|}
+&mean & \multicolumn{9}{c|}{Number of solved instances}\\
+&runtime & $10^0$ &  $10^1$ &  $10^2$ &  $10^3$ &  $10^4$ &  $10^5$ &  $10^6$ &  $10^7$ &  $10^8$\\\hline
+VBS & 2375.9 & 1132 & 3388 & 4636 & 5170 & 5332 & 5434 & 5500 & 5567 & 5608\\\hline
+Single best & 3177.6 & 1126 & 2287 & 3943 & 4908 & 5253 & 5377 & 5452 & 5518 & 5561\\\hline
+LLAMA & 2706.3\\\hline
+\end{tabular}
+\caption{}
+\end{table}
+
+Add Figure 1 which compares single best at $10^8$ ms (Glasgow 2) with LLAMA ?
+
 \section{Conclusion and Future Work}

### 2015-12-09 paper/paper.tex
more experimental setup

@@
 
+We use 10-fold cross-validation to determine the performance of the LLAMA
+models. The entire set of instances was randomly partitioned into 10 subsets of
+approximately equal size. Of the 10 subsets, 9 were combined to form the
+training set for the algorithm selection models, which were evaluated on the
+remaining subset. This process was repeated 10 times for all possible
+combinations of training and test sets. At the end of this process, each problem
+instance in the original set was used exactly once to evaluate the performance
+of the algorithm selection models.
+
 LLAMA's pairwise regression approach with random forest regression gave the best

### 2015-12-09 portfolio/as-graphs.Rnw
include more information

@@
 resvbs = data.frame(model = "virtual best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, vbs)),
+    solved = sum(successes(ldf, vbs)),
     mean.performance = mean(parscores(ldf, vbs, factor = 1)),
@@
 ressb = data.frame(model = "single best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, singleBest)),
+    solved = sum(successes(ldf, singleBest)),
     mean.performance = mean(parscores(ldf, singleBest, factor = 1)),
@@
 resc = data.frame(model = "classification",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, modelc)),
+    solved = sum(successes(ldf, modelc)),
     mean.performance = mean(parscores(ldf, modelc, factor = 1)),
@@
 resr = data.frame(model = "regression",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, modelr)),
+    solved = sum(successes(ldf, modelr)),
     mean.performance = mean(parscores(ldf, modelr, factor = 1)),
@@
 rescp = data.frame(model = "classifyPairs",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, modelcp)),
+    solved = sum(successes(ldf, modelcp)),
     mean.performance = mean(parscores(ldf, modelcp, factor = 1)),
@@
 resrp = data.frame(model = "regressionPairs",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, modelrp)),
+    solved = sum(successes(ldf, modelrp)),
     mean.performance = mean(parscores(ldf, modelrp, factor = 1)),
@@
 
-pmod = mod$predictions
-pmod$portfolio = parscores(ldf, mod, factor = 1)
-pmod$iteration = NULL
-pmod$algorithm = NULL
-pmod$score = NULL
+pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf, mod, factor = 1))
 
@@
 wide = gather_(perfs, "solver", "time", names(perfs)[-1])
-ggplot(wide, aes(x = time, col = solver)) + stat_ecdf() + scale_x_log10() +
-    ylab("fraction of instances solved")
+wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
+    ylab("fraction of instances solved") +
+    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
 @
@@
 resvbs = data.frame(model = "virtual best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf3, vbs)),
+    solved = sum(successes(ldf3, vbs)),
     mean.performance = mean(parscores(ldf3, vbs, factor = 1)),
@@
 ressb = data.frame(model = "single best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf3, singleBest)),
+    solved = sum(successes(ldf3, singleBest)),
     mean.performance = mean(parscores(ldf3, singleBest, factor = 1)),
@@
 resc = data.frame(model = "classification",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf3, modelc)),
+    solved = sum(successes(ldf3, modelc)),
     mean.performance = mean(parscores(ldf3, modelc, factor = 1)),
@@
 resr = data.frame(model = "regression",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf3, modelr)),
+    solved = sum(successes(ldf3, modelr)),
     mean.performance = mean(parscores(ldf3, modelr, factor = 1)),
@@
 rescp = data.frame(model = "classifyPairs",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf3, modelcp)),
+    solved = sum(successes(ldf3, modelcp)),
     mean.performance = mean(parscores(ldf3, modelcp, factor = 1)),
@@
 resrp = data.frame(model = "regressionPairs",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf3, modelrp)),
+    solved = sum(successes(ldf3, modelrp)),
     mean.performance = mean(parscores(ldf3, modelrp, factor = 1)),
@@
 print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolved instances removed."), booktabs = TRUE, include.rownames = FALSE)
+@
+
+<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
+addmcp = rep.int(0, length(presolved))
+addsucc = rep.int(1, length(presolved))
+addtime = subset(ldf$data, ldf$data$instance_id %in% presolved)$lad_features
+
+resvbs = data.frame(model = "virtual best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, vbs)),
+    solved = sum(successes(ldf, vbs)),
+    mean.performance = mean(parscores(ldf, vbs, factor = 1)),
+    median.performance = median(parscores(ldf, vbs, factor = 1)))
+ressb = data.frame(model = "single best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, singleBest)),
+    solved = sum(successes(ldf, singleBest)),
+    mean.performance = mean(parscores(ldf, singleBest, factor = 1)),
+    median.performance = median(parscores(ldf, singleBest, factor = 1)))
+
+resc = data.frame(model = "classification",
+    mean.misclassification.penalty = mean(c(misclassificationPenalties(ldf3, modelc), addmcp)),
+    solved = sum(c(successes(ldf3, modelc), addsucc)),
+    mean.performance = mean(c(parscores(ldf3, modelc, factor = 1), addtime)),
+    median.performance = median(c(parscores(ldf3, modelc, factor = 1), addtime)))
+
+resr = data.frame(model = "regression",
+    mean.misclassification.penalty = mean(c(misclassificationPenalties(ldf3, modelr), addmcp)),
+    solved = sum(c(successes(ldf3, modelr), addsucc)),
+    mean.performance = mean(c(parscores(ldf3, modelr, factor = 1), addtime)),
+    median.performance = median(c(parscores(ldf3, modelr, factor = 1), addtime)))
+
+rescp = data.frame(model = "classifyPairs",
+    mean.misclassification.penalty = mean(c(misclassificationPenalties(ldf3, modelcp), addmcp)),
+    solved = sum(c(successes(ldf3, modelcp), addsucc)),
+    mean.performance = mean(c(parscores(ldf3, modelcp, factor = 1), addtime)),
+    median.performance = median(c(parscores(ldf3, modelcp, factor = 1), addtime)))
+
+resrp = data.frame(model = "regressionPairs",
+    mean.misclassification.penalty = mean(c(misclassificationPenalties(ldf3, modelrp), addmcp)),
+    solved = sum(c(successes(ldf3, modelrp), addsucc)),
+    mean.performance = mean(c(parscores(ldf3, modelrp, factor = 1), addtime)),
+    median.performance = median(c(parscores(ldf3, modelrp, factor = 1), addtime)))
+
+tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
+print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolved instances removed (evaluated on all instances)."), booktabs = TRUE, include.rownames = FALSE)
 @
@@
 
-pmod = mod$predictions
-pmod$portfolio = parscores(ldf3, mod, factor = 1)
-pmod$iteration = NULL
-pmod$algorithm = NULL
-pmod$score = NULL
+pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf3, mod, factor = 1))
 
@@
 wide = gather_(perfs, "solver", "time", names(perfs)[-1])
-ggplot(wide, aes(x = time, col = solver)) + stat_ecdf() + scale_x_log10() +
-    ylab("fraction of instances solved")
+wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
+    ylab("fraction of instances solved") +
+    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+@
+
+<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio for all instances (best model).'>>=
+vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf), function(ss) { ss[1] })
+vbs.agg$virtual.best = parscores(ldf, vbs, factor = 1)
+vbs.agg$iteration = NULL
+vbs.agg$score = NULL
+
+pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf3, mod, factor = 1))
+pmod = rbind(pmod, data.frame(instance_id = factor(presolved), portfolio = subset(ldf$data, ldf$data$instance_id %in% presolved)$lad_features))
+
+perfs = subset(ldf$data, TRUE, c("instance_id", ldf$performance))
+perfs = merge(perfs, vbs.agg, by = "instance_id")
+perfs = merge(perfs, pmod, by = "instance_id")
+wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
+    ylab("fraction of instances solved") +
+    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
 @
@@
 resvbs = data.frame(model = "virtual best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf2, vbs)),
+    solved = sum(successes(ldf2, vbs)),
     mean.performance = mean(parscores(ldf2, vbs, factor = 1)),
@@
 ressb = data.frame(model = "single best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf2, singleBest)),
+    solved = sum(successes(ldf2, singleBest)),
     mean.performance = mean(parscores(ldf2, singleBest, factor = 1)),
@@
 resc = data.frame(model = "classification",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf2, modelc)),
+    solved = sum(successes(ldf2, modelc)),
     mean.performance = mean(parscores(ldf2, modelc, factor = 1)),
@@
 resr = data.frame(model = "regression",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf2, modelr)),
+    solved = sum(successes(ldf2, modelr)),
     mean.performance = mean(parscores(ldf2, modelr, factor = 1)),
@@
 rescp = data.frame(model = "classifyPairs",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf2, modelcp)),
+    solved = sum(successes(ldf2, modelcp)),
     mean.performance = mean(parscores(ldf2, modelcp, factor = 1)),
@@
 resrp = data.frame(model = "regressionPairs",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf2, modelrp)),
+    solved = sum(successes(ldf2, modelrp)),
     mean.performance = mean(parscores(ldf2, modelrp, factor = 1)),
@@
 print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolver."), booktabs = TRUE, include.rownames = FALSE)
+@
+
+<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
+vf2pre = setdiff(ldf3$data$instance_id, ldf2$data$instance_id)
+addmcp = rep.int(0, length(presolved) + length(vf2pre))
+addsucc = rep.int(1, length(presolved) + length(vf2pre))
+addtime = c(subset(ldf$data, ldf$data$instance_id %in% presolved)$lad_features, subset(ldf$data, ldf$data$instance_id %in% vf2pre)$vf2)
+
+resvbs = data.frame(model = "virtual best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, vbs)),
+    solved = sum(successes(ldf, vbs)),
+    mean.performance = mean(parscores(ldf, vbs, factor = 1)),
+    median.performance = median(parscores(ldf, vbs, factor = 1)))
+ressb = data.frame(model = "single best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf, singleBest)),
+    solved = sum(successes(ldf, singleBest)),
+    mean.performance = mean(parscores(ldf, singleBest, factor = 1)),
+    median.performance = median(parscores(ldf, singleBest, factor = 1)))
+
+resc = data.frame(model = "classification",
+    mean.misclassification.penalty = mean(c(misclassificationPenalties(ldf2, modelc), addmcp)),
+    solved = sum(c(successes(ldf2, modelc), addsucc)),
+    mean.performance = mean(c(parscores(ldf2, modelc, factor = 1), addtime)),
+    median.performance = median(c(parscores(ldf2, modelc, factor = 1), addtime)))
+
+resr = data.frame(model = "regression",
+    mean.misclassification.penalty = mean(c(misclassificationPenalties(ldf2, modelr), addmcp)),
+    solved = sum(c(successes(ldf2, modelr), addsucc)),
+    mean.performance = mean(c(parscores(ldf2, modelr, factor = 1), addtime)),
+    median.performance = median(c(parscores(ldf2, modelr, factor = 1), addtime)))
+
+rescp = data.frame(model = "classifyPairs",
+    mean.misclassification.penalty = mean(c(misclassificationPenalties(ldf2, modelcp), addmcp)),
+    solved = sum(c(successes(ldf2, modelcp), addsucc)),
+    mean.performance = mean(c(parscores(ldf2, modelcp, factor = 1), addtime)),
+    median.performance = median(c(parscores(ldf2, modelcp, factor = 1), addtime)))
+
+resrp = data.frame(model = "regressionPairs",
+    mean.misclassification.penalty = mean(c(misclassificationPenalties(ldf2, modelrp), addmcp)),
+    solved = sum(c(successes(ldf2, modelrp), addsucc)),
+    mean.performance = mean(c(parscores(ldf2, modelrp, factor = 1), addtime)),
+    median.performance = median(c(parscores(ldf2, modelrp, factor = 1), addtime)))
+
+tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
+print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolver (evaluated on all instances)."), booktabs = TRUE, include.rownames = FALSE)
 @
@@
 
-pmod = mod$predictions
-pmod$portfolio = parscores(ldf2, mod, factor = 1)
-pmod$iteration = NULL
-pmod$algorithm = NULL
-pmod$score = NULL
+pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf2, mod, factor = 1))
 
@@
 wide = gather_(perfs, "solver", "time", names(perfs)[-1])
-ggplot(wide, aes(x = time, col = solver)) + stat_ecdf() + scale_x_log10() +
-    ylab("fraction of instances solved")
+wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
+    ylab("fraction of instances solved") +
+    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+@
+
+<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio for all instances (best model).'>>=
+vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf), function(ss) { ss[1] })
+vbs.agg$virtual.best = parscores(ldf, vbs, factor = 1)
+vbs.agg$iteration = NULL
+vbs.agg$score = NULL
+
+pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf2, mod, factor = 1))
+pmod = rbind(pmod, data.frame(instance_id = factor(presolved), portfolio = subset(ldf$data, ldf$data$instance_id %in% presolved)$lad_features))
+vf2pre = setdiff(ldf3$data$instance_id, ldf2$data$instance_id)
+pmod = rbind(pmod, data.frame(instance_id = factor(vf2pre), portfolio = subset(ldf$data, ldf$data$instance_id %in% vf2pre)$vf2))
+
+perfs = subset(ldf$data, TRUE, c("instance_id", ldf$performance))
+perfs = merge(perfs, vbs.agg, by = "instance_id")
+perfs = merge(perfs, pmod, by = "instance_id")
+wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
+    ylab("fraction of instances solved") +
+    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
 @

### 2015-12-11 paper/paper.tex
Change Glasgow description a bit

@@
 \item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques (AllDifferent Generalized Arc Consistency and LAD);
-\item Glasgow \cite{McCreesh:2015}, which does an expensive preprocessing based on path length invariant properties, followed by weaker filtering (forward-checking) during search.
-We consider 4 different instanciations of Glasgow, with path lengths varying from 1 to 4: The longer the path, the stronger the reduction, and the more expensive to achieve.
+\item Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
+    invariant properties to generate additional constraints, followed by weaker filtering
+    (forward-checking, and a heuristic AllDifferent propagator which can miss deletions) and
+    conflict-directed backjumping during search.
 \end{itemize}
 
-The Glasgow algorithm has a parameter, which controls the lengths of paths used
-when reasoning about non-adjacent vertices.  In experiments reported in
-\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable
-compromise -- longer paths lead to prohibitively expensive preprocessing on
-larger, denser instances. This is often not the best choice on an instance by
-instance basis: sometimes path-based reasoning gives no benefit at all,
-sometimes considering only paths of length 2 suffices, occasionally paths of
-length 4 are helpful, and even looking at paths of length 3 is relatively
-expensive on some graphs.
+\noindent The Glasgow algorithm has a parameter, which controls the lengths of paths used when
+reasoning about non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice
+of paths of length 3 was used as a reasonable compromise---longer paths lead to prohibitively
+expensive preprocessing on larger, denser instances. This is often not the best choice on an
+instance by instance basis: sometimes path-based reasoning gives no benefit at all, sometimes
+considering only paths of length 2 suffices, occasionally paths of length 4 are helpful, and even
+looking at paths of length 3 is relatively expensive on some graphs.
 
@@
 space. In particular, the relative performance of the algorithms is monotone in
-the time it takes to solve the instance -- path length 1 is best at first, but
+the time it takes to solve the instance---path length 1 is best at first, but
 gets worse as the time to solve the instance increases until path length 2 is

### 2015-12-11 paper/paper.tex
Rework features and selection a bit

@@
 
-Our approach is composed of 3 steps described below: First, we run two presolvers in a static way to quickly solve easy instances; Second, we extract features from instances which are not solved by the first step; Finally, we select an algorithm and run it.
+Our approach is composed of 3 steps described below. First, we run two presolvers in a static way to
+quickly solve easy instances; second, we extract features from instances which are not solved by the
+first step; and finally, we select an algorithm and run it.
 
-\paragraph{Static steps.}
+\subsection{Presolving}
 
-Experimental results reported in Section \ref{expComp} have shown us that incompleteLAD is very fast (7 ms on average) and is able to very quickly solve 1919 instances.
-Therefore, we propose to first run incompleteLAD: If the instance is solved, then we do not process it further.
-Furthermore, running IncompleteLAD allows us to collect features on the number of candidate couples which are removed by the filtering.
+Experimental results reported in Section \ref{expComp} have shown us that incompleteLAD is very fast
+(7 ms on average) and is able to very quickly solve 1919 instances. Therefore, we first run
+incompleteLAD: if the instance is detected to be unsatisfiable, then we do not process it further.
+Otherwise, we record as features the number of candidate couples removed by incompleteLAD in absolute
+terms, as a percentage, and the minimum and maximum on a per-variable basis.
 
-VF2 is also able to solve very quickly many easy instances: Among the 3806 instances which are not solved by incompleteLAD, 1470 are solved by VF2 in less than 50 milliseconds.
-Therefore, after running incompleteLAD, we run the VF2 solver for 50 ms.
-This solves easy instances without the overhead of running algorithm selection and avoids potentially making incorrect solver choices.
-After running incompleteLAD and VF2 for 50 ms, we are left with 2336 hard instances that we consider for algorithm selection.
+VF2 is also able to solve many easy instances very quickly: among the 3806 instances which are not
+solved by incompleteLAD, 1470 are solved by VF2 in less than 50 milliseconds. Therefore, after
+running incompleteLAD, we run the VF2 solver for 50 ms. This solves easy instances without the
+overhead of running algorithm selection and avoids potentially making incorrect solver choices.
+After running incompleteLAD and VF2 for 50 ms, we are left with 2336 hard instances that we consider
+for algorithm selection.
 
-\subsection{Features}
+\subsection{Feature Extraction}
 
-For both the pattern and the target graph, we consider the number of vertices,
-edges and loops, the mean and maximum degree, the density, whether or not every
-vertex has same degree, whether or not the graph is connected, the mean and
-maximum distances between all pairs of vertices, and the proportion of vertex
-pairs which are distance at least 2, 3 and 4 apart. We also consider the number
-of candidate couples removed by incompleteLAD in absolute terms, as
-a percentage, and the minimum and maximum on a per-variable basis. This gives
-us a total of 30 features.
+If presolving does not give us a solution, we extract additional features. For both the pattern and
+the target graph, we consider some basic graph properties, which may be computed very quickly:
 
-In addition to computing feature values, the LAD preprocessing is able to detect
-whether an instance is unsatisfiable. In this case, it is not necessary to run
-an algorithm on the instance.
+\begin{itemize}
+    \item The number of vertices and edges.
+    \item The density---we expect that some kinds of filtering might be expensive and ineffective on
+        dense graphs.
+    \item How many loops (self-adjacent vertices) the graph contains---since loops must be mapped to
+        loops, this could have a strong effect on how easy an instance is.
+    \item The mean and maximum degrees, and whether or not every vertex has the same degree. (The
+        degree-based invariants used by LAD and Glasgow do nothing at the top of search if every
+        vertex has the same degree.)
+    \item Whether or not the graph is connected.
+    \item The mean and maximum distances between all pairs of vertices (if nearly all vertices are
+        close together, path-based reasoning is likely to be ineffective). We also consider the
+        proportion of vertex pairs which are distance at least 2, 3 and 4 apart.
+\end{itemize}
 
-We apply two static steps to each instance before solving it. First, the
-computation of the LAD features can detect unsatisfiability of an instance. If
-an instance is identified as unsatisfiable, we do not process it further.
-Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
-without the overhead of running algorithm selection and avoids potentially
-making incorrect solver choices.
-
-Computing the LAD features detects unsatisfiability on 1919 instances. Running
-VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
-instances that we consider for algorithm selection.
+\subsection{Solver Selection}
 

### 2015-12-11 portfolio/as-graphs.Rnw
more algsel variants, tweaked plots, everything in milliseconds

@@
 library(tidyr)
+library(ggthemes)
+library(scales)
 
@@
 tab = tab[,order(names(tab))]
-print(xtable(tab, digits = 3, caption = "Mean performance by family."), booktabs = TRUE, include.rownames = FALSE)
+print(xtable(tab, digits = 0, caption = "Mean performance by family."), booktabs = TRUE, include.rownames = FALSE)
 
@@
 tab = tab[,order(names(tab))]
-print(xtable(tab, digits = 3, caption = "Median performance by family."), booktabs = TRUE, include.rownames = FALSE)
+print(xtable(tab, digits = 0, caption = "Median performance by family."), booktabs = TRUE, include.rownames = FALSE)
 
@@
 res = summarizeAlgoPerf(asscenario)
-print(xtable(res, digits = 6, display = c("s", "d", "d", rep("g", 8))), booktabs = TRUE)
+print(xtable(res, digits = 0, display = c("s", "d", "d", rep("g", 8))), booktabs = TRUE)
 @
@@
 res = summarizeAlgoRunstatus(asscenario)
-print(xtable(res, digits = 3), booktabs = TRUE)
+print(xtable(res, digits = 0), booktabs = TRUE)
 @
@@
 tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
-print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results."), booktabs = TRUE, include.rownames = FALSE)
+print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results."), booktabs = TRUE, include.rownames = FALSE)
 @
@@
     scale_linetype_manual(values=c(3,1), guide = FALSE) +
-    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
-    ylab("fraction of instances solved") +
-    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='sans', base_size = 12) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.1), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
 @
@@
 tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
-print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolved instances removed."), booktabs = TRUE, include.rownames = FALSE)
+print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results with presolved instances removed."), booktabs = TRUE, include.rownames = FALSE)
 @
@@
 tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
-print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolved instances removed (evaluated on all instances)."), booktabs = TRUE, include.rownames = FALSE)
+print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results with presolved instances removed (evaluated on all instances)."), booktabs = TRUE, include.rownames = FALSE)
 @
@@
 ldf2 = ldf3
-presolvetime = 0.05
+presolvetime = 50
 ldf2$data = subset(ldf3$data, ldf3$data$vf2 > presolvetime)
@@
 tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
-print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolver."), booktabs = TRUE, include.rownames = FALSE)
+print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results with presolver."), booktabs = TRUE, include.rownames = FALSE)
 @
@@
 tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
-print(xtable(tab, digits = 3, caption = "Preliminary algorithm selection results with presolver (evaluated on all instances)."), booktabs = TRUE, include.rownames = FALSE)
+print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results with presolver (evaluated on all instances)."), booktabs = TRUE, include.rownames = FALSE)
 @
@@
     scale_linetype_manual(values=c(3,1), guide = FALSE) +
-    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
-    ylab("fraction of instances solved") +
-    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='sans', base_size = 12) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.1), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+@
+
+\clearpage
+
+\subsection*{Only Glasgow variants}
+
+<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
+ldf.glasgow = ldf2
+ldf.glasgow$performance = c("glasgow1", "glasgow2", "glasgow3", "glasgow4")
+ldf.glasgow$success = c("glasgow1_success", "glasgow2_success", "glasgow3_success", "glasgow4_success")
+
+pnames = ldf.glasgow$performance
+snames = ldf.glasgow$success
+optfun = min
+best = apply(ldf.glasgow$data, 1,
+    function(x) {
+        tosel = pnames
+        nosuccs = sapply(snames[which(as.logical(x[snames]) == FALSE)], function(y) { unlist(strsplit(y, "_"))[1] })
+        tosel = setdiff(pnames, nosuccs)
+        if(length(tosel) == 0) {
+            # nothing was able to solve this instance
+            return(NA)
+        } else {
+            perfs = as.numeric(x[tosel])
+            return(tosel[which(perfs == optfun(perfs))])
+        }
+    })
+# simplify...
+names(best) = NULL
+ldf.glasgow$best = best
+
+resvbs = data.frame(model = "virtual best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, vbs)),
+    solved = sum(successes(ldf.glasgow, vbs)),
+    mean.performance = mean(parscores(ldf.glasgow, vbs, factor = 1)),
+    median.performance = median(parscores(ldf.glasgow, vbs, factor = 1)))
+ressb = data.frame(model = "single best",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, singleBest)),
+    solved = sum(successes(ldf.glasgow, singleBest)),
+    mean.performance = mean(parscores(ldf.glasgow, singleBest, factor = 1)),
+    median.performance = median(parscores(ldf.glasgow, singleBest, factor = 1)))
+
+modelc = classify(rflearner, ldf.glasgow)
+resc = data.frame(model = "classification",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, modelc)),
+    solved = sum(successes(ldf.glasgow, modelc)),
+    mean.performance = mean(parscores(ldf.glasgow, modelc, factor = 1)),
+    median.performance = median(parscores(ldf.glasgow, modelc, factor = 1)))
+times = sapply(modelc$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
+print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classification."), booktabs = TRUE)
+
+modelr = regression(makeLearner("regr.randomForest"), ldf.glasgow)
+resr = data.frame(model = "regression",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, modelr)),
+    solved = sum(successes(ldf.glasgow, modelr)),
+    mean.performance = mean(parscores(ldf.glasgow, modelr, factor = 1)),
+    median.performance = median(parscores(ldf.glasgow, modelr, factor = 1)))
+times = sapply(modelr$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
+print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regression."), booktabs = TRUE)
+
+modelcp = classifyPairs(rflearner, ldf.glasgow)
+rescp = data.frame(model = "classifyPairs",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, modelcp)),
+    solved = sum(successes(ldf.glasgow, modelcp)),
+    mean.performance = mean(parscores(ldf.glasgow, modelcp, factor = 1)),
+    median.performance = median(parscores(ldf.glasgow, modelcp, factor = 1)))
+times = sapply(modelcp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
+print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classificationPairs."), booktabs = TRUE)
+
+modelrp = regressionPairs(makeLearner("regr.randomForest"), ldf.glasgow)
+resrp = data.frame(model = "regressionPairs",
+    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, modelrp)),
+    solved = sum(successes(ldf.glasgow, modelrp)),
+    mean.performance = mean(parscores(ldf.glasgow, modelrp, factor = 1)),
+    median.performance = median(parscores(ldf.glasgow, modelrp, factor = 1)))
+times = sapply(modelrp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
+print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regressionPairs."), booktabs = TRUE)
+
+tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
+print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results with presolver for Glasgow solvers."), booktabs = TRUE, include.rownames = FALSE)
+@
+
+<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio (best model).'>>=
+vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf.glasgow), function(ss) { ss[1] })
+vbs.agg$virtual.best = parscores(ldf.glasgow, vbs, factor = 1)
+vbs.agg$iteration = NULL
+vbs.agg$score = NULL
+
+pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf.glasgow, mod, factor = 1))
+
+perfs = subset(ldf.glasgow$data, TRUE, c("instance_id", ldf.glasgow$performance))
+perfs = merge(perfs, vbs.agg, by = "instance_id")
+perfs = merge(perfs, pmod, by = "instance_id")
+wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='sans', base_size = 12) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.1), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+@
+
+<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio for all instances (best model).'>>=
+vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf), function(ss) { ss[1] })
+vbs.agg$virtual.best = parscores(ldf, vbs, factor = 1)
+vbs.agg$iteration = NULL
+vbs.agg$score = NULL
+
+pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf.glasgow, mod, factor = 1))
+pmod = rbind(pmod, data.frame(instance_id = factor(presolved), portfolio = subset(ldf$data, ldf$data$instance_id %in% presolved)$lad_features))
+vf2pre = setdiff(ldf3$data$instance_id, ldf.glasgow$data$instance_id)
+pmod = rbind(pmod, data.frame(instance_id = factor(vf2pre), portfolio = subset(ldf$data, ldf$data$instance_id %in% vf2pre)$vf2))
+
+perfs = subset(ldf$data, TRUE, c("instance_id", ldf$performance))
+perfs = merge(perfs, vbs.agg, by = "instance_id")
+perfs = merge(perfs, pmod, by = "instance_id")
+wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='sans', base_size = 12) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.1), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+@
+
+\clearpage
+
+\section*{Different timeouts}
+
+Consider timeouts $10^2$\ldots$10^8$ (after filtering and presolving).
+
+<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
+timeoutExps = 2:8
+
+tab = rbind.fill(lapply(timeoutExps, function(to) {
+    ldf.tmp = ldf2
+    if(10^to < asscenario$desc$algorithm_cutoff_time) {
+        for(p in ldf.tmp$performance) {
+            ldf.tmp$data[paste0(p, "_success")] = sapply(ldf.tmp$data[,p], function(x) x <= 10^to)
+            ldf.tmp$data[p] = sapply(ldf.tmp$data[,p], function(x) min(x, 10^to))
+        }
+        pnames = ldf.tmp$performance
+        snames = ldf.tmp$success
+        optfun = min
+        best = apply(ldf.tmp$data, 1,
+            function(x) {
+                tosel = pnames
+                nosuccs = sapply(snames[which(as.logical(x[snames]) == FALSE)], function(y) { unlist(strsplit(y, "_"))[1] })
+                tosel = setdiff(pnames, nosuccs)
+                if(length(tosel) == 0) {
+                    # nothing was able to solve this instance
+                    return(NA)
+                } else {
+                    perfs = as.numeric(x[tosel])
+                    return(tosel[which(perfs == optfun(perfs))])
+                }
+            })
+        names(best) = NULL
+        ldf.tmp$best = best
+    }
+    ldf.tmp$train = NULL
+    ldf.tmp$test = NULL
+    ldf.tmp = cvFolds(ldf.tmp)
+
+    resvbs = data.frame(model = "virtual best",
+        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, vbs)),
+        solved = sum(successes(ldf.tmp, vbs)),
+        mean.performance = mean(parscores(ldf.tmp, vbs, factor = 1)),
+        median.performance = median(parscores(ldf.tmp, vbs, factor = 1)))
+    ressb = data.frame(model = "single best",
+        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, singleBest)),
+        solved = sum(successes(ldf.tmp, singleBest)),
+        mean.performance = mean(parscores(ldf.tmp, singleBest, factor = 1)),
+        median.performance = median(parscores(ldf.tmp, singleBest, factor = 1)))
+
+    modelc = classify(rflearner, ldf.tmp)
+    resc = data.frame(model = "classification",
+        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, modelc)),
+        solved = sum(successes(ldf.tmp, modelc)),
+        mean.performance = mean(parscores(ldf.tmp, modelc, factor = 1)),
+        median.performance = median(parscores(ldf.tmp, modelc, factor = 1)))
+
+    modelr = regression(makeLearner("regr.randomForest"), ldf.tmp)
+    resr = data.frame(model = "regression",
+        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, modelr)),
+        solved = sum(successes(ldf.tmp, modelr)),
+        mean.performance = mean(parscores(ldf.tmp, modelr, factor = 1)),
+        median.performance = median(parscores(ldf.tmp, modelr, factor = 1)))
+
+    modelcp = classifyPairs(rflearner, ldf.tmp)
+    rescp = data.frame(model = "classifyPairs",
+        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, modelcp)),
+        solved = sum(successes(ldf.tmp, modelcp)),
+        mean.performance = mean(parscores(ldf.tmp, modelcp, factor = 1)),
+        median.performance = median(parscores(ldf.tmp, modelcp, factor = 1)))
+
+    modelrp = regressionPairs(makeLearner("regr.randomForest"), ldf.tmp)
+    resrp = data.frame(model = "regressionPairs",
+        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, modelrp)),
+        solved = sum(successes(ldf.tmp, modelrp)),
+        mean.performance = mean(parscores(ldf.tmp, modelrp, factor = 1)),
+        median.performance = median(parscores(ldf.tmp, modelrp, factor = 1)))
+
+    sb = levels(singleBest(ldf.tmp)$algorithm)
+    cbind(data.frame(timeout = 10^to, single.best = sb), rbind(resvbs, ressb, resc, resr, rescp, resrp))
+}))
+
+print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results for different timeouts with presolver."), booktabs = TRUE, include.rownames = FALSE)
 @

### 2015-12-11 portfolio/convertToAslib.R

@@ -17,7 +17,7 @@ times$family = NULL
 times$vbs = NULL
 times$sat = NULL
 
-timeout = 1e8/1000
+timeout = 1e8
 algorithms = names(times)[-1]
 
 feats.cheap.pattern = names(features.cheap)[grep("pattern.", names(features.cheap), fixed = TRUE)][-1]
@@ -37,14 +37,14 @@ feats.lad = feats.lad[!(feats.lad %in% c(costs.lad, presolved.lad))]
 
 # convert ms to s to avoid integer overflows later
 for (alg in algorithms) {
-    times[,alg] = times[,alg] / 1000
+    times[,alg] = times[,alg]
     times[,alg][times[,alg] >= timeout] = NA
 }
-features.cheap[,costs.cheap.pattern] = features.cheap[,costs.cheap.pattern] / 1000
-features.cheap[,costs.cheap.target] = features.cheap[,costs.cheap.target] / 1000
-features.distance[,costs.distance.pattern] = features.distance[,costs.distance.pattern] / 1000
-features.distance[,costs.distance.target] = features.distance[,costs.distance.target] / 1000
-features.lad[,costs.lad] = features.lad[,costs.lad] / 1000
+features.cheap[,costs.cheap.pattern] = features.cheap[,costs.cheap.pattern] 
+features.cheap[,costs.cheap.target] = features.cheap[,costs.cheap.target]
+features.distance[,costs.distance.pattern] = features.distance[,costs.distance.pattern]
+features.distance[,costs.distance.target] = features.distance[,costs.distance.target]
+features.lad[,costs.lad] = features.lad[,costs.lad]
 
 # convert to aslib
 desc = makeS3Obj("ASScenarioDesc",

### 2015-12-11 paper/paper.tex

@@
 \emph{pattern} graph to vertices of a large \emph{target} graph which preserves adjacency. This
-\NP-complete problem has lots of important practical applications, for example in computer vision,
-biochemistry, and model checking. There exist various exact algorithms, which have been compared on
-a large benchmark in  \cite{McCreesh:2015}. These experiments indicated that the single best
-algorithm depends on the CPU time limit considered: for very small (resp. larger) CPU time limits,
-VF2 \cite{Cordella:2004} (resp. the Glasgow algorithm \cite{McCreesh:2015}) has better success
-rates. Furthermore, they also showed us that on an instance by instance basis, other algorithms may
-become better.
+\NP-complete problem has lots of important practical applications, for example in computer vision
+\cite{cviu11,pr15}, biochemistry \cite{Giugno:2013}, and model checking \cite{Sevegnani:2015}. There
+exist various exact algorithms, which have been compared on a large benchmark in
+\cite{McCreesh:2015}. These experiments indicated that the single best algorithm depends on the CPU
+time limit considered: for very small (resp. larger) CPU time limits, VF2 \cite{Cordella:2004}
+(resp. the Glasgow algorithm \cite{McCreesh:2015}) has better success rates. Furthermore, they also
+showed us that on an instance by instance basis, other algorithms may become better.
 
@@
     next 50 graphs.
-\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}:
-    bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs
-    with uniform edge probabilities for class 7.
+\item Classes 4 to 8 contain randomly generated graphs coming from
+    \cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
+    for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
 \item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
@@
 The figure illustrates the potential for portfolios and algorithm selection we have. There is clearly no single solver that dominates throughout.
-Furthermore, the Virtual Best Solver (VBS), which considers the best algorithm for each instance separately, obtains much better results, showing us that the algorithms have complementary performance.
-The difference between VBS and single best is more particularly important for rather small CPU time limits (smaller than 1000 ms).
-In many applications, it is important to have the fastest possible algorithm.
-For example, in pattern recognition applications such as those reported in \cite{pr15,cviu11}, we often have to solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern image in a large database of target images, for example), so that having an algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference.
-Therefore, it is important to select the best algorithm for each instance, even if the instance is an easy one.
+
+Furthermore, the Virtual Best Solver (VBS), which considers the best algorithm for each instance
+separately, obtains much better results, showing us that the algorithms have complementary
+performance. The difference between VBS and single best is more particularly important for rather
+small CPU time limits (smaller than 1000 ms). In many applications, it is important to have the
+fastest possible algorithm. For example, in pattern recognition applications such as those reported
+in \cite{pr15,cviu11}, and in chemical applications such as \cite{Giugno:2013}, we often have to
+solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern
+image or molecule in a large database of target images or compounds, for example), so that having an
+algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference.
+Therefore, it is important to select the best algorithm for each instance, even if the instance is
+an easy one.
 

### 2015-12-11 portfolio/convertToAslib.R
use times to compute features as features

@@ -20,22 +20,21 @@ times$sat = NULL
 timeout = 1e8
 algorithms = names(times)[-1]
 
-feats.cheap.pattern = names(features.cheap)[grep("pattern.", names(features.cheap), fixed = TRUE)][-1]
+feats.cheap.pattern = names(features.cheap)[grep("pattern.", names(features.cheap), fixed = TRUE)]
 costs.cheap.pattern = "cheap.pattern.time"
-feats.cheap.target = names(features.cheap)[grep("target.", names(features.cheap), fixed = TRUE)][-1]
+feats.cheap.target = names(features.cheap)[grep("target.", names(features.cheap), fixed = TRUE)]
 costs.cheap.target = "cheap.target.time"
 
-feats.distance.pattern = names(features.distance)[grep("pattern.", names(features.distance), fixed = TRUE)][-1]
+feats.distance.pattern = names(features.distance)[grep("pattern.", names(features.distance), fixed = TRUE)]
 costs.distance.pattern = "distance.pattern.time"
-feats.distance.target = names(features.distance)[grep("target.", names(features.distance), fixed = TRUE)][-1]
+feats.distance.target = names(features.distance)[grep("target.", names(features.distance), fixed = TRUE)]
 costs.distance.target = "distance.target.time"
 
 feats.lad = names(features.lad)[grep("lad.", names(features.lad), fixed = TRUE)]
 costs.lad = "lad.time"
 presolved.lad = "lad.detected.inconsistent"
-feats.lad = feats.lad[!(feats.lad %in% c(costs.lad, presolved.lad))]
+feats.lad = feats.lad[!(feats.lad %in% c(presolved.lad))]
 
-# convert ms to s to avoid integer overflows later
 for (alg in algorithms) {
     times[,alg] = times[,alg]
     times[,alg][times[,alg] >= timeout] = NA
@@ -63,7 +62,7 @@ desc = makeS3Obj("ASScenarioDesc",
   features_cutoff_time = NA,
   features_cutoff_memory = NA,
 
-  number_of_feature_steps = 5,
+  number_of_feature_steps = 5L,
   default_steps = c("cheap_pattern", "cheap_target", "distance_pattern", "distance_target", "lad_features"),
   feature_steps =
       list(cheap_pattern = list(provides = feats.cheap.pattern),

### 2015-12-11 paper/paper.tex
Patrick doesn't love us any more

@@
         and Physical Sciences Research Council [grant number EP/K503058/1]}\inst{2}
-    \and Patrick Prosser\inst{2}
     \and Christine Solnon\inst{3}}
@@
 
-\paragraph{Filtering for subgraph isomorphism.}
+\subsection{Filtering for subgraph isomorphism}
 
@@
 
-\paragraph{Propagation of invariant properties.}
+\subsection{Propagation of invariant properties}
 
@@
 
-\paragraph{Algorithms selected in our portfolio.}
+\subsection{Per-Instance Algorithm Selection}
+
+The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
+algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
+best on a given problem instance. Algorithm selection systems usually build machine learning models
+of the algorithms or the portfolio they are contained in to forecast which algorithm to use in a
+particular context. Using the model predictions, one or more algorithms from the portfolio are
+selected to be run sequentially or in parallel.
+
+Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
+the most prominent and successful systems that employs this approach is
+SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
+years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
+salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
+The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
+information on algorithm selection.
+
+\subsection{Problem Instances}
+
+We consider 5725 instances, which are available in a simple text
+format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These
+instances are grouped into 12 classes.
+
+\begin{itemize}
+\item Class 1 contains randomly generated scale-free graphs \cite{constraints10}.
+\item Classes 2 and 3 contain various kinds of graphs coming from \cite{LV02}: class 2 contains small
+    instances generated from the 50 first graphs, whereas class 3 contains larger
+    instances where pattern graphs are coming from class 2, and target graphs are coming from the
+    next 50 graphs.
+\item Classes 4 to 8 contain randomly generated graphs coming from
+    \cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
+    for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
+\item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
+\item Class 11 contains instances coming from meshes modeling 3D objects \cite{cviu11}.
+\item Class 12 contains random graph instances belonging to the phase transition.
+\end{itemize}
+
+Note that Classes 3 and 12 were not considered in the previous experimental
+study reported in \cite{McCreesh:2015}.  This set of instances is also much
+larger than that of Battiti and Mascia~\cite{battiti-mascia07}, who were the
+first to propose algorithm portfolios for subgraph isomorphism problems.
+Battiti and Mascia only considered a pure parallel portfolio consisting of two
+solvers, and without a selection mechanism. Their problem set consisted
+entirely of easy satisfiable instances, which is where this style of approach
+has most success.
+
+\subsection{Experimental setup}
+
+We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
+Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
+the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
+\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
+timeout of $10^8$ milliseconds (a little over a day).
+
+\subsection{Algorithms selected in our portfolio}
+
 We have selected a set of algorithms with complementary performance:
@@
 
-\section{Experimental Comparison of Algorithms of the Portfolio}\label{expComp}
-
-\paragraph{Benchmark.}
-
-We consider 5725 instances, which are available in a simple text
-format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These
-instances are grouped into 12 classes.
-
-\begin{itemize}
-\item Class 1 contains randomly generated scale-free graphs \cite{constraints10}.
-\item Classes 2 and 3 contain various kinds of graphs coming from \cite{LV02}: class 2 contains small
-    instances generated from the 50 first graphs, whereas class 3 contains larger
-    instances where pattern graphs are coming from class 2, and target graphs are coming from the
-    next 50 graphs.
-\item Classes 4 to 8 contain randomly generated graphs coming from
-    \cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
-    for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
-\item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
-\item Class 11 contains instances coming from meshes modeling 3D objects \cite{cviu11}.
-\item Class 12 contains random graph instances belonging to the phase transition.
-\end{itemize}
-
-Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.
-
-\paragraph{Experimental setup.} We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
-Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
-the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
-\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
-timeout of $10^8$ milliseconds (a little over a day).
+\subsection{Results from Individual Algorithms} \label{expComp}
 
@@
 \includegraphics[width=\textwidth]{courbe.pdf}
-\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms of the portfolio, and a Virtual Best Solver (VBS).\label{expTime}}
+\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms of the portfolio, and a Virtual Best Solver (VBS).\label{expTimeGraph}}
 \end{figure}
 
-\paragraph{Results.}Figure \ref{expTime} displays the evolution of the number of instances solved with respect to CPU time. It shows us that the best solver depends on the time limit
+Figure \ref{expTimeGraph} displays the evolution of the number of instances solved with respect to CPU time. It shows us that the best solver depends on the time limit
 considered.
@@
 \end{table}
-
-\subsection{Per-Instance Algorithm Selection}
-
-The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
-algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
-best on a given problem instance. Algorithm selection systems usually build machine learning models
-of the algorithms or the portfolio they are contained in to forecast which algorithm to use in a
-particular context. Using the model predictions, one or more algorithms from the portfolio are
-selected to be run sequentially or in parallel.
-
-Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
-the most prominent and successful systems that employs this approach is
-SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
-years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
-salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
-The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
-information on algorithm selection.
-
-\medskip
-
-Battiti and Mascia~\cite{battiti-mascia07} were the first to propose
-algorithm portfolios for subgraph isomorphism problems. However, they only
-consider a pure parallel portfolio without a selection mechanism consisting of
-only two solvers. In addition, our instance set is much larger and considers
-both hard and unsatisfiable instances.
 
@@
 4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
-single best if displayed in the last column ($\delta$).\label{expTime}}
+single best if displayed in the last column ($\delta$).\label{expTimeTable}}
 \end{table}
 
-Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging
+Table \ref{expTimeTable} displays the number of instances solved at different CPU time limits, ranging
 from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
@@
 showing us that the algorithms have complementary performance. The difference between VBS and single
-best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU
+best (column $\delta$ of Table \ref{expTimeTable}) is more particularly important for rather small CPU
 time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the

### 2015-12-12 paper/paper.tex
more and updated results

@@
 \centering
-\begin{tabular}{lr}
+\begin{tabular}{lrrr}
   \toprule
-& mean runtime\\
+model & mean MCP & solved instances & mean performance\\
   \midrule
-virtual best & 5822.809\\
-  single best & 7788.719\\
-  LLAMA & 6509.947\\
+virtual best & 0 & 2219 & 5822809\\
+single best & 1965158 & 2172 & 7787967\\
+LLAMA & 610835 & 2206 & 6435301\\
    \bottomrule
@@
 \vspace{1ex}
-\caption{Algorithm selection performance on reduced set of 2336
-instances.}\label{tab:res}
-\end{table}
+\caption{Algorithm selection performance on the reduced set of 2336
+instances. MCP is the misclassification penalty; that is, the additional time
+required to solve a set of instances because of choosing solvers that perform
+worse than the best.}\label{tab:res} \end{table}
 
-We are able to close more than 60\% of the performance gap between the single
-best and the virtual best solver, our lower and upper bounds. For comparison, we
-show the results achieved on the entire set of instances without
-unsatisfiability detection and presolving in Table~\ref{tab:resfull}. We achieve
-similar results with more than 58\% of the gap between the single best and the
-virtual best closed.
+We are able to solve 34 more instances than the single best solver within the
+timeout, with only an additional 13 to the virtual best. In terms of average
+performance, we are able to close almost 70\% of the gap between the single best
+and the virtual best solver.
+
+Table~\ref{tab:res} shows the performance of the selection model. The
+performance of the entire system we propose here is shown in
+Table~\ref{tab:resfull}. The additional instances are solved either by
+incompleteLAD or VF2 presolving, as explained above. Our system is able to close
+more than 50\% of the gap between single and virtual best.
 
@@
 \centering
-\begin{tabular}{lr}
+\begin{tabular}{lrrr}
   \toprule
-& mean runtime\\
+model & mean MCP & solved instances & mean performance\\
   \midrule
-virtual best & 2375.913\\
-  single best & 3177.569\\
-  LLAMA & 2706.749\\
+virtual best & 0 & 5608 & 2375913\\
+single best & 798660 & 5562 & 3174573\\
+LLAMA & 249242 & 5595 & 2625831\\
    \bottomrule
@@
 \vspace{1ex}
-\caption{Algorithm selection performance on the full set of 5725
+\caption{Algorithm selection system performance on the full set of 5725
 instances.}\label{tab:resfull}
 \end{table}
-
-Figure~\ref{fig:scatter} shows detailed results for each of the 2336 instances
-in the reduced set. Our algorithm selection approach is by no means able to
-identify the best solver on a majority of instances, although we are able to
-achieve performance improvements overall. There seems to be no particular
-pattern as to where algorithm selection performs well and where it fails.
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figures/perfScatter}
-\caption{Single best algorithm vs.\ portfolio performance on all instances by
-family. Points above the diagonal indicate that the portfolio achieved better
-performance, while the single best solver was better on instances below the
-diagonal.}
-\label{fig:scatter}
-\end{figure}
 
@@
 through feature computation. As the instances become more difficult to solve,
-its performance improves. For instances that take more than 1,000 seconds to
-solve, the portfolio eclipses any individual solver and closes most of the gap
-to the virtual best solver.
+its performance improves.
 
@@
 \caption{Empirical cumulative distribution function of solvers, virtual best
-solvers, and the LLAMA portfolio.}
+solvers, and the LLAMA portfolio on the reduced set of 2336 instances. The
+actual solvers are shown as dashed lines.}
 \label{fig:portfolio-ecdf}
 \end{figure}
+
+Figure~\ref{fig:portfolio-ecdf-full} shows the ECDF of the system including
+incompleteLAD and VF2 presolving on the full set of instances. Note that for
+very easy instances, we are better than the virtual best solver because
+incompleteLAD is not included in our portfolio. The performance on small
+instances is much better than the LLAMA selector alone (cf.\
+Figure~\ref{fig:portfolio-ecdf}) and the region where the portfolio performs
+worse than the individual solvers is now limited to approximately $10^2$ to
+$10^4$ milliseconds.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figures/portfolio-ecdf-full}
+\caption{Empirical cumulative distribution function of solvers, virtual best
+solvers, and the portfolio system on the full set of 5725 instances. The
+actual solvers are shown as dashed lines.}
+\label{fig:portfolio-ecdf-full}
+\end{figure}
+
+We train the algorithm selection model specifically for the timeout of $10^8$
+milliseconds. In particular, we are interested in minimising the performance
+difference to the virtual best. Problem instances that take longer to solver
+contribute more to this difference than easy instances and therefore carry more
+weight for the algorithm selection model. That is, choosing the wrong solver for
+a hard instances is much worse than choosing the wrong solver for an easy
+instance.
+
+Even though Figures~\ref{fig:portfolio-ecdf} and~\ref{fig:portfolio-ecdf-full}
+show that on easy instances, the portfolio does not perform well, it does over
+the entire set of instances (cf.\ Tables~\ref{tab:res} and~\ref{tab:resfull}).
+
+\subsection{Portfolios for different timeouts}
+
+The question arises of whether the success of our approach is limited to the
+particular timeout we have chosen. Table~\ref{tab:resTimeouts} shows the
+performance of our portfolio for timeouts ranging from $10^2$ to $10^8$. We
+do not include incompleteLAD and VF2 presolving here, as these components do not
+change for different timeouts.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{llrrr}
+  \toprule
+timeout & solver & mean MCP & solved instances & mean performance\\
+  \midrule
+$10^2$ & virtual best & 0 & 1253 & 59\\
+ & LLAMA & 3 & 685 & 78\\
+ & glasgow1 & 11 & 903 & 71\\
+  \midrule
+$10^3$ & virtual best & 0 & 1781 & 322\\
+ & LLAMA & 22 & 1284 & 569\\
+ & glasgow1 & 126 & 1539 & 449\\
+  \midrule
+$10^4$ & virtual best & 0 & 1943 & 2018\\
+ & LLAMA & 186 & 1771 & 3342\\
+ & glasgow2 & 683 & 1855 & 2701\\
+  \midrule
+$10^5$ & virtual best & 0 & 2045 & 14348\\
+ & LLAMA & 1346 & 2022 & 17231\\
+ & glasgow2 & 3019 & 1988 & 17368\\
+  \midrule
+$10^6$ & virtual best & 0 & 2111 & 108060\\
+ & LLAMA & 10826 & 2087 & 120459\\
+ & glasgow2 & 22729 & 2063 & 130789\\
+  \midrule
+$10^7$ & virtual best & 0 & 2178 & 815577\\
+ & LLAMA & 91012 & 2156 & 908200\\
+ & glasgow2 & 198944 & 2129 & 1014522\\
+  \midrule
+$10^8$ & virtual best & 0 & 2219 & 5822809\\
+ & LLAMA & 740765 & 2204 & 6565232\\
+ & glasgow2 & 1969418 & 2172 & 7792227\\
+\bottomrule
+\end{tabular}
+\vspace{1ex}
+\caption{Algorithm selection performance for different timeouts.}\label{tab:resTimeouts}
+\end{table}
+
+The results show that while for the misclassification penalty, LLAMA is always
+better than the single best algorithm, for timeouts $10^2$ to $10^4$ it is worse
+in terms of the number of solved instances and mean performance. An explanation
+for this is that empirical runtime measurements are inherently noisy, which
+affects the accuracy of the measurements particularly in this range.
 

### 2015-12-12 paper/paper.tex
reformatting of table and small fixes

@@
 \usepackage{xcolor,colortbl}
+\usepackage{tabularx}
 
@@
 
-\section{Definitions and notations}
+\section{Definitions and Notations}
 
@@
 
-\section{Subgraph isomorphism algorithms}
+\section{Subgraph Isomorphism Algorithms}
 
@@
 
-\subsection{Filtering for subgraph isomorphism}
+\subsection{Filtering for Subgraph Isomorphism}
 
@@
 
-\subsection{Propagation of invariant properties}
+\subsection{Propagation of Invariant Properties}
 
@@
 
-\subsection{Experimental setup}
+\subsection{Experimental Setup}
 
@@
 
-\subsection{Algorithms selected in our portfolio}
+\subsection{Portfolio Composition}
 
@@
 
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figures/ecdf-glasgow}
-\caption{Empirical cumulative distribution function for the Glasgow algorithm
-with path lengths 1, 2, 3, and 4.}
-\label{fig:ecdf-glasgow}
-\end{figure}
-
-Figure~\ref{fig:ecdf-glasgow} shows the cumulative distribution function for the
-Glasgow algorithm with four different path lengths. As the instances become
-harder to solve, longer paths become better. This is what we expect, as more
-reasoning is expensive, but potentially increases the reduction of the search
-space. In particular, the relative performance of the algorithms is monotone in
-the time it takes to solve the instance---path length 1 is best at first, but
-gets worse as the time to solve the instance increases until path length 2 is
-better. After that point, length 1 is never better than length 2 again. The same
-is true for the other path lengths.
-
-The figure illustrates the potential for portfolios and algorithm selection we
-have. There is clearly no path length that dominates throughout. Therefore, we
-consider 4 different instances of the Glasgow algorithm here, with path lengths
-varying from 1 to 4.
-
 We also introduce two new variants of LAD.
@@
 
-\subsection{Results from Individual Algorithms} \label{expComp}
+\subsection{Results for Individual Algorithms} \label{expComp}
 
@@
 For all classes, there are always at least 2 algorithms which are the best for at least one instance of the class.
-In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for Class 3).
+In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for class 3).
 
@@
 \begin{center}
-\begin{tabular}{|c||r||r|r|r||r|r|r|r|}
-\hline
-Class & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c|}{Glasgow}\\
-&&incomplete&default&path&1&2&3&4\\\hline\hline
-1 &        0 &       20 &        0 &        0 &       80 &        0 &        0 &        0 \\\hline
-2 &      201 &       92 &      189 &      270 &      520 &      180 &       53 &       15 \\\hline
-3 &      112 &     1608 &      617 &      959 &      396 &      195 &       21 &        0 \\\hline
-4 &      270 &        0 &        0 &        0 &        5 &        0 &        0 &        0 \\\hline
-5 &      266 &        0 &        1 &        3 &       31 &        0 &        0 &        0 \\\hline
-6 &       71 &        0 &        0 &        0 &        7 &       14 &        1 &        0 \\\hline
-7 &      270 &        0 &        0 &        0 &        5 &        0 &        0 &        0 \\\hline
-8 &        0 &        0 &        0 &        1 &      195 &       69 &        6 &        0 \\\hline
-9 &       77 &        3 &        0 &       19 &      103 &        1 &        0 &        0 \\\hline
-10 &       13 &        0 &        2 &        2 &        7 &        0 &        0 &        0 \\\hline
-11 &        2 &      142 &       71 &       17 &       23 &        0 &        0 &        0 \\\hline
-12 &        0 &        0 &        1 &        2 &      158 &        6 &        1 &        0 \\\hline\hline
-Total & 1282 & 1865 & 881 & 1273 & 1530 & 465 & 82& 15\\\hline
-\end{tabular}
+\begin{tabularx}{.8\textwidth}{XXrXrrrXrrrr}
+\toprule
+Class && VF2 && \multicolumn{3}{c}{LAD} && \multicolumn{4}{c}{\hspace*{2em}Glasgow}\\
+&&&&incomplete&default&path&&1&2&3&4\\
+\midrule
+1 &&        0 &&       20 &        0 &        0 &&       80 &        0 &        0 &        0 \\
+2 &&      201 &&       92 &      189 &      270 &&      520 &      180 &       53 &       15 \\
+3 &&      112 &&     1608 &      617 &      959 &&      396 &      195 &       21 &        0 \\
+4 &&      270 &&        0 &        0 &        0 &&        5 &        0 &        0 &        0 \\
+5 &&      266 &&        0 &        1 &        3 &&       31 &        0 &        0 &        0 \\
+6 &&       71 &&        0 &        0 &        0 &&        7 &       14 &        1 &        0 \\
+7 &&      270 &&        0 &        0 &        0 &&        5 &        0 &        0 &        0 \\
+8 &&        0 &&        0 &        0 &        1 &&      195 &       69 &        6 &        0 \\
+9 &&       77 &&        3 &        0 &       19 &&      103 &        1 &        0 &        0 \\
+10 &&       13 &&        0 &        2 &        2 &&        7 &        0 &        0 &        0 \\
+11 &&        2 &&      142 &       71 &       17 &&       23 &        0 &        0 &        0 \\
+12 &&        0 &&        0 &        1 &        2 &&      158 &        6 &        1 &        0 \\
+\midrule
+Total && 1282 && 1865 & 881 & 1273 && 1530 & 465 & 82& 15\\
+\bottomrule
+\end{tabularx}
 \end{center}
@@
 
-\subsection{Solver Selection}
+\subsection{Selection Model}
 
@@
 
-\subsection{Portfolios for different timeouts}
+\subsection{Portfolios for Different Timeouts}
 

### 2015-12-12 paper/paper.tex
clarify results for different timeouts

@@
 better than the single best algorithm, for timeouts $10^2$ to $10^4$ it is worse
-in terms of the number of solved instances and mean performance. An explanation
-for this is that empirical runtime measurements are inherently noisy, which
-affects the accuracy of the measurements particularly in this range.
+in terms of the number of solved instances and mean performance. The
+misclassification penalty does not take the cost of computing features into
+account, while the mean performance does -- for these small timeouts, the cost
+of computing the features dominates any improvement through algorithm selection.
+Another explanation for this is that empirical runtime measurements are
+inherently noisy, which affects the accuracy of the measurements and hence the
+performance of the algorithm selection models particularly in this range.
 

### 2015-12-14 paper/paper.tex
General editing

@@
 \usepackage{tabularx}
+\usepackage{siunitx}
 
@@
 \begin{abstract}
-Subgraph isomorphism is a computationally challenging problem with lots of important practical
+Subgraph isomorphism is a computationally challenging problem with important practical
 applications, for example in computer vision, biochemistry, and model checking. There are a number
-of state-of-the-art algorithms for solving it, each of which has its own performance
+of state-of-the-art algorithms for solving the problem, each of which has its own performance
 characteristics. As with many other hard problems, the single best choice of algorithm overall is
-rarely the best algorithm for each instance. We develop a portfolio approach that leverages novel
-features to characterise subgraph isomorphism problems to dynamically decide which algorithm to use
-on a per-instance basis. We demonstrate significant performance improvements on a large set of hard
-benchmark problems. In addition, we show how algorithm selection models can be leveraged to gain new
-insights into what affects the performance of an algorithm.
+rarely the best algorithm on an instance-by-instance. We develop a portfolio approach which leverages
+novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm
+to use on a per-instance basis. We demonstrate significant performance improvements on a large set
+of hard benchmark problems. In addition, we show how algorithm selection models can be leveraged to
+gain new insights into what affects the performance of an algorithm.
 \end{abstract}
@@
 
-The subgraph isomorphism problem is to find an injective mapping from vertices of a small
-\emph{pattern} graph to vertices of a large \emph{target} graph which preserves adjacency. This
-\NP-complete problem has lots of important practical applications, for example in computer vision
+The subgraph isomorphism problem is to find an adjacency-preserving injective mapping from vertices
+of a small \emph{pattern} graph to vertices of a large \emph{target} graph. This \NP-complete
+problem has many important practical applications, for example in computer vision
 \cite{cviu11,pr15}, biochemistry \cite{Giugno:2013}, and model checking \cite{Sevegnani:2015}. There
-exist various exact algorithms, which have been compared on a large benchmark in
-\cite{McCreesh:2015}. These experiments indicated that the single best algorithm depends on the CPU
-time limit considered: for very small (resp. larger) CPU time limits, VF2 \cite{Cordella:2004}
-(resp. the Glasgow algorithm \cite{McCreesh:2015}) has better success rates. Furthermore, they also
-showed us that on an instance by instance basis, other algorithms may become better.
+exist various exact algorithms, which have been compared on a large suite of problem instances by
+McCreesh and Prosser \cite{McCreesh:2015}. These experiments indicated that the single best
+algorithm depends on the CPU time limit considered: for very small CPU time limits, VF2
+\cite{Cordella:2004} is the best choice, whereas the Glasgow algorithm \cite{McCreesh:2015} has
+better success rates for larger time limits.  They also showed that on an instance by instance
+basis, other algorithms are often better.
 
@@
 
-A \emph{graph} $G=(N,E)$ consists of a \emph{node set} $N$ and an \emph{edge set} $E \subseteq N \times N$, where an edge $(u,u')$ is a couple of nodes.
-The number of neighbors of a node $u$ is the degree of $u$, $d^\circ(u)=\#\{ (u,u')\in E\}$.
-In this paper, we implicitely consider non directed graphs, such that $(u,u')\in E\Leftrightarrow (u',u)\in E$.
-The extension to directed graphs is rather straightforward, and all algorithms compared in this paper can handle directed graphs as well.
+A \emph{graph} $G=(N,E)$ consists of a \emph{node set} $N$ and an \emph{edge set} $E \subseteq N
+\times N$, where an edge $(u,u')$ is a pair of nodes. The number of neighbors of a node $u$ is
+called the degree of $u$, $d^\circ(u)=\#\{ (u,u')\in E\}$. In this paper, we implicitly consider
+non-directed graphs, such that $(u,u')\in E\Leftrightarrow (u',u)\in E$. The extension to directed
+graphs is rather straightforward, and all algorithms compared in this paper can handle directed
+graphs as well.
 
-A \emph{subgraph isomorphism problem} between a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$ consists in deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
-More precisely, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a different target node to each pattern node, and that preserves pattern edges, i.e., $\forall (u,u') \in E_p, (f(u),f(u')) \in E_t$.
+Given a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$, the \emph{subgraph
+isomorphism problem} consists of deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
+More precisely, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a
+different target node to each pattern node, and that preserves pattern edges, i.e.\ $\forall (u,u')
+\in E_p, (f(u),f(u')) \in E_t$.
 
-Note that the subgraph is not necessarily induced so that two pattern nodes that are not linked by an edge may be matched to two target nodes which are linked by an edge.
-This problem is also called subgraph monomorphism or subgraph matching in the literature.
-%{\em Induced subgraph isomorphism} is a particular case of subgraph isomorphism such that the subgraph isomorphism function must also preserve target edges between matched nodes, i.e., 
-%\[\forall (u,u') \in N_p^2, (u,u')\in E_p \Leftrightarrow(f(u),f(u')) \in E_t\]
+Note that the subgraph is not necessarily induced, so that two pattern nodes that are not linked by
+an edge may be mapped to two target nodes which are linked by an edge. This problem is also called
+\emph{subgraph monomorphism} or \emph{subgraph matching} in the literature.
 
-In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of subgraph isomorphism problem. 
-We also define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\# E_t$, and $d_p$ and $d_t$ the maximal degrees of the graphs $G_p$ and $G_t$.
+In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of a
+subgraph isomorphism problem.  We also define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\#
+E_t$, and $d_p$ and $d_t$ to be the maximum degrees of the graphs $G_p$ and $G_t$.
 
@@
 
-Subgraph isomorphism problems may be solved by a systematic exploration of the search space composed of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one incrementally extends a partial matching by matching a non matched pattern node to a non matched target node until either some edges are not matched by the current matching (the search must backtrack to a previous choice point and go on with another extension) or all pattern nodes have been matched (a solution has been found).
-To reduce the search space, this exhaustive exploration is combined with filtering techniques that aim at removing candidate couples of non matched pattern-target nodes $(u,v)\in N_p\times N_t$.
-Different filterings may be considered; some are stronger than others (they remove more candidate couples), but also have higher time complexities.
+Subgraph isomorphism problems may be solved by a systematic exploration of the search space composed
+of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one
+incrementally extends a partial matching by matching a non-matched pattern node to a non-matched
+target node until either some edges are not matched by the current matching (so the search must
+backtrack to a previous choice point and go on with another extension), or all pattern nodes have
+been matched (a solution has been found). To reduce the search space, this exhaustive exploration is
+combined with filtering techniques that aim at removing candidate pairs of non-matched
+pattern-target nodes $(u,v)\in N_p\times N_t$. Different filtering techniques may be considered;
+some are stronger than others (they remove more candidate pairs), but also have higher time
+complexities.
 
@@
 
-Simple filterings basically propagate difference constraints (which ensure that the matching is injective) and edge constraints (which ensure that the matching preserves pattern edges): each time a pattern node $u\in N_p$ is matched with a target node $v\in N_t$, one removes every candidate couple $(u',v')\in N_p\times N_t$ such that either $v'=v$ (difference constraint) or $(u,u')$ is a pattern edge whereas $(v,v')$ is not a target edge (edge constraint).
-This simple filtering (called {\em Forward-Checking}) is very fast to achieve: in ${\cal O}(n_p)$ for difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints.
-It is used, for example, in McGregor's algorithm \cite{mcgregor79} and in VF2 \cite{Cordella:2004}.
+The simplest form of filtering is simply to propagate difference constraints (which ensure that the
+matching is injective) and edge constraints (which ensure that the matching preserves pattern
+edges): each time a pattern node $u\in N_p$ is matched with a target node $v\in N_t$, one removes
+every candidate pair $(u',v')\in N_p\times N_t$ such that either $v'=v$ (difference constraint) or
+$(u,u')$ is a pattern edge but $(v,v')$ is not a target edge (edge constraint). This simple
+filtering (called \emph{Forward-Checking}) is very fast to achieve: in ${\cal O}(n_p)$ for
+difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints. It is used, for
+example, in McGregor's algorithm \cite{mcgregor79} and in VF2 \cite{Cordella:2004}.
 
-R\'egin has introduced in \cite{regin} a stronger filtering for difference constraints, which ensures that all pattern nodes can be matched with different target nodes, all together.
-This filtering (called {\em AllDifferent Generalized Arc Consistency}) removes more candidate couples than when each difference contraint is propagated separately which Forward-Checking.
-However, it is also more time consuming as it is done in ${\cal O}(n_p^2\cdot n_t^2)$.
+R\'egin \cite{regin} introduced a stronger filtering for difference constraints, which ensures that
+all pattern nodes can be matched with different target nodes, all together. This filtering (called
+\emph{AllDifferent Generalized Arc Consistency}) removes more candidate pairs than when each
+difference constraint is propagated separately which Forward-Checking. However, it is also more time
+consuming as it is done in ${\cal O}(n_p^2\cdot n_t^2)$ time.
 
-Various filterings have been introduced for edge constraints.
-Ullman \cite{ullman} has introduced a filtering which basically ensures that for each pattern edge $(u,u')\in E_p$ and each candidate couple $(u,v)\in N_p\times N_t$, there exists a candidate couple $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a target edge.
-Candidate couples $(u,v)$ that do not satisfy this property are iteratively removed until a fixed point is reached.
-This filtering (called {\em Arc Consistency}) removes more candidate couples than Forward-Checking, but it is also more time consuming as it is done in ${\cal O}(e_p\cdot n_t^2)$ when using the algorithm AC4  \cite{MH86}.
+Various filtering techniques have been tried for edge constraints. Ullman \cite{ullman} introduced a
+filtering which ensures that for each pattern edge $(u,u')\in E_p$ and each candidate pair $(u,v)\in
+N_p\times N_t$, there exists a candidate couple $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a
+target edge. Candidate couples $(u,v)$ that do not satisfy this property are iteratively removed
+until a fixed point is reached. This filtering (called \emph{Arc Consistency}) removes more
+candidate pairs than Forward-Checking, but it is also more time consuming as it is done in ${\cal
+O}(e_p\cdot n_t^2)$ when using the algorithm AC4  \cite{MH86}.
 
-A stronger filtering may be obtained by propagating edge constraints in a more global way, as proposed in \cite{LV02}.
-The idea is to check for each candidate couple $(u,v)\in N_p\times N_t$ that the number of pattern nodes adjacent to $u$ is smaller than or equal to the number of target nodes that are both adjacent to $v$ and that may be matched with nodes adjacent to $u$.
-This is done in ${\cal O}(n_p^2\cdot n_t^2)$.
-This idea has been generalized in \cite{Solnon:2010} where, for each candidate couple $(u,v)\in N_p\times N_t$, a redundant Local AllDifferent (LAD) constraint ensures that each neighbour of $u$ may be matched with a different neighbour of $v$.
-This is done in ${\cal O}(n_p\cdot n_t\cdot d_p^2\cdot d_t^2)$.
+Stronger filtering may be obtained by propagating edge constraints in a more global way, as proposed
+by Larrosa and Valiente \cite{LV02}. The idea is to check for each candidate pair $(u,v)\in
+N_p\times N_t$ that the number of pattern nodes adjacent to $u$ is smaller than or equal to the
+number of target nodes that are both adjacent to $v$ and that may be matched with nodes adjacent to
+$u$. This is done in ${\cal O}(n_p^2\cdot n_t^2)$. This idea was generalised by Solnon
+\cite{Solnon:2010} where, for each candidate pair $(u,v)\in N_p\times N_t$, a redundant Local
+AllDifferent (LAD) constraint ensures that each neighbour of $u$ may be matched with a different
+neighbour of $v$. This is done in ${\cal O}(n_p\cdot n_t\cdot d_p^2\cdot d_t^2)$.
 
@@
 
-Some filterings exploit invariant properties, {\em i.e.}, properties associated with nodes such that nodes may be matched only if they have compatible properties.
-A classical property is the degree: a pattern node $u\in N_p$ may be matched with a target node $v\in N_t$ only if $d^\circ(u)\leq d^\circ(v)$.
-This property is usually used at the beginning of the search to reduce the set of candidate couples to $\{(u,v)\in N_p\times N_t\;|\;d^\circ(u)\leq d^\circ(v)\}$.
-Other examples of invariant properties are the number of cycles of length $k$ passing through the node, and the number of cliques of size $k$ containing the node, which must be smaller for a pattern node than for its matched target node.
-Invariant properties may also be associated with couples of nodes.
-For example, the number of paths of length $k$ between two pattern nodes is smaller than or equal to the number of paths of length $k$ between the target nodes they may be matched with.
+Some filtering techniques exploit invariant properties, i.e.\ properties associated with nodes such
+that nodes may be matched only if they have compatible properties. A classical property is the
+degree: a pattern node $u\in N_p$ may be matched with a target node $v\in N_t$ only if
+$d^\circ(u)\leq d^\circ(v)$. This property is usually used at the beginning of the search to reduce
+the set of candidate pairs to $\{(u,v)\in N_p\times N_t\;|\;d^\circ(u)\leq d^\circ(v)\}$.  Other
+examples of invariant properties are the number of cycles of length $k$ passing through the node,
+and the number of cliques of size $k$ containing the node, which must be smaller for a pattern node
+than for its matched target node.  Invariant properties may also be associated with pairs of nodes.
+For example, the number of paths of length $k$ between two pattern nodes is smaller than or equal to
+the number of paths of length $k$ between the target nodes with which they may be matched.
 
 These invariant properties are used, for example:
+
 \begin{itemize}
-\item  In \cite{battiti-mascia07}, to remove candidate couples $(u,v)\in N_p\times N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of paths starting from  target node $v$~;
-\item In \cite{Audemard:2014} to generalize the LAD constraint proposed in \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be matched with all different compatible target nodes, where compatibility is defined with respect to invariant properties~;
-\item In \cite{McCreesh:2015} to filter the set of candidate couples before starting the search.
+\item  By Battiti and Mascia \cite{battiti-mascia07}, to remove candidate pairs $(u,v)\in N_p\times
+    N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of
+    paths starting from  target node $v$;
+\item By Audemard et al.\ \cite{Audemard:2014} to generalise the LAD constraint proposed by Solnon
+    \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be matched with all
+    different compatible target nodes, where compatibility is defined with respect to invariant
+    properties;
+\item by McCreesh and Prosser \cite{McCreesh:2015} to filter the set of candidate couples before
+    starting the search, and to generate additional implied adjacency-like constraints which are
+    processed during search.
 \end{itemize}
-In \cite{Audemard:2014} , the length of paths is not limited, and it is iteratively incremented until no more couple is removed.
-In \cite{battiti-mascia07} and \cite{McCreesh:2015},  algorithms are parameterized by the maximum path length considered when counting paths: Larger values for this parameter remove more candidate couples, but are also more time consuming.
-Experiments reported in \cite{battiti-mascia07} show us that the best setting depends on the instance considered, and that a portfolio running several randomized
-versions in time-sharing decreases the total CPU-time needed to find a solution for feasible instances.
-In experiments reported in \cite{McCreesh:2015}, this parameter has been set to 3, as this setting seemed to be a reasonable compromise.
+
+\noindent Audemard et al.\ \cite{Audemard:2014} do not limit the length of paths considered, and
+iteratively increment the length until no more pairs are removed. Battiti and Mascia
+\cite{battiti-mascia07}, and McCreesh and Prosser \cite{McCreesh:2015} parameterise their algorithms
+by the maximum path length considered when counting paths: larger values for this parameter remove
+more candidate pairs, but are also more time consuming. Battiti and Mascia's experiments show that
+the best setting depends on the instance considered, and that a portfolio running several randomised
+versions in time-sharing decreases the total CPU time needed to find a solution for feasible
+instances. McCreesh and Prosser simply set the parameter to 3, as this setting seemed to be a
+reasonable compromise.
 
@@
 best on a given problem instance. Algorithm selection systems usually build machine learning models
-of the algorithms or the portfolio they are contained in to forecast which algorithm to use in a
+of the algorithms or the in which portfolio they are contained to forecast which algorithm to use in a
 particular context. Using the model predictions, one or more algorithms from the portfolio are
@@
 We consider 5725 instances, which are available in a simple text
-format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These
-instances are grouped into 12 classes.
+format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
+classes.
 
@@
 \item Class 1 contains randomly generated scale-free graphs \cite{constraints10}.
-\item Classes 2 and 3 contain various kinds of graphs coming from \cite{LV02}: class 2 contains small
-    instances generated from the 50 first graphs, whereas class 3 contains larger
-    instances where pattern graphs are coming from class 2, and target graphs are coming from the
-    next 50 graphs.
-\item Classes 4 to 8 contain randomly generated graphs coming from
+\item Classes 2 and 3 contain various kinds of graphs selected by Larrosa and Valiente \cite{LV02}:
+    class 2 contains small instances generated from the 50 first graphs, and class 3 contains
+    larger instances where pattern graphs are coming from class 2, and target graphs are coming from
+    the next 50 graphs.
+\item Classes 4 to 8 contain randomly generated graphs coming from a database of graphs commonly
+    used for benchmarking subgraph isomorphism algorithms
     \cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
-    for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
+    for classes 5 and 6, and random graphs with uniform edge probabilities for class 7. All of these
+    instances are satisfiable.
 \item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
 \item Class 11 contains instances coming from meshes modeling 3D objects \cite{cviu11}.
-\item Class 12 contains random graph instances belonging to the phase transition.
+\item Class 12 contains random graph instances chosen to be close to the satisfiable-unsatisfiable
+    phase transition---these instances are expected to be particularly challenging, despite their
+    small size.
 \end{itemize}
 
-Note that Classes 3 and 12 were not considered in the previous experimental
-study reported in \cite{McCreesh:2015}.  This set of instances is also much
-larger than that of Battiti and Mascia~\cite{battiti-mascia07}, who were the
-first to propose algorithm portfolios for subgraph isomorphism problems.
-Battiti and Mascia only considered a pure parallel portfolio consisting of two
-solvers, and without a selection mechanism. Their problem set consisted
-entirely of easy satisfiable instances, which is where this style of approach
-has most success.
+Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
+Prosser \cite{McCreesh:2015}.  This set of instances is also much larger than that of Battiti and
+Mascia~\cite{battiti-mascia07}, who were the first to propose algorithm portfolios for subgraph
+isomorphism problems.  Battiti and Mascia only considered a pure parallel portfolio consisting of
+two solvers, and without a selection mechanism. Their problem set consisted entirely of easy
+satisfiable instances, which is where this style of approach has most success.
 
@@
 \begin{itemize}
-\item VF2 \cite{Cordella:2004}, which performs a weak filtering so that it is especially fast on trivially satisfiable instances;
-\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques (AllDifferent Generalized Arc Consistency and LAD);
+\item VF2 \cite{Cordella:2004}, which performs a weak filtering that it is especially fast on
+    trivially satisfiable instances;
+\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques
+    (AllDifferent Generalized Arc Consistency and LAD);
 \item Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
@@
 \noindent The Glasgow algorithm has a parameter, which controls the lengths of paths used when
-reasoning about non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice
-of paths of length 3 was used as a reasonable compromise---longer paths lead to prohibitively
-expensive preprocessing on larger, denser instances. This is often not the best choice on an
-instance by instance basis: sometimes path-based reasoning gives no benefit at all, sometimes
-considering only paths of length 2 suffices, occasionally paths of length 4 are helpful, and even
-looking at paths of length 3 is relatively expensive on some graphs.
+reasoning about non-adjacent vertices.  In experiments reported by McCreesh and Prosser
+\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer
+paths lead to prohibitively expensive preprocessing on larger, denser instances. This is often not
+the best choice on an instance by instance basis: sometimes path-based reasoning gives no benefit at
+all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are
+helpful, and even looking at paths of length 3 is relatively expensive on some graphs.
 
-We also introduce two new variants of LAD.
-The first variant is a weaker LAD filtering (called incompleteLAD) which is applied once, without performing a backtracking search, and very quickly detects inconsistencies on many instances: For each pattern node $u$, we check that there exists at least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
-IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when it does not detect inconsistency, the instance may either be consistent or not.
-As a counterpart of this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.
+We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called
+incompleteLAD) which is applied once, without performing a backtracking search, and very quickly
+detects inconsistencies on many instances: for each pattern node $u$, we check that there exists at
+least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor
+$v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
+IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when
+it does not detect inconsistency, the instance may either be satisfiable or not. As a counterpart of
+this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.
 
-The second variant of LAD is called pathLAD and it combines the LAD constraints introduced in \cite{Solnon:2010} with the exploitation of path length properties, as proposed in \cite{Audemard:2014}.
-The idea is to label each edge $(u,v)$ with the number of paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$, and to add the constraint that the label of a pattern node (resp.
-edge) must be smaller than or equal to the label of its associated target node (resp.
-edge).
+The second variant of LAD is called pathLAD, which combines the LAD constraints introduced by Solnon
+\cite{Solnon:2010} with the exploitation of path length properties proposed by Audemard et
+al.\ \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
+between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
+and to add the constraint that the label of a pattern node (resp.\ edge) must be smaller than or
+equal to the label of its associated target node (resp.\ edge).
 
@@
 \includegraphics[width=\textwidth]{courbe.pdf}
-\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms of the portfolio, and a Virtual Best Solver (VBS).\label{expTimeGraph}}
+\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms
+of the portfolio, and a Virtual Best Solver (VBS).\label{expTimeGraph}}
 \end{figure}
 
-Figure \ref{expTimeGraph} displays the evolution of the number of instances solved with respect to CPU time. It shows us that the best solver depends on the time limit
-considered.
-IncompleteLAD is able to solve easy inconsistent instances very quickly, in a few milliseconds.
-Hence, for time limits lower than 5 ms, the best solver is incompleteLAD.
-However, it is not able to solve harder inconsistent instances, neither can it solve consistent instances.
+Figure \ref{expTimeGraph} displays the evolution of the number of instances solved with respect to
+CPU time. It shows us that the best solver depends on the time limit considered. IncompleteLAD is
+able to solve easy unsastisfiable instances very quickly, in a few milliseconds. Hence, for time
+limits lower than \SI{5}{\ms}, the best solver is incompleteLAD. However, it is not able to solve harder
+unsatisfiable instances, nor can it solve satsifiable instances.
 
-PathLAD and Glasgow(1) outperform incompleteLAD when considering longer time limits: PathLAD is the best solver for time limits larger than 5ms and lower than 40 ms, and Glasgow(1) is the best solver for time limits larger than 40ms and lower than 3000 ms.
+PathLAD and Glasgow(1) outperform incompleteLAD when considering longer time limits: PathLAD is the
+best solver for time limits larger than \SI{5}{\ms} and lower than \SI{40}{\ms}, and Glasgow(1) is the best solver
+for time limits larger than \SI{40}{\ms} and lower than \SI{3000}{\ms}.
 
-Finally, for CPU time limits larger than 3000 ms, the best solver becomes Glasgow(2).
-As we increase the CPU time limit, variants of Glasgow with longer paths (Glasgow(3) and Glasgow(4)) become better.
-This is what we expect, as more reasoning is expensive, but potentially increases the reduction of the search space.
-In particular, the relative performance of Glasgow variants is monotone in the time it takes to solve the instance – Glasgow(1) is best at first, but gets worse as the time to solve the instance increases until Glasgow(2) is better.
-After that point, Glasgow(1) is never better than Glasgow(2) again.
-The same is true for path lengths 3 and 4.
+For CPU time limits larger than \SI{3000}{\ms}, the best solver becomes Glasgow(2).  As we
+increase the CPU time limit, variants of Glasgow with longer paths (Glasgow(3) and Glasgow(4))
+become better. This is what we expect, as more reasoning is expensive, but potentially increases the
+reduction of the search space. In particular, the relative performance of Glasgow variants is
+monotone in the time it takes to solve the instance---Glasgow(1) is best at first, but gets worse as
+the time to solve the instance increases until Glasgow(2) is better. After that point, Glasgow(1) is
+never better than Glasgow(2) again. The same is true for path lengths 3 and 4.
 
-The figure illustrates the potential for portfolios and algorithm selection we have. There is clearly no single solver that dominates throughout.
+The figure illustrates the potential for portfolios and algorithm selection we have. There is
+clearly no single solver that dominates throughout.
 
@@
 performance. The difference between VBS and single best is more particularly important for rather
-small CPU time limits (smaller than 1000 ms). In many applications, it is important to have the
-fastest possible algorithm. For example, in pattern recognition applications such as those reported
-in \cite{pr15,cviu11}, and in chemical applications such as \cite{Giugno:2013}, we often have to
-solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern
-image or molecule in a large database of target images or compounds, for example), so that having an
-algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference.
-Therefore, it is important to select the best algorithm for each instance, even if the instance is
-an easy one.
+small CPU time limits (smaller than \SI{1000}{\ms}). In many applications, it is important to have the
+fastest possible algorithm. For example, in pattern recognition \cite{pr15,cviu11} and chemical
+\cite{Giugno:2013} applications, we often have to solve the subgraph isomorphism problem repeatedly
+for a very large number of graphs (in order to find a pattern image or molecule in a large database
+of target images or compounds, for example), so having an algorithm that is able to solve an
+instance in \SI{100}{\ms} instead of \SI{1000}{\ms} makes a big difference.  Therefore, it is important to select
+the best algorithm for each instance, even if the instance is an easy one.
 
-Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class instances are coming from, unlike SAT.
-For all classes, there are always at least 2 algorithms which are the best for at least one instance of the class.
-In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for class 3).
+Table~\ref{expClass} shows us that we cannot simply select algorithms based upon which class from
+which instances are coming (unlike in SAT, where this kind of approach can be reasonable). For all
+classes, there are always at least two algorithms which are the best for at least one instance of
+the class. In particular, for classes 2 and 3, each algorithm is the best for at least one instance
+(except Glasgow(4) for class 3).
 
@@
 
-Our approach is composed of 3 steps described below. First, we run two presolvers in a static way to
-quickly solve easy instances; second, we extract features from instances which are not solved by the
-first step; and finally, we select an algorithm and run it.
+Our approach is composed of three steps. First, we run two presolvers in a static way to quickly
+solve easy instances; second, we extract features from instances which are not solved by the first
+step; and third, we select an algorithm and run it.
 
@@
 
-Experimental results reported in Section \ref{expComp} have shown us that incompleteLAD is very fast
-(7 ms on average) and is able to very quickly solve 1919 instances. Therefore, we first run
+Experimental results reported in Section~\ref{expComp} have shown us that incompleteLAD is very fast
+(\SI{7}{\ms} on average) and is able to solve 1919 instances very quickly. Therefore, we first run
 incompleteLAD: if the instance is detected to be unsatisfiable, then we do not process it further.
@@
 solved by incompleteLAD, 1470 are solved by VF2 in less than 50 milliseconds. Therefore, after
-running incompleteLAD, we run the VF2 solver for 50 ms. This solves easy instances without the
-overhead of running algorithm selection and avoids potentially making incorrect solver choices.
-After running incompleteLAD and VF2 for 50 ms, we are left with 2336 hard instances that we consider
-for algorithm selection.
+running incompleteLAD, we run the VF2 solver for \SI{50}{\ms}. This solves easy instances without
+the overhead of running algorithm selection and avoids potentially making incorrect solver choices.
+After running incompleteLAD and VF2 for \SI{50}{\ms}, we are left with 2336 hard instances that we
+consider for algorithm selection.
 
@@
 
-To decide which solver to run on the remaining instances, we use
-LLAMA~\cite{kotthoff_llama_2013}. LLAMA supports the most common algorithm
-selection approaches used in the literature. We performed a set of preliminary
-experiments to determine the approach that works best in this case.
+To decide which solver to run on the remaining instances, we use LLAMA~\cite{kotthoff_llama_2013}.
+LLAMA supports the most common algorithm selection approaches used in the literature. We performed a
+set of preliminary experiments to determine the approach that works best in this case.
 
-We use 10-fold cross-validation to determine the performance of the LLAMA
-models. The entire set of instances was randomly partitioned into 10 subsets of
-approximately equal size. Of the 10 subsets, 9 were combined to form the
-training set for the algorithm selection models, which were evaluated on the
-remaining subset. This process was repeated 10 times for all possible
-combinations of training and test sets. At the end of this process, each problem
-instance in the original set was used exactly once to evaluate the performance
-of the algorithm selection models.
+We use 10-fold cross-validation to determine the performance of the LLAMA models. The entire set of
+instances was randomly partitioned into 10 subsets of approximately equal size. Of the 10 subsets, 9
+were combined to form the training set for the algorithm selection models, which were evaluated on
+the remaining subset. This process was repeated 10 times for all possible combinations of training
+and test sets. At the end of this process, each problem instance in the original set was used
+exactly once to evaluate the performance of the algorithm selection models.
 
-LLAMA's pairwise regression approach with random forest regression gave the best
-performance. The idea is very similar to the pairwise classification used
-in~\cite{xu_satzilla_2008}. For each pair of algorithms in our portfolio, we
-train a model that predicts the performance difference between them. That is, if
-the first algorithm is better than the second, the difference is positive,
-otherwise negative. The algorithm with the highest cumulative performance
-difference, i.e.\ the most positive difference over all other algorithms, is
-chosen to be run.
+LLAMA's pairwise regression approach with random forest regression gave the best performance. The
+idea is very similar to the pairwise classification used by Xu et al.\ \cite{xu_satzilla_2008}. For
+each pair of algorithms in our portfolio, we train a model that predicts the performance difference
+between them. That is, if the first algorithm is better than the second, the difference is positive,
+otherwise negative. The algorithm with the highest cumulative performance difference, i.e.\ the most
+positive difference over all other algorithms, is chosen to be run.
 
-As this approach gives very good performance already, we did not tune the
-parameters of the random forest machine learning algorithm. It is possible that
-overall performance can be improved by doing so and we make no claims that the
-particular algorithm selection approach we use in this paper cannot be improved.
+As this approach gives very good performance already, we did not tune the parameters of the random
+forest machine learning algorithm. It is possible that overall performance can be improved by doing
+so and we make no claims that the particular algorithm selection approach we use in this paper
+cannot be improved.
 
-The data we use in this paper is available as ASlib~\cite{aslib} scenario
-GRAPHS-2015.
+The data we use in this paper is available as ASlib~\cite{aslib} scenario GRAPHS-2015.
 
@@
 achieve. The single best solver is the one solver from the portfolio that has the overall best
-performance across the entire set of instances, at the CPU time limit of $10^8$ ms, {\em i.e.}, Glasgow(2). We consider it a lower bound on performance.
+performance across the entire set of instances, at the CPU time limit of \SI{10e8}{\ms},
+i.e.\ Glasgow(2). We consider it a lower bound on performance.
+We are able to solve 34 more instances than the single best solver within the timeout, with only an
+additional 13 to the virtual best. In terms of average performance, we are able to close almost
+\SI{70}{\percent} of the gap between the single best and the virtual best solver.
 
 \begin{table}[ht]
-\centering
+    \centering\setlength{\tabcolsep}{1em}
 \begin{tabular}{lrrr}
@@
 LLAMA & 610835 & 2206 & 6435301\\
-   \bottomrule
+   \bottomrule \\
 \end{tabular}
-\vspace{1ex}
 \caption{Algorithm selection performance on the reduced set of 2336
@@
 
-We are able to solve 34 more instances than the single best solver within the
-timeout, with only an additional 13 to the virtual best. In terms of average
-performance, we are able to close almost 70\% of the gap between the single best
-and the virtual best solver.
-
-Table~\ref{tab:res} shows the performance of the selection model. The
-performance of the entire system we propose here is shown in
-Table~\ref{tab:resfull}. The additional instances are solved either by
-incompleteLAD or VF2 presolving, as explained above. Our system is able to close
-more than 50\% of the gap between single and virtual best.
+Table~\ref{tab:res} shows the performance of the selection model. The performance of the entire
+system we propose here is shown in Table~\ref{tab:resfull}. The additional instances are solved
+either by incompleteLAD or VF2 presolving, as explained above. Our system is able to close more than
+\SI{50}{\percent} of the gap between single and virtual best.
 
 \begin{table}[ht]
-\centering
+    \centering\setlength{\tabcolsep}{1em}
 \begin{tabular}{lrrr}
@@
 LLAMA & 249242 & 5595 & 2625831\\
-   \bottomrule
+   \bottomrule \\
 \end{tabular}
-\vspace{1ex}
 \caption{Algorithm selection system performance on the full set of 5725
@@
 misclassification penalty does not take the cost of computing features into
-account, while the mean performance does -- for these small timeouts, the cost
+account, while the mean performance does---for these small timeouts, the cost
 of computing the features dominates any improvement through algorithm selection.
@@
 
-Table \ref{expTimeTable} displays the number of instances solved at different CPU time limits, ranging
-from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
-considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very
-quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
-preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2
+Table~\ref{expTimeTable} displays the number of instances solved at different CPU time limits,
+ranging from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time
+limit considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances
+very quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
+preprocessing is an incomplete approach which can only detect rather trivial inconsistencies; VF2
 performs a basic backtracking search and is very fast on easy instances, because it does not compute
 expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
-and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow
-1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS),
-which considers the best algorithm for each instance separately, obtains much better results,
-showing us that the algorithms have complementary performance. The difference between VBS and single
-best (column $\delta$ of Table \ref{expTimeTable}) is more particularly important for rather small CPU
-time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the
-limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For
-example, in pattern recognition applications, we often have to solve subgraph isomorphism problems
-for a very large number of graphs (in order to find a pattern image in a large database of target
-images, for example), so that having an algorithm that is able to solve an instance in 100 ms
-instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm
-for each instance.
+and LAD preprocessing when considering longer time limits: \SI{10}{\ms} for default and new LAD and
+Glasgow 1, and \SI{100}{\ms} (resp. \SI{1000}{\ms} and \SI{10000}{\ms}) for Glasgow(2) (resp.\ 3 and
+4). The Virtual Best Solver (VBS), which considers the best algorithm for each instance separately,
+obtains much better results, showing us that the algorithms have complementary performance. The
+difference between VBS and single best (column $\delta$ of Table \ref{expTimeTable}) is more
+particularly important for rather small CPU time limits. In particular, VBS solves 693 more
+instances than the single best (Glasgow(1)) when the limit is \SI{100}{\ms}. In many applications,
+it is important to have the fastest possible algorithm. For example, in pattern recognition
+applications, we often have to solve subgraph isomorphism problems for a very large number of graphs
+(in order to find a pattern image in a large database of target images, for example), so that having
+an algorithm that is able to solve an instance in \SI{100}{\ms} instead of \SI{1000}{\ms} makes a
+big difference. Therefore, it is important to select the best algorithm for each instance.
 
-Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the
-upper bound of what an algorithm selection approach can achieve), and the single best solver, at
-different CPU time limits (the single best corresponds to results highlighted in blue in Table 1).
-Comment Table 3: Mean runtime but also number of solved instances at different CPU time limits.
-Logically, LLAMA may solve less instances than single best for small time limits because it spends
-time to compute features and choose a solver, but it should be better for larger time limits...
+Table~\ref{selectionvsvbs} shows the performance of our algorithm selection approach, compared to
+the VBS (which is the upper bound of what an algorithm selection approach can achieve), and the
+single best solver, at different CPU time limits (the single best corresponds to results highlighted
+in blue in Table 1).  Comment Table 3: Mean runtime but also number of solved instances at different
+CPU time limits.  Logically, LLAMA may solve less instances than single best for small time limits
+because it spends time to compute features and choose a solver, but it should be better for larger
+time limits...
 
@@
 \end{tabular}
-\caption{}
+\caption{}\label{selectionvsvbs}
 \end{table}
 
-Add Figure 1 which compares single best at $10^8$ ms (Glasgow 2) with LLAMA ?
+Add Figure 1 which compares single best at \SI{10e8}{\ms}(Glasgow(2)) with LLAMA ?
 

### 2015-12-14 paper/paper.tex
Typeset algorithm names

@@
 \usepackage{siunitx}
+\usepackage{xspace}
+\usepackage{scalefnt}
 
@@
 \Crefname{proposition}{Proposition}{Propositions}
+
+\newcommand{\VFtwo}{$\textsc{Vf{\scalefont{0.8}2}}$\xspace}
+\newcommand{\Glasgow}{$\textsc{Glasgow}$\xspace}
+\newcommand{\LAD}{$\textsc{Lad}$\xspace}
+\newcommand{\IncompleteLAD}{$\textsc{IncompleteLad}$\xspace}
+\newcommand{\PathLAD}{$\textsc{PathLad}$\xspace}
+\newcommand{\GlasgowOne}{$\textsc{Glasgow{\scalefont{0.8}1}}$\xspace}
+\newcommand{\GlasgowTwo}{$\textsc{Glasgow{\scalefont{0.8}2}}$\xspace}
+\newcommand{\GlasgowThree}{$\textsc{Glasgow{\scalefont{0.8}3}}$\xspace}
+\newcommand{\GlasgowFour}{$\textsc{Glasgow{\scalefont{0.8}4}}$\xspace}
+\newcommand{\LLAMA}{$\textsc{Llama}$\xspace}
 
@@
 McCreesh and Prosser \cite{McCreesh:2015}. These experiments indicated that the single best
-algorithm depends on the CPU time limit considered: for very small CPU time limits, VF2
-\cite{Cordella:2004} is the best choice, whereas the Glasgow algorithm \cite{McCreesh:2015} has
+algorithm depends on the CPU time limit considered: for very small CPU time limits, \VFtwo
+\cite{Cordella:2004} is the best choice, whereas the \Glasgow algorithm \cite{McCreesh:2015} has
 better success rates for larger time limits.  They also showed that on an instance by instance
@@
 difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints. It is used, for
-example, in McGregor's algorithm \cite{mcgregor79} and in VF2 \cite{Cordella:2004}.
+example, in McGregor's algorithm \cite{mcgregor79} and in \VFtwo \cite{Cordella:2004}.
 
@@
 all pattern nodes can be matched with different target nodes, all together. This filtering (called
-\emph{AllDifferent Generalized Arc Consistency}) removes more candidate pairs than when each
+\emph{All-Different Generalized Arc Consistency}) removes more candidate pairs than when each
 difference constraint is propagated separately which Forward-Checking. However, it is also more time
@@
 $u$. This is done in ${\cal O}(n_p^2\cdot n_t^2)$. This idea was generalised by Solnon
-\cite{Solnon:2010} where, for each candidate pair $(u,v)\in N_p\times N_t$, a redundant Local
-AllDifferent (LAD) constraint ensures that each neighbour of $u$ may be matched with a different
+\cite{Solnon:2010}'s \LAD algorithm, where, for each candidate pair $(u,v)\in N_p\times N_t$, a redundant Local
+All-Different constraint ensures that each neighbour of $u$ may be matched with a different
 neighbour of $v$. This is done in ${\cal O}(n_p\cdot n_t\cdot d_p^2\cdot d_t^2)$.
@@
     paths starting from  target node $v$;
-\item By Audemard et al.\ \cite{Audemard:2014} to generalise the LAD constraint proposed by Solnon
-    \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be matched with all
-    different compatible target nodes, where compatibility is defined with respect to invariant
-    properties;
+\item By Audemard et al.\ \cite{Audemard:2014} to generalise the locally all-different constraint
+    proposed by Solnon \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be
+    matched with all different compatible target nodes, where compatibility is defined with respect
+    to invariant properties;
 \item by McCreesh and Prosser \cite{McCreesh:2015} to filter the set of candidate couples before
@@
 We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
-Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
-the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
+Scientific Linux 6.5. We used the C++ implementation of the \Glasgow algorithm \cite{McCreesh:2015},
+the C implementation of \LAD \cite{Solnon:2010}, and the VFLib C implementation of \VFtwo
 \cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
@@
 \begin{itemize}
-\item VF2 \cite{Cordella:2004}, which performs a weak filtering that it is especially fast on
+\item \VFtwo \cite{Cordella:2004}, which performs a weak filtering that it is especially fast on
     trivially satisfiable instances;
-\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques
-    (AllDifferent Generalized Arc Consistency and LAD);
-\item Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
+\item \LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques
+    (All-Different Generalized Arc Consistency and Locally All-Different);
+\item \Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
     invariant properties to generate additional constraints, followed by weaker filtering
-    (forward-checking, and a heuristic AllDifferent propagator which can miss deletions) and
+    (forward-checking, and a heuristic All-Different propagator which can miss deletions) and
     conflict-directed backjumping during search.
@@
 
-\noindent The Glasgow algorithm has a parameter, which controls the lengths of paths used when
+\noindent The \Glasgow algorithm has a parameter, which controls the lengths of paths used when
 reasoning about non-adjacent vertices.  In experiments reported by McCreesh and Prosser
@@
 all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are
-helpful, and even looking at paths of length 3 is relatively expensive on some graphs.
+helpful, and even looking at paths of length 3 is relatively expensive on some graphs. We thus all
+four choices, naming these variants \GlasgowOne through \GlasgowFour.
 
-We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called
-incompleteLAD) which is applied once, without performing a backtracking search, and very quickly
+We also introduce two new variants of \LAD. The first, called \IncompleteLAD, is a weaker
+filtering which is applied once, without performing a backtracking search, and very quickly
 detects inconsistencies on many instances: for each pattern node $u$, we check that there exists at
@@
 $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
-IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when
+\IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when
 it does not detect inconsistency, the instance may either be satisfiable or not. As a counterpart of
@@
 
-The second variant of LAD is called pathLAD, which combines the LAD constraints introduced by Solnon
-\cite{Solnon:2010} with the exploitation of path length properties proposed by Audemard et
-al.\ \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
-between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
-and to add the constraint that the label of a pattern node (resp.\ edge) must be smaller than or
-equal to the label of its associated target node (resp.\ edge).
+The second variant of \LAD is called \PathLAD, which combines the locally all-different constraints
+introduced by Solnon \cite{Solnon:2010} with the exploitation of path length properties proposed by
+Audemard et al.\ \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of
+paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3
+passing through $u$, and to add the constraint that the label of a pattern node (resp.\ edge) must
+be smaller than or equal to the label of its associated target node (resp.\ edge).
 
@@
 Figure \ref{expTimeGraph} displays the evolution of the number of instances solved with respect to
-CPU time. It shows us that the best solver depends on the time limit considered. IncompleteLAD is
-able to solve easy unsastisfiable instances very quickly, in a few milliseconds. Hence, for time
-limits lower than \SI{5}{\ms}, the best solver is incompleteLAD. However, it is not able to solve harder
-unsatisfiable instances, nor can it solve satsifiable instances.
+CPU time. It shows us that the best solver depends on the time limit considered. \IncompleteLAD is
+able to solve easy unsatisfiable instances very quickly, in a few milliseconds. Hence, for time
+limits lower than \SI{5}{\ms}, the best solver is \IncompleteLAD. However, it is not able to solve harder
+unsatisfiable instances, nor can it solve satisfiable instances.
 
-PathLAD and Glasgow(1) outperform incompleteLAD when considering longer time limits: PathLAD is the
-best solver for time limits larger than \SI{5}{\ms} and lower than \SI{40}{\ms}, and Glasgow(1) is the best solver
+\PathLAD and \GlasgowOne outperform \IncompleteLAD when considering longer time limits: \PathLAD is the
+best solver for time limits larger than \SI{5}{\ms} and lower than \SI{40}{\ms}, and \GlasgowOne is the best solver
 for time limits larger than \SI{40}{\ms} and lower than \SI{3000}{\ms}.
 
-For CPU time limits larger than \SI{3000}{\ms}, the best solver becomes Glasgow(2).  As we
-increase the CPU time limit, variants of Glasgow with longer paths (Glasgow(3) and Glasgow(4))
+For CPU time limits larger than \SI{3000}{\ms}, the best solver becomes \GlasgowTwo.  As we
+increase the CPU time limit, variants of \Glasgow with longer paths (\GlasgowThree and \GlasgowFour)
 become better. This is what we expect, as more reasoning is expensive, but potentially increases the
 reduction of the search space. In particular, the relative performance of Glasgow variants is
-monotone in the time it takes to solve the instance---Glasgow(1) is best at first, but gets worse as
-the time to solve the instance increases until Glasgow(2) is better. After that point, Glasgow(1) is
-never better than Glasgow(2) again. The same is true for path lengths 3 and 4.
+monotone in the time it takes to solve the instance---\GlasgowOne is best at first, but gets worse as
+the time to solve the instance increases until \GlasgowTwo is better. After that point, \GlasgowOne is
+never better than \GlasgowTwo again. The same is true for path lengths 3 and 4.
 
@@
 the class. In particular, for classes 2 and 3, each algorithm is the best for at least one instance
-(except Glasgow(4) for class 3).
+(except \GlasgowFour for class 3).
 
@@
 \toprule
-Class && VF2 && \multicolumn{3}{c}{LAD} && \multicolumn{4}{c}{\hspace*{2em}Glasgow}\\
+Class && \VFtwo && \multicolumn{3}{c}{\LAD} && \multicolumn{4}{c}{\hspace*{2em}\Glasgow}\\
 &&&&incomplete&default&path&&1&2&3&4\\
@@
 
-Experimental results reported in Section~\ref{expComp} have shown us that incompleteLAD is very fast
+Experimental results reported in Section~\ref{expComp} have shown us that \IncompleteLAD is very fast
 (\SI{7}{\ms} on average) and is able to solve 1919 instances very quickly. Therefore, we first run
-incompleteLAD: if the instance is detected to be unsatisfiable, then we do not process it further.
-Otherwise, we record as features the number of candidate couples removed by incompleteLAD in absolute
+\IncompleteLAD: if the instance is detected to be unsatisfiable, then we do not process it further.
+Otherwise, we record as features the number of candidate couples removed by \IncompleteLAD in absolute
 terms, as a percentage, and the minimum and maximum on a per-variable basis.
 
-VF2 is also able to solve many easy instances very quickly: among the 3806 instances which are not
-solved by incompleteLAD, 1470 are solved by VF2 in less than 50 milliseconds. Therefore, after
-running incompleteLAD, we run the VF2 solver for \SI{50}{\ms}. This solves easy instances without
+\VFtwo is also able to solve many easy instances very quickly: among the 3806 instances which are not
+solved by \IncompleteLAD, 1470 are solved by \VFtwo in less than 50 milliseconds. Therefore, after
+running \IncompleteLAD, we run the \VFtwo solver for \SI{50}{\ms}. This solves easy instances without
 the overhead of running algorithm selection and avoids potentially making incorrect solver choices.
-After running incompleteLAD and VF2 for \SI{50}{\ms}, we are left with 2336 hard instances that we
+After running \IncompleteLAD and \VFtwo for \SI{50}{\ms}, we are left with 2336 hard instances that we
 consider for algorithm selection.
@@
     \item The mean and maximum degrees, and whether or not every vertex has the same degree. (The
-        degree-based invariants used by LAD and Glasgow do nothing at the top of search if every
+        degree-based invariants used by \LAD and \Glasgow do nothing at the top of search if every
         vertex has the same degree.)
@@
 
-To decide which solver to run on the remaining instances, we use LLAMA~\cite{kotthoff_llama_2013}.
-LLAMA supports the most common algorithm selection approaches used in the literature. We performed a
+To decide which solver to run on the remaining instances, we use \LLAMA~\cite{kotthoff_llama_2013}.
+\LLAMA supports the most common algorithm selection approaches used in the literature. We performed a
 set of preliminary experiments to determine the approach that works best in this case.
 
-We use 10-fold cross-validation to determine the performance of the LLAMA models. The entire set of
+We use 10-fold cross-validation to determine the performance of the \LLAMA models. The entire set of
 instances was randomly partitioned into 10 subsets of approximately equal size. Of the 10 subsets, 9
@@
 
-LLAMA's pairwise regression approach with random forest regression gave the best performance. The
+\LLAMA's pairwise regression approach with random forest regression gave the best performance. The
 idea is very similar to the pairwise classification used by Xu et al.\ \cite{xu_satzilla_2008}. For
@@
 performance across the entire set of instances, at the CPU time limit of \SI{10e8}{\ms},
-i.e.\ Glasgow(2). We consider it a lower bound on performance.
+i.e.\ GlasgowTwo. We consider it a lower bound on performance.
 We are able to solve 34 more instances than the single best solver within the timeout, with only an
@@
 single best & 1965158 & 2172 & 7787967\\
-LLAMA & 610835 & 2206 & 6435301\\
+\LLAMA & 610835 & 2206 & 6435301\\
    \bottomrule \\
@@
 system we propose here is shown in Table~\ref{tab:resfull}. The additional instances are solved
-either by incompleteLAD or VF2 presolving, as explained above. Our system is able to close more than
+either by \IncompleteLAD or \VFtwo presolving, as explained above. Our system is able to close more than
 \SI{50}{\percent} of the gap between single and virtual best.
@@
 single best & 798660 & 5562 & 3174573\\
-LLAMA & 249242 & 5595 & 2625831\\
+\LLAMA & 249242 & 5595 & 2625831\\
    \bottomrule \\
@@
 Figure~\ref{fig:portfolio-ecdf} shows the cumulative distribution function for
-the individual solvers, the virtual best solver, and the LLAMA portfolio. The
+the individual solvers, the virtual best solver, and the \LLAMA portfolio. The
 virtual best solver clearly dominates with a significant margin to the next
-best. VF2 is far below all other solvers. The portfolio does not perform well
+best. \VFtwo is far below all other solvers. The portfolio does not perform well
 for instances that can be solved quickly because of the overhead incurred
@@
 \caption{Empirical cumulative distribution function of solvers, virtual best
-solvers, and the LLAMA portfolio on the reduced set of 2336 instances. The
+solvers, and the \LLAMA portfolio on the reduced set of 2336 instances. The
 actual solvers are shown as dashed lines.}
@@
 Figure~\ref{fig:portfolio-ecdf-full} shows the ECDF of the system including
-incompleteLAD and VF2 presolving on the full set of instances. Note that for
+\IncompleteLAD and \VFtwo presolving on the full set of instances. Note that for
 very easy instances, we are better than the virtual best solver because
-incompleteLAD is not included in our portfolio. The performance on small
-instances is much better than the LLAMA selector alone (cf.\
+\IncompleteLAD is not included in our portfolio. The performance on small
+instances is much better than the \LLAMA selector alone (cf.\
 Figure~\ref{fig:portfolio-ecdf}) and the region where the portfolio performs
@@
 performance of our portfolio for timeouts ranging from $10^2$ to $10^8$. We
-do not include incompleteLAD and VF2 presolving here, as these components do not
+do not include \IncompleteLAD and \VFtwo presolving here, as these components do not
 change for different timeouts.
@@
 $10^2$ & virtual best & 0 & 1253 & 59\\
- & LLAMA & 3 & 685 & 78\\
- & glasgow1 & 11 & 903 & 71\\
+ & \LLAMA & 3 & 685 & 78\\
+ & \Glasgow1 & 11 & 903 & 71\\
   \midrule
 $10^3$ & virtual best & 0 & 1781 & 322\\
- & LLAMA & 22 & 1284 & 569\\
- & glasgow1 & 126 & 1539 & 449\\
+ & \LLAMA & 22 & 1284 & 569\\
+ & \Glasgow1 & 126 & 1539 & 449\\
   \midrule
 $10^4$ & virtual best & 0 & 1943 & 2018\\
- & LLAMA & 186 & 1771 & 3342\\
- & glasgow2 & 683 & 1855 & 2701\\
+ & \LLAMA & 186 & 1771 & 3342\\
+ & \Glasgow2 & 683 & 1855 & 2701\\
   \midrule
 $10^5$ & virtual best & 0 & 2045 & 14348\\
- & LLAMA & 1346 & 2022 & 17231\\
- & glasgow2 & 3019 & 1988 & 17368\\
+ & \LLAMA & 1346 & 2022 & 17231\\
+ & \Glasgow2 & 3019 & 1988 & 17368\\
   \midrule
 $10^6$ & virtual best & 0 & 2111 & 108060\\
- & LLAMA & 10826 & 2087 & 120459\\
- & glasgow2 & 22729 & 2063 & 130789\\
+ & \LLAMA & 10826 & 2087 & 120459\\
+ & Glasgow2 & 22729 & 2063 & 130789\\
   \midrule
 $10^7$ & virtual best & 0 & 2178 & 815577\\
- & LLAMA & 91012 & 2156 & 908200\\
- & glasgow2 & 198944 & 2129 & 1014522\\
+ & \LLAMA & 91012 & 2156 & 908200\\
+ & Glasgow2 & 198944 & 2129 & 1014522\\
   \midrule
 $10^8$ & virtual best & 0 & 2219 & 5822809\\
- & LLAMA & 740765 & 2204 & 6565232\\
- & glasgow2 & 1969418 & 2172 & 7792227\\
+ & \LLAMA & 740765 & 2204 & 6565232\\
+ & Glasgow2 & 1969418 & 2172 & 7792227\\
 \bottomrule
@@
 
-The results show that while for the misclassification penalty, LLAMA is always
+The results show that while for the misclassification penalty, \LLAMA is always
 better than the single best algorithm, for timeouts $10^2$ to $10^4$ it is worse
@@
 \hline
-$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
-&&pre&def&new&1&2&3&4&&\\\hline
+$t$ & \VFtwo & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{\Glasgow}& VBS & $\delta$\\
+&&incomplete&classic&path&1&2&3&4&&\\\hline
 $10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
@@
 \caption{Number of solved instances at different CPU time limits: Each line displays a time limit
-$t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD
-preprocessing, default LAD, and new LAD, Glasgow with the lengths of paths limited to 1, 2, 3 and
+$t$ (in milliseconds) followed by the number of instances solved within this time limit by \VFtwo,
+\IncompleteLAD, classical \LAD, and \PathLAD, \Glasgow with the lengths of paths limited to 1, 2, 3 and
 4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
@@
 ranging from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time
-limit considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances
-very quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
-preprocessing is an incomplete approach which can only detect rather trivial inconsistencies; VF2
+limit considered. Simple approaches like \IncompleteLAD and \VFtwo are able to solve easy instances
+very quickly, in a few milliseconds. However, they are not able to solve harder instances:
+\IncompleteLAD is an incomplete approach which can only detect rather trivial inconsistencies; \VFtwo
 performs a basic backtracking search and is very fast on easy instances, because it does not compute
-expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
-and LAD preprocessing when considering longer time limits: \SI{10}{\ms} for default and new LAD and
-Glasgow 1, and \SI{100}{\ms} (resp. \SI{1000}{\ms} and \SI{10000}{\ms}) for Glasgow(2) (resp.\ 3 and
+expensive invariants. Classical \LAD, \PathLAD, and the \Glasgow algorithms solve more instances than \VFtwo
+and \IncompleteLAD when considering longer time limits: \SI{10}{\ms} for \LAD and \PathLAD and
+\GlasgowOne, and \SI{100}{\ms} (resp. \SI{1000}{\ms} and \SI{10000}{\ms}) for \GlasgowTwo (resp.\ 3 and
 4). The Virtual Best Solver (VBS), which considers the best algorithm for each instance separately,
@@
 particularly important for rather small CPU time limits. In particular, VBS solves 693 more
-instances than the single best (Glasgow(1)) when the limit is \SI{100}{\ms}. In many applications,
+instances than the single best (\GlasgowOne) when the limit is \SI{100}{\ms}. In many applications,
 it is important to have the fastest possible algorithm. For example, in pattern recognition
@@
 in blue in Table 1).  Comment Table 3: Mean runtime but also number of solved instances at different
-CPU time limits.  Logically, LLAMA may solve less instances than single best for small time limits
+CPU time limits.  Logically, \LLAMA may solve less instances than single best for small time limits
 because it spends time to compute features and choose a solver, but it should be better for larger
@@
 Single best & 3177.6 & 1126 & 2287 & 3943 & 4908 & 5253 & 5377 & 5452 & 5518 & 5561\\\hline
-LLAMA & 2706.3\\\hline
+\LLAMA & 2706.3\\\hline
 \end{tabular}
@@
 
-Add Figure 1 which compares single best at \SI{10e8}{\ms}(Glasgow(2)) with LLAMA ?
+Add Figure 1 which compares single best at \SI{10e8}{\ms}(\GlasgowTwo) with \LLAMA ?
 

### 2015-12-14 paper/paper.tex
Remove packages we don't use
Silly mistakes

@@
 \usepackage{microtype}
-\usepackage{cleveref}                  % no need to type Figure etc
 \usepackage{xcolor,colortbl}
@@
 \usepackage{scalefnt}
-
-\usepackage{todonotes}
+\usepackage{graphicx}
 \usepackage{booktabs}
-
-% lncs style
-\crefname{algocf}{Algorithm}{Algorithms}
-\Crefname{algocf}{Algorithm}{Algorithms}
-\crefname{figure}{Fig.}{Figs.}
-\Crefname{figure}{Fig.}{Figs.}
-\crefname{table}{Table}{Tables}
-\Crefname{table}{Table}{Tables}
-\crefname{proposition}{Proposition}{Propositions}
-\Crefname{proposition}{Proposition}{Propositions}
 
@@
 achieve. The single best solver is the one solver from the portfolio that has the overall best
-performance across the entire set of instances, at the CPU time limit of \SI{10e8}{\ms},
-i.e.\ GlasgowTwo. We consider it a lower bound on performance.
+performance across the entire set of instances, at the CPU time limit of \SI{e8}{\ms},
+i.e.\ \GlasgowTwo. We consider it a lower bound on performance.
 We are able to solve 34 more instances than the single best solver within the timeout, with only an
@@
 
-Add Figure 1 which compares single best at \SI{10e8}{\ms}(\GlasgowTwo) with \LLAMA ?
+Add Figure 1 which compares single best at \SI{e8}{\ms}(\GlasgowTwo) with \LLAMA ?
 

### 2015-12-14 paper/paper.tex
first draft of conclusion

--- 2015-12-14 paper/paper.tex
@@
 
+The problem of identifying subgraph isomorphisms is a hard computation problem
+that has many applications in diverse areas. In this paper, we presented a
+portfolio of six algorithms from the literature and a novel variant of the \LAD
+algorithm. We introduced a set of novel features to characterise subgraph
+isomorphism problems and leveraged them to select the most appropriate algorithm
+from the portfolio for each instance.
+
+We demonstrated that our portfolio achieves substantial performance improvements
+over the single algorithm that has the best performance on our benchmark set. We
+showed that combining a portfolio approach with a novel incomplete variant of
+\LAD that is able to detect inconsistencies and a presolver boosts performance
+even further.
+
 Mention parallel, directed, labelled in here briefly.

### 2015-12-15 paper/paper.tex
Christine's changes

@@
-% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :
-
 \documentclass{llncs}
@@
         and Physical Sciences Research Council [grant number EP/K503058/1]}\inst{2}
-    \and Christine Solnon\inst{3}}
+    \and Christine Solnon\inst{3}\thanks{This work has been supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01).}}
 
@@
 better success rates for larger time limits.  They also showed that on an instance by instance
-basis, other algorithms are often better.
+basis, other algorithms are often better. This motivates us for introducing and evaluating a per-instance algorithm selection  approach for this problem.
 
-In this paper, we report experimental results on a larger set of 8 algorithms, including new
-variants, and a larger benchmark. We show that 4 of these algorithms are single best algorithms,
-depending on the CPU time limit considered, and that combining preprocessing with an algorithm
-selection on an instance by instance basis allows us to achieve better overall performance than any
-single algorithm.
+The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
+algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
+best on a given problem instance. Algorithm selection systems usually build machine learning models
+of the algorithms or the in which portfolio they are contained to forecast which algorithm to use in a
+particular context. Using the model predictions, one or more algorithms from the portfolio are
+selected to be run sequentially or in parallel.
+
+Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
+the most prominent and successful systems that employs this approach is
+SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
+years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
+salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
+The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
+information on algorithm selection.
+
+\paragraph{Overview of the paper.}
+We formally define the subgraph isomorphism problem in Section 2. 
+In Section 3, we describe the main existing algorithms for solving this problem, and we also introduce two new algorithms which are derived from the LAD algorithm of \cite{Solnon:2010}.
+
+In Section 4, we experimentally compare some algorithms.
+We describe a large benchmark composed of 5725 instances grouped into 12 classes: 10 of these classes have already been used in the experimental study reported in  \cite{McCreesh:2015}, and 2 are new ones.
+We evaluate 8 algorithms on this benchmark (6 existing algorithms, and our 2 new algorithms), and we show that they have very compementary performance.
+In particular, we show that 4 of these algorithms are single best algorithms, depending on the CPU time limit considered.
+
+In Section 5, we describe the features that are used to describe instances, and we describe our algorithm selection approach.
+It combines a presolving step, which allows us to very quickly solve easy instances, with the algorithm selection approach \LLAMA~\cite{kotthoff_llama_2013}.
+
+In Section 6, we experimentally evaluate our selection approach, and show that it is able to close almost 70\% of the gap between the single best and the virtual best solver.
 
@@
 \times N$, where an edge $(u,u')$ is a pair of nodes. The number of neighbors of a node $u$ is
-called the degree of $u$, $d^\circ(u)=\#\{ (u,u')\in E\}$. In this paper, we implicitly consider
+called the degree of $u$, denoted $d^\circ(u)=\#\{ (u,u')\in E\}$. In this paper, we implicitly consider
 non-directed graphs, such that $(u,u')\in E\Leftrightarrow (u',u)\in E$. The extension to directed
@@
 \in E_p, (f(u),f(u')) \in E_t$.
-
 Note that the subgraph is not necessarily induced, so that two pattern nodes that are not linked by
@@
 
-In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of a
-subgraph isomorphism problem.  We also define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\#
+In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of 
+subgraph isomorphism.  We  define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\#
 E_t$, and $d_p$ and $d_t$ to be the maximum degrees of the graphs $G_p$ and $G_t$.
@@
 
-\subsection{Per-Instance Algorithm Selection}
+\section{Experimental Comparison of Individual Algorithms}
 
-The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
-algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
-best on a given problem instance. Algorithm selection systems usually build machine learning models
-of the algorithms or the in which portfolio they are contained to forecast which algorithm to use in a
-particular context. Using the model predictions, one or more algorithms from the portfolio are
-selected to be run sequentially or in parallel.
+\subsection{Considered Existing Algorithms}
 
-Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
-the most prominent and successful systems that employs this approach is
-SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
-years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
-salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
-The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
-information on algorithm selection.
+We have selected a set of algorithms with complementary performance. These algorithms will compose our portfolio in the per-instance selection approach described in Section 5. 
+\begin{itemize}
+\item \VFtwo \cite{Cordella:2004} performs a weak filtering that it is especially fast on
+    trivially satisfiable instances;
+\item \LAD \cite{Solnon:2010} combines two strong but expensive filtering techniques
+    (All-Different Generalized Arc Consistency and Locally All-Different);
+\item \Glasgow \cite{McCreesh:2015} does expensive preprocessing based on path length
+    invariant properties to generate additional constraints, followed by weaker filtering
+    (forward-checking, and a heuristic All-Different propagator which can miss deletions) and
+    conflict-directed backjumping during search.
+\end{itemize}
+
+\noindent The \Glasgow algorithm has a parameter, which controls the lengths of paths used when
+reasoning about non-adjacent vertices.  In experiments reported by McCreesh and Prosser
+\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer
+paths lead to prohibitively expensive preprocessing on larger, denser instances. This is often not
+the best choice on an instance by instance basis: sometimes path-based reasoning gives no benefit at
+all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are
+helpful, and even looking at paths of length 3 is relatively expensive on some graphs. We thus consider all
+four choices, naming these variants \GlasgowOne through \GlasgowFour.
+
+\subsection{New algorithms}
+
+We also introduce two new variants of \LAD. The first, called \IncompleteLAD, is a weaker
+filtering which is applied once, without performing a backtracking search, and very quickly
+detects inconsistencies on many instances: for each pattern node $u$, we check that there exists at
+least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor
+$v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
+\IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when
+it does not detect inconsistency, the instance may either be satisfiable or not. As a counterpart of
+this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.
+
+The second variant of \LAD is called \PathLAD. It combines the locally all-different constraints
+introduced by Solnon \cite{Solnon:2010} with the exploitation of path length properties proposed by
+Audemard et al.\ \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of
+paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3
+passing through $u$, and to add the constraint that the label of a pattern node (resp.\ edge) must
+be smaller than or equal to the label of its associated target node (resp.\ edge).
 
@@
 
-We consider 5725 instances, which are available in a simple text
+We consider a large benchmark of 5725 instances, which are available in a simple text
 format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
@@
 
-\subsection{Portfolio Composition}
-
-We have selected a set of algorithms with complementary performance:
-\begin{itemize}
-\item \VFtwo \cite{Cordella:2004}, which performs a weak filtering that it is especially fast on
-    trivially satisfiable instances;
-\item \LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques
-    (All-Different Generalized Arc Consistency and Locally All-Different);
-\item \Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
-    invariant properties to generate additional constraints, followed by weaker filtering
-    (forward-checking, and a heuristic All-Different propagator which can miss deletions) and
-    conflict-directed backjumping during search.
-\end{itemize}
-
-\noindent The \Glasgow algorithm has a parameter, which controls the lengths of paths used when
-reasoning about non-adjacent vertices.  In experiments reported by McCreesh and Prosser
-\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer
-paths lead to prohibitively expensive preprocessing on larger, denser instances. This is often not
-the best choice on an instance by instance basis: sometimes path-based reasoning gives no benefit at
-all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are
-helpful, and even looking at paths of length 3 is relatively expensive on some graphs. We thus all
-four choices, naming these variants \GlasgowOne through \GlasgowFour.
-
-We also introduce two new variants of \LAD. The first, called \IncompleteLAD, is a weaker
-filtering which is applied once, without performing a backtracking search, and very quickly
-detects inconsistencies on many instances: for each pattern node $u$, we check that there exists at
-least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor
-$v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
-\IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when
-it does not detect inconsistency, the instance may either be satisfiable or not. As a counterpart of
-this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.
-
-The second variant of \LAD is called \PathLAD, which combines the locally all-different constraints
-introduced by Solnon \cite{Solnon:2010} with the exploitation of path length properties proposed by
-Audemard et al.\ \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of
-paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3
-passing through $u$, and to add the constraint that the label of a pattern node (resp.\ edge) must
-be smaller than or equal to the label of its associated target node (resp.\ edge).
-
-\subsection{Results for Individual Algorithms} \label{expComp}
+\subsection{Results} \label{expComp}
 
@@
 \begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
+%\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
 \caption{Empirical cumulative distribution function of solvers, virtual best
@@
 \begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figures/portfolio-ecdf-full}
+%\includegraphics[width=\textwidth]{figures/portfolio-ecdf-full}
 \caption{Empirical cumulative distribution function of solvers, virtual best

### 2015-12-15 paper/paper.tex
Turn graphs back on
Consistent ordering for thanks and inst
we don't need no stinking CRLFs
funding acks

@@
 \author{
-    Lars Kotthoff\inst{1}
+    Lars Kotthoff\thanks{This work was supported by an NSERC E.W.R.\ Steacie
+    Fellowship and under the NSERC Discovery Grant Program.}\inst{1}
     \and Ciaran McCreesh\thanks{This work was supported by the Engineering
         and Physical Sciences Research Council [grant number EP/K503058/1]}\inst{2}
-    \and Christine Solnon\inst{3}\thanks{This work has been supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01).}}
+    \and Christine Solnon\thanks{This work has been supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01).}\inst{3}}
 
@@
 \begin{figure}[!ht]
-%\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
+\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
 \caption{Empirical cumulative distribution function of solvers, virtual best
@@
 \begin{figure}[!ht]
-%\includegraphics[width=\textwidth]{figures/portfolio-ecdf-full}
+\includegraphics[width=\textwidth]{figures/portfolio-ecdf-full}
 \caption{Empirical cumulative distribution function of solvers, virtual best

### 2015-12-15 paper/paper.tex
intro pass

@@
 \usepackage{booktabs}
+\usepackage[sort]{cite}
 
@@
 novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm
-to use on a per-instance basis. We demonstrate significant performance improvements on a large set
+to use on a per-instance basis. We demonstrate substantial performance improvements on a large set
 of hard benchmark problems. In addition, we show how algorithm selection models can be leveraged to
@@
 better success rates for larger time limits.  They also showed that on an instance by instance
-basis, other algorithms are often better. This motivates us for introducing and evaluating a per-instance algorithm selection  approach for this problem.
+basis, other algorithms are often better.
 
 The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
-algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
+algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the
+algorithm expected to perform
 best on a given problem instance. Algorithm selection systems usually build machine learning models
-of the algorithms or the in which portfolio they are contained to forecast which algorithm to use in a
-particular context. Using the model predictions, one or more algorithms from the portfolio are
-selected to be run sequentially or in parallel.
+of the algorithms or the portfolio which they are contained in to forecast which algorithm to use in a
+particular context. Using the predictions, one or more algorithms from the
+portfolio can be selected to be run sequentially or in parallel.
 
@@
 \paragraph{Overview of the paper.}
-We formally define the subgraph isomorphism problem in Section 2. 
-In Section 3, we describe the main existing algorithms for solving this problem, and we also introduce two new algorithms which are derived from the LAD algorithm of \cite{Solnon:2010}.
+We formally define the subgraph isomorphism problem in Section~\ref{sec:defs}. 
+In Section~\ref{sec:algs}, we describe the main existing algorithms for solving this problem, and we also introduce two new algorithms which are derived from the LAD algorithm of \cite{Solnon:2010}.
 
-In Section 4, we experimentally compare some algorithms.
-We describe a large benchmark composed of 5725 instances grouped into 12 classes: 10 of these classes have already been used in the experimental study reported in  \cite{McCreesh:2015}, and 2 are new ones.
-We evaluate 8 algorithms on this benchmark (6 existing algorithms, and our 2 new algorithms), and we show that they have very compementary performance.
-In particular, we show that 4 of these algorithms are single best algorithms, depending on the CPU time limit considered.
+In Section~\ref{sec:algs}, we experimentally compare eight state-of-the-art
+algorithms. We introduce a large benchmark set composed of 5725 instances
+composed of twelve classes. Ten of these classes have been introduced in the
+experimental study reported in~\cite{McCreesh:2015}. We evaluate the algorithms,
+of which two are novel, on this benchmark set, and show that they have very
+complementary performance. In particular, we show that depending on the CPU time
+limit, different algorithms achieve the best performance on the entire benchmark
+set.
 
-In Section 5, we describe the features that are used to describe instances, and we describe our algorithm selection approach.
+In Section~\ref{sec:algsel}, we describe the features that are used to describe instances, and we describe our algorithm selection approach.
 It combines a presolving step, which allows us to very quickly solve easy instances, with the algorithm selection approach \LLAMA~\cite{kotthoff_llama_2013}.
 
-In Section 6, we experimentally evaluate our selection approach, and show that it is able to close almost 70\% of the gap between the single best and the virtual best solver.
+In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
+approach, and show that it is able to close almost 70\% of the gap between the
+single best and the virtual best solver.
 
-\section{Definitions and Notations}
+We conclude and give directions for future work in Section~\ref{sec:concs}.
+
+\section{Definitions and Notations}\label{sec:defs}
 
@@
 
-\section{Subgraph Isomorphism Algorithms}
+\section{Subgraph Isomorphism Algorithms}\label{sec:algs}
 
@@
 
-\section{Experimental Comparison of Individual Algorithms}
+\section{Experimental Comparison of Individual Algorithms}\label{sec:exps}
 
@@
 
-\section{Algorithm Selection Approach}
+\section{Algorithm Selection Approach}\label{sec:algsel}
 
@@
 
-\section{Experimental Evaluation of Algorithm Selection}
+\section{Experimental Evaluation of Algorithm Selection}\label{sec:algsel-exps}
 
@@
 
-\section{Conclusion and Future Work}
+\section{Conclusion and Future Work}\label{sec:concs}
 

### 2015-12-15 paper/paper.tex
pass up to experiments

@@
 \usepackage[sort]{cite}
+\usepackage{enumitem}
+\setlist{nosep}
 
@@
 of a small \emph{pattern} graph to vertices of a large \emph{target} graph. This \NP-complete
-problem has many important practical applications, for example in computer vision
-\cite{cviu11,pr15}, biochemistry \cite{Giugno:2013}, and model checking \cite{Sevegnani:2015}. There
+problem has many important practical applications, for example in computer
+vision~\cite{cviu11,pr15}, biochemistry~\cite{Giugno:2013}, and model checking~\cite{Sevegnani:2015}. There
 exist various exact algorithms, which have been compared on a large suite of problem instances by
-McCreesh and Prosser \cite{McCreesh:2015}. These experiments indicated that the single best
-algorithm depends on the CPU time limit considered: for very small CPU time limits, \VFtwo
-\cite{Cordella:2004} is the best choice, whereas the \Glasgow algorithm \cite{McCreesh:2015} has
+McCreesh and Prosser~\cite{McCreesh:2015}. These experiments indicated that the single best
+algorithm depends on the CPU time limit considered: for very small CPU time
+limits, \VFtwo{}~\cite{Cordella:2004} is the best choice, whereas the \Glasgow algorithm~\cite{McCreesh:2015} has
 better success rates for larger time limits.  They also showed that on an instance by instance
@@
 We formally define the subgraph isomorphism problem in Section~\ref{sec:defs}. 
-In Section~\ref{sec:algs}, we describe the main existing algorithms for solving this problem, and we also introduce two new algorithms which are derived from the LAD algorithm of \cite{Solnon:2010}.
+In Section~\ref{sec:algs}, we describe the main existing algorithms for solving this problem, and we also introduce two new algorithms which are derived from the LAD algorithm of~\cite{Solnon:2010}.
 
@@
 isomorphism problem} consists of deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
-More precisely, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a
-different target node to each pattern node, and that preserves pattern edges, i.e.\ $\forall (u,u')
+More formally, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a
+different target node to each pattern node, and preserves pattern edges, i.e.\ $\forall (u,u')
 \in E_p, (f(u),f(u')) \in E_t$.
-Note that the subgraph is not necessarily induced, so that two pattern nodes that are not linked by
+Note that the subgraph is not necessarily induced, so that two pattern nodes not linked by
 an edge may be mapped to two target nodes which are linked by an edge. This problem is also called
@@
 difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints. It is used, for
-example, in McGregor's algorithm \cite{mcgregor79} and in \VFtwo \cite{Cordella:2004}.
+example, in McGregor's algorithm~\cite{mcgregor79} and in \VFtwo{}~\cite{Cordella:2004}.
 
-R\'egin \cite{regin} introduced a stronger filtering for difference constraints, which ensures that
+R\'egin~\cite{regin} introduced a stronger filtering for difference constraints, which ensures that
 all pattern nodes can be matched with different target nodes, all together. This filtering (called
@@
 difference constraint is propagated separately which Forward-Checking. However, it is also more time
-consuming as it is done in ${\cal O}(n_p^2\cdot n_t^2)$ time.
+consuming as requires ${\cal O}(n_p^2\cdot n_t^2)$ time.
 
-Various filtering techniques have been tried for edge constraints. Ullman \cite{ullman} introduced a
+Various filtering techniques have been tried for edge constraints. Ullman~\cite{ullman} introduced a
 filtering which ensures that for each pattern edge $(u,u')\in E_p$ and each candidate pair $(u,v)\in
@@
 until a fixed point is reached. This filtering (called \emph{Arc Consistency}) removes more
-candidate pairs than Forward-Checking, but it is also more time consuming as it is done in ${\cal
-O}(e_p\cdot n_t^2)$ when using the algorithm AC4  \cite{MH86}.
+candidate pairs than Forward-Checking, but it is also more time consuming as it
+runs in ${\cal
+O}(e_p\cdot n_t^2)$ when using AC4~\cite{MH86}.
 
 Stronger filtering may be obtained by propagating edge constraints in a more global way, as proposed
-by Larrosa and Valiente \cite{LV02}. The idea is to check for each candidate pair $(u,v)\in
+by Larrosa and Valiente~\cite{LV02}. The idea is to check for each candidate pair $(u,v)\in
 N_p\times N_t$ that the number of pattern nodes adjacent to $u$ is smaller than or equal to the
 number of target nodes that are both adjacent to $v$ and that may be matched with nodes adjacent to
-$u$. This is done in ${\cal O}(n_p^2\cdot n_t^2)$. This idea was generalised by Solnon
-\cite{Solnon:2010}'s \LAD algorithm, where, for each candidate pair $(u,v)\in N_p\times N_t$, a redundant Local
+$u$. This is done in ${\cal O}(n_p^2\cdot n_t^2)$. This idea was generalised by
+Solnon's \LAD algorithm~\cite{Solnon:2010}, where, for each candidate pair $(u,v)\in N_p\times N_t$, a redundant Local
 All-Different constraint ensures that each neighbour of $u$ may be matched with a different
@@
 
-These invariant properties are used, for example:
-
+These invariant properties are used, for example,
 \begin{itemize}
-\item  By Battiti and Mascia \cite{battiti-mascia07}, to remove candidate pairs $(u,v)\in N_p\times
+\item by Battiti and Mascia~\cite{battiti-mascia07}, to remove candidate pairs $(u,v)\in N_p\times
     N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of
     paths starting from  target node $v$;
-\item By Audemard et al.\ \cite{Audemard:2014} to generalise the locally all-different constraint
-    proposed by Solnon \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be
+\item by Audemard et al.~\cite{Audemard:2014} to generalise the locally all-different constraint
+    proposed by Solnon~\cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be
     matched with all different compatible target nodes, where compatibility is defined with respect
     to invariant properties;
-\item by McCreesh and Prosser \cite{McCreesh:2015} to filter the set of candidate couples before
+\item by McCreesh and Prosser~\cite{McCreesh:2015} to filter the set of candidate couples before
     starting the search, and to generate additional implied adjacency-like constraints which are
@@
 
-\noindent Audemard et al.\ \cite{Audemard:2014} do not limit the length of paths considered, and
-iteratively increment the length until no more pairs are removed. Battiti and Mascia
-\cite{battiti-mascia07}, and McCreesh and Prosser \cite{McCreesh:2015} parameterise their algorithms
+Audemard et al.~\cite{Audemard:2014} do not limit the length of paths considered, and
+iteratively increment the length until no more pairs are removed. Battiti and Mascia~\cite{battiti-mascia07}, and McCreesh and Prosser~\cite{McCreesh:2015} parameterise their algorithms
 by the maximum path length considered when counting paths: larger values for this parameter remove
@@
 versions in time-sharing decreases the total CPU time needed to find a solution for feasible
-instances. McCreesh and Prosser simply set the parameter to 3, as this setting seemed to be a
-reasonable compromise.
+instances. McCreesh and Prosser simply set the parameter to 3, as this setting
+presented the best overall performance in their case.
 
@@
 \begin{itemize}
-\item \VFtwo \cite{Cordella:2004} performs a weak filtering that it is especially fast on
+\item \VFtwo{}~\cite{Cordella:2004} performs a weak filtering that it is especially fast on
     trivially satisfiable instances;
-\item \LAD \cite{Solnon:2010} combines two strong but expensive filtering techniques
+\item \LAD{}~\cite{Solnon:2010} combines two strong but expensive filtering techniques
     (All-Different Generalized Arc Consistency and Locally All-Different);
-\item \Glasgow \cite{McCreesh:2015} does expensive preprocessing based on path length
+\item \Glasgow{}~\cite{McCreesh:2015} does expensive preprocessing based on path length
     invariant properties to generate additional constraints, followed by weaker filtering
@@
 \noindent The \Glasgow algorithm has a parameter, which controls the lengths of paths used when
-reasoning about non-adjacent vertices.  In experiments reported by McCreesh and Prosser
-\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer
+reasoning about non-adjacent vertices.  In experiments reported by McCreesh and
+Prosser~\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer
 paths lead to prohibitively expensive preprocessing on larger, denser instances. This is often not
@@
 The second variant of \LAD is called \PathLAD. It combines the locally all-different constraints
-introduced by Solnon \cite{Solnon:2010} with the exploitation of path length properties proposed by
-Audemard et al.\ \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of
+introduced by Solnon~\cite{Solnon:2010} with the exploitation of path length properties proposed by
+Audemard et al.~\cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of
 paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3
@@
 \begin{itemize}
-\item Class 1 contains randomly generated scale-free graphs \cite{constraints10}.
-\item Classes 2 and 3 contain various kinds of graphs selected by Larrosa and Valiente \cite{LV02}:
+\item Class 1 contains randomly generated scale-free graphs~\cite{constraints10}.
+\item Classes 2 and 3 contain various kinds of graphs selected by Larrosa and Valiente~\cite{LV02}:
     class 2 contains small instances generated from the 50 first graphs, and class 3 contains
@@
 \item Classes 4 to 8 contain randomly generated graphs coming from a database of graphs commonly
-    used for benchmarking subgraph isomorphism algorithms
-    \cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
+    used for benchmarking subgraph isomorphism algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
     for classes 5 and 6, and random graphs with uniform edge probabilities for class 7. All of these
     instances are satisfiable.
-\item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
-\item Class 11 contains instances coming from meshes modeling 3D objects \cite{cviu11}.
+\item Classes 9 and 10 contain instances coming from segmented images~\cite{pr15,cviu11}.
+\item Class 11 contains instances coming from meshes modeling 3D objects~\cite{cviu11}.
 \item Class 12 contains random graph instances chosen to be close to the satisfiable-unsatisfiable
@@
 Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
-Prosser \cite{McCreesh:2015}.  This set of instances is also much larger than that of Battiti and
+Prosser~\cite{McCreesh:2015}.  This set of instances is also much larger than that of Battiti and
 Mascia~\cite{battiti-mascia07}, who were the first to propose algorithm portfolios for subgraph
@@
 We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
-Scientific Linux 6.5. We used the C++ implementation of the \Glasgow algorithm \cite{McCreesh:2015},
-the C implementation of \LAD \cite{Solnon:2010}, and the VFLib C implementation of \VFtwo
-\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
+Scientific Linux 6.5. We used the C++ implementation of the \Glasgow algorithm~\cite{McCreesh:2015},
+the C implementation of \LAD{}~\cite{Solnon:2010}, and the VFLib C
+implementation of \VFtwo{}~\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
 timeout of $10^8$ milliseconds (a little over a day).
@@
 small CPU time limits (smaller than \SI{1000}{\ms}). In many applications, it is important to have the
-fastest possible algorithm. For example, in pattern recognition \cite{pr15,cviu11} and chemical
-\cite{Giugno:2013} applications, we often have to solve the subgraph isomorphism problem repeatedly
+fastest possible algorithm. For example, in pattern
+recognition~\cite{pr15,cviu11} and chemical~\cite{Giugno:2013} applications, we often have to solve the subgraph isomorphism problem repeatedly
 for a very large number of graphs (in order to find a pattern image or molecule in a large database
@@
 \LLAMA's pairwise regression approach with random forest regression gave the best performance. The
-idea is very similar to the pairwise classification used by Xu et al.\ \cite{xu_satzilla_2008}. For
+idea is very similar to the pairwise classification used by Xu et al.~\cite{xu_satzilla_2008}. For
 each pair of algorithms in our portfolio, we train a model that predicts the performance difference

### 2015-12-15 paper/paper.tex
proofread up to section 6

@@
 
-\subsection{Considered Existing Algorithms}
+We consider six algorithms from the literature and propose two novel ones.
 
-We have selected a set of algorithms with complementary performance. These algorithms will compose our portfolio in the per-instance selection approach described in Section 5. 
+\subsection{Algorithms from the Literature}
+
+We have selected a set of algorithms with complementary performance. These
+algorithms will compose our portfolio in the per-instance selection approach
+described in Section~\ref{sec:algsel}.
 \begin{itemize}
-\item \VFtwo{}~\cite{Cordella:2004} performs a weak filtering that it is especially fast on
+\item \VFtwo{}~\cite{Cordella:2004} performs weak filtering that it is especially fast on
     trivially satisfiable instances;
@@
 helpful, and even looking at paths of length 3 is relatively expensive on some graphs. We thus consider all
-four choices, naming these variants \GlasgowOne through \GlasgowFour.
+lengths up to 4, naming these variants \GlasgowOne through \GlasgowFour.
 
-\subsection{New algorithms}
+\subsection{Novel Algorithms}
 
-We also introduce two new variants of \LAD. The first, called \IncompleteLAD, is a weaker
+We introduce two new variants of \LAD. The first, called \IncompleteLAD, does weaker
 filtering which is applied once, without performing a backtracking search, and very quickly
@@
 $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
-\IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when
-it does not detect inconsistency, the instance may either be satisfiable or not. As a counterpart of
-this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.
+\IncompleteLAD is an incomplete algorithm that checks a sufficient, but not
+necessary, condition for inconsistency: when it does not detect inconsistency,
+the instance may still be unsatisfiable. Its main benefit is that it runs very
+fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.
 
@@
 
-We consider a large benchmark of 5725 instances, which are available in a simple text
+We consider a large benchmark set of 5725 instances, which are available in a simple text
 format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
@@
 \item Classes 2 and 3 contain various kinds of graphs selected by Larrosa and Valiente~\cite{LV02}:
-    class 2 contains small instances generated from the 50 first graphs, and class 3 contains
-    larger instances where pattern graphs are coming from class 2, and target graphs are coming from
+    class 2 contains small instances generated from the 50 first graphs of class
+    1 and class 3 contains
+    larger instances with pattern graphs from class 2 and target graphs from
     the next 50 graphs.
-\item Classes 4 to 8 contain randomly generated graphs coming from a database of graphs commonly
-    used for benchmarking subgraph isomorphism algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
-    for classes 5 and 6, and random graphs with uniform edge probabilities for class 7. All of these
+\item Classes 4 to 8 contain randomly generated graphs from a database of graphs commonly
+    used for benchmarking subgraph isomorphism algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 4 and 5, regular meshes
+    for classes 6 and 7, and random graphs with uniform edge probabilities for class 8. All of these
     instances are satisfiable.
-\item Classes 9 and 10 contain instances coming from segmented images~\cite{pr15,cviu11}.
-\item Class 11 contains instances coming from meshes modeling 3D objects~\cite{cviu11}.
+\item Classes 9 and 10 contain instances from segmented images~\cite{pr15,cviu11}.
+\item Class 11 contains instances from meshes modeling 3D objects~\cite{cviu11}.
 \item Class 12 contains random graph instances chosen to be close to the satisfiable-unsatisfiable
@@
 two solvers, and without a selection mechanism. Their problem set consisted entirely of easy
-satisfiable instances, which is where this style of approach has most success.
+satisfiable instances.
 
@@
 implementation of \VFtwo{}~\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
-timeout of $10^8$ milliseconds (a little over a day).
+timeout of $10^8$ milliseconds ($\approx$ 27.8 hours).
 
-\subsection{Results} \label{expComp}
+\subsection{Results}\label{expComp}
 
@@
 \includegraphics[width=\textwidth]{courbe.pdf}
-\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms
-of the portfolio, and a Virtual Best Solver (VBS).\label{expTimeGraph}}
+\caption{Number of solved instances over CPU time for the eight algorithms we
+consider in this paper, and the virtual best solver (VBS) that shows the best
+solver on an instance by instance basis.\label{expTimeGraph}}
 \end{figure}
 
-Figure \ref{expTimeGraph} displays the evolution of the number of instances solved with respect to
+Figure~\ref{expTimeGraph} displays the evolution of the number of instances solved with respect to
 CPU time. It shows us that the best solver depends on the time limit considered. \IncompleteLAD is
@@
 
-\PathLAD and \GlasgowOne outperform \IncompleteLAD when considering longer time limits: \PathLAD is the
+\PathLAD and \GlasgowOne outperform \IncompleteLAD for longer time limits: \PathLAD is the
 best solver for time limits larger than \SI{5}{\ms} and lower than \SI{40}{\ms}, and \GlasgowOne is the best solver
@@
 
-For CPU time limits larger than \SI{3000}{\ms}, the best solver becomes \GlasgowTwo.  As we
+\GlasgowTwo becomes the best solver for time limits larger than \SI{3000}{\ms}.
+As we
 increase the CPU time limit, variants of \Glasgow with longer paths (\GlasgowThree and \GlasgowFour)
-become better. This is what we expect, as more reasoning is expensive, but potentially increases the
-reduction of the search space. In particular, the relative performance of Glasgow variants is
+become better. This is what we expect, as more reasoning is expensive, but
+increases the potential
+reduction of the search space. In particular, the relative performance of \Glasgow variants is
 monotone in the time it takes to solve the instance---\GlasgowOne is best at first, but gets worse as
@@
 
-Furthermore, the Virtual Best Solver (VBS), which considers the best algorithm for each instance
-separately, obtains much better results, showing us that the algorithms have complementary
-performance. The difference between VBS and single best is more particularly important for rather
-small CPU time limits (smaller than \SI{1000}{\ms}). In many applications, it is important to have the
-fastest possible algorithm. For example, in pattern
-recognition~\cite{pr15,cviu11} and chemical~\cite{Giugno:2013} applications, we often have to solve the subgraph isomorphism problem repeatedly
-for a very large number of graphs (in order to find a pattern image or molecule in a large database
-of target images or compounds, for example), so having an algorithm that is able to solve an
-instance in \SI{100}{\ms} instead of \SI{1000}{\ms} makes a big difference.  Therefore, it is important to select
-the best algorithm for each instance, even if the instance is an easy one.
+Furthermore, the virtual best solver (VBS), which considers the best algorithm
+for each instance separately, obtains much better results, showing us that the
+algorithms have complementary performance. The difference between VBS and single
+best is particularly pronounced for CPU time limits less than \SI{1000}{\ms}. In
+many applications, it is important to have the fastest possible algorithm even
+if the absolute differences in CPU time are small. For example, in pattern
+recognition~\cite{pr15,cviu11} and chemical~\cite{Giugno:2013} applications, we
+often have to solve the subgraph isomorphism problem repeatedly for a very large
+number of graphs (in order to find a pattern image or molecule in a large
+database of target images or compounds, for example), so having an algorithm
+that is able to solve an instance in \SI{100}{\ms} instead of \SI{1000}{\ms}
+makes a big difference.  Therefore, it is important to select the best algorithm
+for each instance, even if the instance is an easy one.
 
-Table~\ref{expClass} shows us that we cannot simply select algorithms based upon which class from
-which instances are coming (unlike in SAT, where this kind of approach can be reasonable). For all
+Table~\ref{expClass} shows us that we cannot simply select algorithms based on
+the instance class. For all
 classes, there are always at least two algorithms which are the best for at least one instance of
@@
 Experimental results reported in Section~\ref{expComp} have shown us that \IncompleteLAD is very fast
-(\SI{7}{\ms} on average) and is able to solve 1919 instances very quickly. Therefore, we first run
-\IncompleteLAD: if the instance is detected to be unsatisfiable, then we do not process it further.
-Otherwise, we record as features the number of candidate couples removed by \IncompleteLAD in absolute
-terms, as a percentage, and the minimum and maximum on a per-variable basis.
+(\SI{7}{\ms} on average) and is able to solve 1919 instances from our benchmark
+set very quickly. Therefore, we first run
+\IncompleteLAD: if unsatisfiability is detected, we do not need to process it further.
+Otherwise, we record the number of candidate pairs removed and as a percentage
+of all pairs and the minimum and maximum values on a per-variable basis. The
+recorded values are then used as features for algorithm selection.
 
@@
 the overhead of running algorithm selection and avoids potentially making incorrect solver choices.
-After running \IncompleteLAD and \VFtwo for \SI{50}{\ms}, we are left with 2336 hard instances that we
-consider for algorithm selection.
+
+After the presolving step, we are left with 2336 hard instances that we consider
+for algorithm selection.
 
@@
 If presolving does not give us a solution, we extract additional features. For both the pattern and
-the target graph, we consider some basic graph properties, which may be computed very quickly:
-
+the target graph, we consider some basic graph properties, which can be computed
+very quickly,
 \begin{itemize}
-    \item The number of vertices and edges.
-    \item The density---we expect that some kinds of filtering might be expensive and ineffective on
-        dense graphs.
-    \item How many loops (self-adjacent vertices) the graph contains---since loops must be mapped to
-        loops, this could have a strong effect on how easy an instance is.
-    \item The mean and maximum degrees, and whether or not every vertex has the same degree. (The
+    \item the number of vertices and edges;
+    \item the density---we expect that some kinds of filtering might be expensive and ineffective on
+        dense graphs;
+    \item how many loops (self-adjacent vertices) the graph contains---since loops must be mapped to
+        loops, this could have a strong effect on how easy an instance is;
+    \item the mean and maximum degrees, and whether or not every vertex has the same degree (the
         degree-based invariants used by \LAD and \Glasgow do nothing at the top of search if every
-        vertex has the same degree.)
-    \item Whether or not the graph is connected.
-    \item The mean and maximum distances between all pairs of vertices (if nearly all vertices are
-        close together, path-based reasoning is likely to be ineffective). We also consider the
-        proportion of vertex pairs which are distance at least 2, 3 and 4 apart.
+        vertex has the same degree);
+    \item whether or not the graph is connected;
+    \item the mean and maximum distances between all pairs of vertices (if nearly all vertices are
+        close together, path-based reasoning is likely to be ineffective) and
+        the proportion of vertex pairs which are at least 2, 3 and 4 apart.
 \end{itemize}
@@
 
-To decide which solver to run on the remaining instances, we use \LLAMA~\cite{kotthoff_llama_2013}.
+We use \LLAMA~\cite{kotthoff_llama_2013} to build our algorithm selection model.
 \LLAMA supports the most common algorithm selection approaches used in the literature. We performed a
-set of preliminary experiments to determine the approach that works best in this case.
+set of preliminary experiments to determine the approach that works best here.
 
@@
 \LLAMA's pairwise regression approach with random forest regression gave the best performance. The
-idea is very similar to the pairwise classification used by Xu et al.~\cite{xu_satzilla_2008}. For
+idea is very similar to the pairwise classification models used by Xu et al.~\cite{xu_satzilla_2008}. For
 each pair of algorithms in our portfolio, we train a model that predicts the performance difference
-between them. That is, if the first algorithm is better than the second, the difference is positive,
+between them. If the first algorithm is better than the second, the difference is positive,
 otherwise negative. The algorithm with the highest cumulative performance difference, i.e.\ the most
@@
 \GlasgowOne, and \SI{100}{\ms} (resp. \SI{1000}{\ms} and \SI{10000}{\ms}) for \GlasgowTwo (resp.\ 3 and
-4). The Virtual Best Solver (VBS), which considers the best algorithm for each instance separately,
-obtains much better results, showing us that the algorithms have complementary performance. The
+4). The virtual best solver obtains much better results, showing us that the algorithms have complementary performance. The
 difference between VBS and single best (column $\delta$ of Table \ref{expTimeTable}) is more

### 2015-12-15 paper/paper.tex
complete pass

@@
 performance across the entire set of instances, at the CPU time limit of \SI{e8}{\ms},
-i.e.\ \GlasgowTwo. We consider it a lower bound on performance.
+i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
+algorithm selection approach.
 We are able to solve 34 more instances than the single best solver within the timeout, with only an
@@
 virtual best & 0 & 2219 & 5822809\\
-single best & 1965158 & 2172 & 7787967\\
 \LLAMA & 610835 & 2206 & 6435301\\
+\GlasgowTwo & 1965158 & 2172 & 7787967\\
    \bottomrule \\
@@
 instances. MCP is the misclassification penalty; that is, the additional time
-required to solve a set of instances because of choosing solvers that perform
-worse than the best.}\label{tab:res} \end{table}
+required to solve an instance because of choosing solvers that perform
+worse than the best.}\label{tab:res}
+\end{table}
 
-Table~\ref{tab:res} shows the performance of the selection model. The performance of the entire
-system we propose here is shown in Table~\ref{tab:resfull}. The additional instances are solved
-either by \IncompleteLAD or \VFtwo presolving, as explained above. Our system is able to close more than
-\SI{50}{\percent} of the gap between single and virtual best.
+Table~\ref{tab:res} shows the performance of the selection model on its own. The performance of the entire
+algorithm selection system including the preprocessing is shown in
+Table~\ref{tab:resfull}.
+Our system is able to close more than \SI{50}{\percent} of the gap between
+single and virtual best.
 
@@
 virtual best & 0 & 5608 & 2375913\\
-single best & 798660 & 5562 & 3174573\\
 \LLAMA & 249242 & 5595 & 2625831\\
+\GlasgowTwo & 798660 & 5562 & 3174573\\
    \bottomrule \\
@@
 
-Figure~\ref{fig:portfolio-ecdf} shows the cumulative distribution function for
-the individual solvers, the virtual best solver, and the \LLAMA portfolio. The
-virtual best solver clearly dominates with a significant margin to the next
-best. \VFtwo is far below all other solvers. The portfolio does not perform well
-for instances that can be solved quickly because of the overhead incurred
-through feature computation. As the instances become more difficult to solve,
-its performance improves.
+Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time
+for the individual solvers, the virtual best solver, and the \LLAMA algorithm
+selection approach. The performance of \VFtwo is far behind all other solvers.
+The portfolio does not perform well for instances that can be solved quickly
+because of the overhead incurred through feature computation. As the instances
+become more difficult to solve, its performance improves.
 
@@
 \includegraphics[width=\textwidth]{figures/portfolio-ecdf}
-\caption{Empirical cumulative distribution function of solvers, virtual best
-solvers, and the \LLAMA portfolio on the reduced set of 2336 instances. The
+\caption{Number of solved instances over time for solvers, virtual best
+solvers, and \LLAMA on the reduced set of 2336 instances. The
 actual solvers are shown as dashed lines.}
@@
 
-Figure~\ref{fig:portfolio-ecdf-full} shows the ECDF of the system including
+Figure~\ref{fig:portfolio-ecdf-full} shows the number of solved instances over
+time for the algorithm selection system including
 \IncompleteLAD and \VFtwo presolving on the full set of instances. Note that for
@@
 worse than the individual solvers is now limited to approximately $10^2$ to
-$10^4$ milliseconds.
+$10^5$ milliseconds.
 
@@
 \includegraphics[width=\textwidth]{figures/portfolio-ecdf-full}
-\caption{Empirical cumulative distribution function of solvers, virtual best
-solvers, and the portfolio system on the full set of 5725 instances. The
+\caption{Number of solved instances over time for solvers, virtual best
+solvers, and \LLAMA on the full set of 5725 instances. The
 actual solvers are shown as dashed lines.}
@@
 Even though Figures~\ref{fig:portfolio-ecdf} and~\ref{fig:portfolio-ecdf-full}
-show that on easy instances, the portfolio does not perform well, it does over
+show that on easy instances, the model does not perform well, it does over
 the entire set of instances (cf.\ Tables~\ref{tab:res} and~\ref{tab:resfull}).
@@
 particular timeout we have chosen. Table~\ref{tab:resTimeouts} shows the
-performance of our portfolio for timeouts ranging from $10^2$ to $10^8$. We
-do not include \IncompleteLAD and \VFtwo presolving here, as these components do not
-change for different timeouts.
+performance of our algorithm selection approach for timeouts ranging from $10^2$
+to $10^8$. We do not include the instances solved by \IncompleteLAD and \VFtwo
+presolving here, as these components do not change for different timeouts.
 
@@
  & \LLAMA & 3 & 685 & 78\\
- & \Glasgow1 & 11 & 903 & 71\\
+ & \GlasgowOne & 11 & 903 & 71\\
   \midrule
@@
  & \LLAMA & 22 & 1284 & 569\\
- & \Glasgow1 & 126 & 1539 & 449\\
+ & \GlasgowOne & 126 & 1539 & 449\\
   \midrule
@@
  & \LLAMA & 186 & 1771 & 3342\\
- & \Glasgow2 & 683 & 1855 & 2701\\
+ & \GlasgowTwo & 683 & 1855 & 2701\\
   \midrule
@@
  & \LLAMA & 1346 & 2022 & 17231\\
- & \Glasgow2 & 3019 & 1988 & 17368\\
+ & \GlasgowTwo & 3019 & 1988 & 17368\\
   \midrule
@@
  & \LLAMA & 10826 & 2087 & 120459\\
- & Glasgow2 & 22729 & 2063 & 130789\\
+ & \GlasgowTwo & 22729 & 2063 & 130789\\
   \midrule
@@
  & \LLAMA & 91012 & 2156 & 908200\\
- & Glasgow2 & 198944 & 2129 & 1014522\\
+ & \GlasgowTwo & 198944 & 2129 & 1014522\\
   \midrule
@@
  & \LLAMA & 740765 & 2204 & 6565232\\
- & Glasgow2 & 1969418 & 2172 & 7792227\\
+ & \GlasgowTwo & 1969418 & 2172 & 7792227\\
 \bottomrule

### 2015-12-15 paper/paper.tex
more consistent terminology

@@
 characteristics. As with many other hard problems, the single best choice of algorithm overall is
-rarely the best algorithm on an instance-by-instance. We develop a portfolio approach which leverages
+rarely the best algorithm on an instance-by-instance. We develop an algorithm
+selection approach which leverages
 novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm
@@
 selection approach. The performance of \VFtwo is far behind all other solvers.
-The portfolio does not perform well for instances that can be solved quickly
-because of the overhead incurred through feature computation. As the instances
-become more difficult to solve, its performance improves.
+The algorithm selection model does not perform well for instances that can be
+solved quickly because of the overhead incurred through feature computation. As
+the instances become more difficult to solve, its performance improves.
 
@@
 instances is much better than the \LLAMA selector alone (cf.\
-Figure~\ref{fig:portfolio-ecdf}) and the region where the portfolio performs
+Figure~\ref{fig:portfolio-ecdf}) and the region where \LLAMA performs
 worse than the individual solvers is now limited to approximately $10^2$ to
@@
 portfolio of six algorithms from the literature and a novel variant of the \LAD
-algorithm. We introduced a set of novel features to characterise subgraph
-isomorphism problems and leveraged them to select the most appropriate algorithm
-from the portfolio for each instance.
+algorithm. We introduced a
+set of novel features to characterise subgraph isomorphism problems and
+leveraged them to select the most appropriate algorithm from the portfolio for
+each instance.
 
-We demonstrated that our portfolio achieves substantial performance improvements
-over the single algorithm that has the best performance on our benchmark set. We
-showed that combining a portfolio approach with a novel incomplete variant of
-\LAD that is able to detect inconsistencies and a presolver boosts performance
-even further.
+We demonstrated that our algorithm selection approach achieves substantial
+performance improvements over the single algorithm that has the best performance
+on our benchmark set. We showed that combining an algorithm selection approach
+with a novel incomplete variant of \LAD that is able to detect inconsistencies
+and a presolver boosts performance even further.
 

### 2015-12-16 paper/paper.tex
mention that incomplete LAD isn't included in portfolio
Gnuplotify cumulative plot
We shall kill 6.1. With fire! And blood!

@@
 \usepackage{enumitem}
+\usepackage{tikz}
+\usepackage{gnuplot-lua-tikz}
 \setlist{nosep}
+
+\newcommand{\VFtwoNS}{$\textsc{Vf{\scalefont{0.8}2}}$}
+\newcommand{\GlasgowNS}{$\textsc{Glasgow}$}
+\newcommand{\LADNS}{$\textsc{Lad}$}
+\newcommand{\IncompleteLADNS}{$\textsc{IncompleteLad}$}
+\newcommand{\PathLADNS}{$\textsc{PathLad}$}
+\newcommand{\GlasgowOneNS}{$\textsc{Glasgow{\scalefont{0.8}1}}$}
+\newcommand{\GlasgowTwoNS}{$\textsc{Glasgow{\scalefont{0.8}2}}$}
+\newcommand{\GlasgowThreeNS}{$\textsc{Glasgow{\scalefont{0.8}3}}$}
+\newcommand{\GlasgowFourNS}{$\textsc{Glasgow{\scalefont{0.8}4}}$}
+\newcommand{\LLAMANS}{$\textsc{Llama}$}
 
@@
 \begin{figure}[t]
-\includegraphics[width=\textwidth]{courbe.pdf}
+    \centering
+    \input{graph-cumulative}\vspace{-0.5cm}
+
     \caption{Number of solved instances over CPU time for the eight algorithms we
@@
 the overhead of running algorithm selection and avoids potentially making incorrect solver choices.
+
+We include \VFtwo in the portfolio, as it may solve an instance given more time,
+but not \IncompleteLAD.
 
@@
 
-\subsection{Portfolios for Different Timeouts}
-
-The question arises of whether the success of our approach is limited to the
-particular timeout we have chosen. Table~\ref{tab:resTimeouts} shows the
-performance of our algorithm selection approach for timeouts ranging from $10^2$
-to $10^8$. We do not include the instances solved by \IncompleteLAD and \VFtwo
-presolving here, as these components do not change for different timeouts.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{llrrr}
-  \toprule
-timeout & solver & mean MCP & solved instances & mean performance\\
-  \midrule
-$10^2$ & virtual best & 0 & 1253 & 59\\
- & \LLAMA & 3 & 685 & 78\\
- & \GlasgowOne & 11 & 903 & 71\\
-  \midrule
-$10^3$ & virtual best & 0 & 1781 & 322\\
- & \LLAMA & 22 & 1284 & 569\\
- & \GlasgowOne & 126 & 1539 & 449\\
-  \midrule
-$10^4$ & virtual best & 0 & 1943 & 2018\\
- & \LLAMA & 186 & 1771 & 3342\\
- & \GlasgowTwo & 683 & 1855 & 2701\\
-  \midrule
-$10^5$ & virtual best & 0 & 2045 & 14348\\
- & \LLAMA & 1346 & 2022 & 17231\\
- & \GlasgowTwo & 3019 & 1988 & 17368\\
-  \midrule
-$10^6$ & virtual best & 0 & 2111 & 108060\\
- & \LLAMA & 10826 & 2087 & 120459\\
- & \GlasgowTwo & 22729 & 2063 & 130789\\
-  \midrule
-$10^7$ & virtual best & 0 & 2178 & 815577\\
- & \LLAMA & 91012 & 2156 & 908200\\
- & \GlasgowTwo & 198944 & 2129 & 1014522\\
-  \midrule
-$10^8$ & virtual best & 0 & 2219 & 5822809\\
- & \LLAMA & 740765 & 2204 & 6565232\\
- & \GlasgowTwo & 1969418 & 2172 & 7792227\\
-\bottomrule
-\end{tabular}
-\vspace{1ex}
-\caption{Algorithm selection performance for different timeouts.}\label{tab:resTimeouts}
-\end{table}
-
-The results show that while for the misclassification penalty, \LLAMA is always
-better than the single best algorithm, for timeouts $10^2$ to $10^4$ it is worse
-in terms of the number of solved instances and mean performance. The
-misclassification penalty does not take the cost of computing features into
-account, while the mean performance does---for these small timeouts, the cost
-of computing the features dominates any improvement through algorithm selection.
-Another explanation for this is that empirical runtime measurements are
-inherently noisy, which affects the accuracy of the measurements and hence the
-performance of the algorithm selection models particularly in this range.
-
-\begin{table}[t]
-\begin{center}
-\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
-\hline
-$t$ & \VFtwo & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{\Glasgow}& VBS & $\delta$\\
-&&incomplete&classic&path&1&2&3&4&&\\\hline
-$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
-$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
-$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
-$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
-$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
-$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
-$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
-$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
-$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
-\end{tabular}
-\end{center}
-\caption{Number of solved instances at different CPU time limits: Each line displays a time limit
-$t$ (in milliseconds) followed by the number of instances solved within this time limit by \VFtwo,
-\IncompleteLAD, classical \LAD, and \PathLAD, \Glasgow with the lengths of paths limited to 1, 2, 3 and
-4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
-single best if displayed in the last column ($\delta$).\label{expTimeTable}}
-\end{table}
-
-Table~\ref{expTimeTable} displays the number of instances solved at different CPU time limits,
-ranging from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time
-limit considered. Simple approaches like \IncompleteLAD and \VFtwo are able to solve easy instances
-very quickly, in a few milliseconds. However, they are not able to solve harder instances:
-\IncompleteLAD is an incomplete approach which can only detect rather trivial inconsistencies; \VFtwo
-performs a basic backtracking search and is very fast on easy instances, because it does not compute
-expensive invariants. Classical \LAD, \PathLAD, and the \Glasgow algorithms solve more instances than \VFtwo
-and \IncompleteLAD when considering longer time limits: \SI{10}{\ms} for \LAD and \PathLAD and
-\GlasgowOne, and \SI{100}{\ms} (resp. \SI{1000}{\ms} and \SI{10000}{\ms}) for \GlasgowTwo (resp.\ 3 and
-4). The virtual best solver obtains much better results, showing us that the algorithms have complementary performance. The
-difference between VBS and single best (column $\delta$ of Table \ref{expTimeTable}) is more
-particularly important for rather small CPU time limits. In particular, VBS solves 693 more
-instances than the single best (\GlasgowOne) when the limit is \SI{100}{\ms}. In many applications,
-it is important to have the fastest possible algorithm. For example, in pattern recognition
-applications, we often have to solve subgraph isomorphism problems for a very large number of graphs
-(in order to find a pattern image in a large database of target images, for example), so that having
-an algorithm that is able to solve an instance in \SI{100}{\ms} instead of \SI{1000}{\ms} makes a
-big difference. Therefore, it is important to select the best algorithm for each instance.
-
-Table~\ref{selectionvsvbs} shows the performance of our algorithm selection approach, compared to
-the VBS (which is the upper bound of what an algorithm selection approach can achieve), and the
-single best solver, at different CPU time limits (the single best corresponds to results highlighted
-in blue in Table 1).  Comment Table 3: Mean runtime but also number of solved instances at different
-CPU time limits.  Logically, \LLAMA may solve less instances than single best for small time limits
-because it spends time to compute features and choose a solver, but it should be better for larger
-time limits...
-
-\begin{table}
-\begin{tabular}{|l|r|rrrrrrrrr|}
-&mean & \multicolumn{9}{c|}{Number of solved instances}\\
-&runtime & $10^0$ &  $10^1$ &  $10^2$ &  $10^3$ &  $10^4$ &  $10^5$ &  $10^6$ &  $10^7$ &  $10^8$\\\hline
-VBS & 2375.9 & 1132 & 3388 & 4636 & 5170 & 5332 & 5434 & 5500 & 5567 & 5608\\\hline
-Single best & 3177.6 & 1126 & 2287 & 3943 & 4908 & 5253 & 5377 & 5452 & 5518 & 5561\\\hline
-\LLAMA & 2706.3\\\hline
-\end{tabular}
-\caption{}\label{selectionvsvbs}
-\end{table}
-
-Add Figure 1 which compares single best at \SI{e8}{\ms}(\GlasgowTwo) with \LLAMA ?
-
 \section{Conclusion and Future Work}\label{sec:concs}

### 2015-12-16 paper/paper.tex
Muck around with float placement
More graphery. Will improve these later.

@@
 
-\begin{figure}[t]
+\begin{figure}[p]
     \centering
@@
 \end{figure}
+
+\begin{table}[p]
+\centering
+\begin{tabularx}{.8\textwidth}{XXrXrrrXrrrr}
+\toprule
+Class && \VFtwo && \multicolumn{3}{c}{\LAD} && \multicolumn{4}{c}{\hspace*{2em}\Glasgow}\\
+&&&&incomplete&default&path&&1&2&3&4\\
+\midrule
+1 &&        0 &&       20 &        0 &        0 &&       80 &        0 &        0 &        0 \\
+2 &&      201 &&       92 &      189 &      270 &&      520 &      180 &       53 &       15 \\
+3 &&      112 &&     1608 &      617 &      959 &&      396 &      195 &       21 &        0 \\
+4 &&      270 &&        0 &        0 &        0 &&        5 &        0 &        0 &        0 \\
+5 &&      266 &&        0 &        1 &        3 &&       31 &        0 &        0 &        0 \\
+6 &&       71 &&        0 &        0 &        0 &&        7 &       14 &        1 &        0 \\
+7 &&      270 &&        0 &        0 &        0 &&        5 &        0 &        0 &        0 \\
+8 &&        0 &&        0 &        0 &        1 &&      195 &       69 &        6 &        0 \\
+9 &&       77 &&        3 &        0 &       19 &&      103 &        1 &        0 &        0 \\
+10 &&       13 &&        0 &        2 &        2 &&        7 &        0 &        0 &        0 \\
+11 &&        2 &&      142 &       71 &       17 &&       23 &        0 &        0 &        0 \\
+12 &&        0 &&        0 &        1 &        2 &&      158 &        6 &        1 &        0 \\
+\midrule
+Total && 1282 && 1865 & 881 & 1273 && 1530 & 465 & 82& 15\\
+\bottomrule \\
+\end{tabularx}
+\caption{Number of times each algorithm is best, for each class.\label{expClass}}
+\end{table}
 
@@
 (except \GlasgowFour for class 3).
-
-\begin{table}[t]
-\begin{center}
-\begin{tabularx}{.8\textwidth}{XXrXrrrXrrrr}
-\toprule
-Class && \VFtwo && \multicolumn{3}{c}{\LAD} && \multicolumn{4}{c}{\hspace*{2em}\Glasgow}\\
-&&&&incomplete&default&path&&1&2&3&4\\
-\midrule
-1 &&        0 &&       20 &        0 &        0 &&       80 &        0 &        0 &        0 \\
-2 &&      201 &&       92 &      189 &      270 &&      520 &      180 &       53 &       15 \\
-3 &&      112 &&     1608 &      617 &      959 &&      396 &      195 &       21 &        0 \\
-4 &&      270 &&        0 &        0 &        0 &&        5 &        0 &        0 &        0 \\
-5 &&      266 &&        0 &        1 &        3 &&       31 &        0 &        0 &        0 \\
-6 &&       71 &&        0 &        0 &        0 &&        7 &       14 &        1 &        0 \\
-7 &&      270 &&        0 &        0 &        0 &&        5 &        0 &        0 &        0 \\
-8 &&        0 &&        0 &        0 &        1 &&      195 &       69 &        6 &        0 \\
-9 &&       77 &&        3 &        0 &       19 &&      103 &        1 &        0 &        0 \\
-10 &&       13 &&        0 &        2 &        2 &&        7 &        0 &        0 &        0 \\
-11 &&        2 &&      142 &       71 &       17 &&       23 &        0 &        0 &        0 \\
-12 &&        0 &&        0 &        1 &        2 &&      158 &        6 &        1 &        0 \\
-\midrule
-Total && 1282 && 1865 & 881 & 1273 && 1530 & 465 & 82& 15\\
-\bottomrule
-\end{tabularx}
-\end{center}
-\caption{Number of times each algorithm is best, for each class.\label{expClass}}
-\end{table}
 
@@
 
-\begin{table}[ht]
+Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time
+for the individual solvers, the virtual best solver, and the \LLAMA algorithm
+selection approach. The performance of \VFtwo is far behind all other solvers.
+The algorithm selection model does not perform well for instances that can be
+solved quickly because of the overhead incurred through feature computation. As
+the instances become more difficult to solve, its performance improves.
+
+\begin{table}[p]
     \centering\setlength{\tabcolsep}{1em}
@@
 
+\begin{figure}[p]
+\input{graph-reduced}
+
+\caption{Number of solved instances over time for the virtual best
+solver, \LLAMA, and individual solvers (shown as dotted lines, in the same
+colours as Figure~\ref{expTimeGraph}) on the reduced set of 2336 instances.}
+\label{fig:portfolio-ecdf}
+\end{figure}
+
 Table~\ref{tab:res} shows the performance of the selection model on its own. The performance of the entire
@@
 
-\begin{table}[ht]
+Figure~\ref{fig:portfolio-ecdf-full} shows the number of solved instances over
+time for the algorithm selection system including
+\IncompleteLAD and \VFtwo presolving on the full set of instances. Note that for
+very easy instances, we are better than the virtual best solver because
+\IncompleteLAD is not included in our portfolio. The performance on small
+instances is much better than the \LLAMA selector alone (cf.\
+Figure~\ref{fig:portfolio-ecdf}) and the region where \LLAMA performs
+worse than the individual solvers is now limited to approximately $10^2$ to
+$10^5$ milliseconds.
+
+\begin{table}[p]
     \centering\setlength{\tabcolsep}{1em}
@@
 
-Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time
-for the individual solvers, the virtual best solver, and the \LLAMA algorithm
-selection approach. The performance of \VFtwo is far behind all other solvers.
-The algorithm selection model does not perform well for instances that can be
-solved quickly because of the overhead incurred through feature computation. As
-the instances become more difficult to solve, its performance improves.
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
-\caption{Number of solved instances over time for solvers, virtual best
-solvers, and \LLAMA on the reduced set of 2336 instances. The
-actual solvers are shown as dashed lines.}
-\label{fig:portfolio-ecdf}
-\end{figure}
-
-Figure~\ref{fig:portfolio-ecdf-full} shows the number of solved instances over
-time for the algorithm selection system including
-\IncompleteLAD and \VFtwo presolving on the full set of instances. Note that for
-very easy instances, we are better than the virtual best solver because
-\IncompleteLAD is not included in our portfolio. The performance on small
-instances is much better than the \LLAMA selector alone (cf.\
-Figure~\ref{fig:portfolio-ecdf}) and the region where \LLAMA performs
-worse than the individual solvers is now limited to approximately $10^2$ to
-$10^5$ milliseconds.
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figures/portfolio-ecdf-full}
+\begin{figure}[p]
+    \input{graph-full}
 \caption{Number of solved instances over time for solvers, virtual best

### 2015-12-16 paper/paper.tex
A bit of consistency in cite styles

@@
 We formally define the subgraph isomorphism problem in Section~\ref{sec:defs}.
-In Section~\ref{sec:algs}, we describe the main existing algorithms for solving this problem, and we also introduce two new algorithms which are derived from the LAD algorithm of~\cite{Solnon:2010}.
+In Section~\ref{sec:algs}, we describe the main existing algorithms for solving
+this problem, and we also introduce two new algorithms which are derived from
+Solnon's \LAD algorithm \cite{Solnon:2010}.
 
@@
 composed of twelve classes. Ten of these classes have been introduced in the
-experimental study reported in~\cite{McCreesh:2015}. We evaluate the algorithms,
-of which two are novel, on this benchmark set, and show that they have very
-complementary performance. In particular, we show that depending on the CPU time
-limit, different algorithms achieve the best performance on the entire benchmark
-set.
+experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}. We
+evaluate the algorithms, of which two are novel, on this benchmark set, and
+show that they have very complementary performance. In particular, we show that
+depending on the CPU time limit, different algorithms achieve the best
+performance on the entire benchmark set.
 
-In Section~\ref{sec:algsel}, we describe the features that are used to describe instances, and we describe our algorithm selection approach.
-It combines a presolving step, which allows us to very quickly solve easy instances, with the algorithm selection approach \LLAMA~\cite{kotthoff_llama_2013}.
+In Section~\ref{sec:algsel}, we describe the features that are used to describe
+instances, and we describe our algorithm selection approach. It combines a
+presolving step, which allows us to very quickly solve easy instances, with the
+algorithm selection approach \LLAMA~\cite{kotthoff_llama_2013}.
 

### 2015-12-16 paper/paper.tex
Christine's changes

@@
 algorithms. We introduce a large benchmark set composed of 5725 instances
-composed of twelve classes. Ten of these classes have been introduced in the
+grouped into twelve classes. Ten of these classes have been introduced in the
 experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}. We
-evaluate the algorithms, of which two are novel, on this benchmark set, and
-show that they have very complementary performance. In particular, we show that
-depending on the CPU time limit, different algorithms achieve the best
-performance on the entire benchmark set.
+evaluate the algorithms on this benchmark set, and show that they have very
+complementary performance. In particular, we show that depending on the CPU
+time limit, different algorithms achieve the best performance on the entire
+benchmark set.
 
@@
 Note that the subgraph is not necessarily induced, so that two pattern nodes not linked by
-an edge may be mapped to two target nodes which are linked by an edge. This problem is also called
-\emph{subgraph monomorphism} or \emph{subgraph matching} in the literature.
+an edge may be mapped to two target nodes which are linked by an edge.
 
@@
 
-The simplest form of filtering is simply to propagate difference constraints (which ensure that the
+The simplest form of filtering is to propagate difference constraints (which ensure that the
 matching is injective) and edge constraints (which ensure that the matching preserves pattern
@@
 difference constraint is propagated separately which Forward-Checking. However, it is also more time
-consuming as requires ${\cal O}(n_p^2\cdot n_t^2)$ time.
+consuming as it requires ${\cal O}(n_p^2\cdot n_t^2)$ time.
 
@@
 \begin{itemize}
-\item \VFtwo{}~\cite{Cordella:2004} performs weak filtering that it is especially fast on
+\item \VFtwo{}~\cite{Cordella:2004} performs weak filtering that is especially fast on
     trivially satisfiable instances;
@@
 \item Class 1 contains randomly generated scale-free graphs~\cite{constraints10}.
-\item Classes 2 and 3 contain various kinds of graphs selected by Larrosa and Valiente~\cite{LV02}:
-    class 2 contains small instances generated from the 50 first graphs of class
-    1 and class 3 contains
-    larger instances with pattern graphs from class 2 and target graphs from
-    the next 50 graphs.
-\item Classes 4 to 8 contain randomly generated graphs from a database of graphs commonly
-    used for benchmarking subgraph isomorphism algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 4 and 5, regular meshes
-    for classes 6 and 7, and random graphs with uniform edge probabilities for class 8. All of these
-    instances are satisfiable.
+\item Classes 2 and 3 contain instances built from a database containing
+    various kinds of graph gathered by Larrosa and Valiente~\cite{LV02}: class
+    2 contains small instances generated from the first 50 graphs of the
+    database, and class 3 contains larger instances with pattern graphs from
+    the first 50 graphs of the database and target graphs from the next 50
+    graphs.
+\item Classes 4 to 8 contain randomly generated graphs from a database of
+    graphs commonly used for benchmarking subgraph isomorphism
+    algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for
+    classes 4 and 5, regular meshes for classes 6 and 7, and random graphs with
+    uniform edge probabilities for class 8. All of these instances are
+    satisfiable.
 \item Classes 9 and 10 contain instances from segmented images~\cite{pr15,cviu11}.
@@
 makes a big difference.  Therefore, it is important to select the best algorithm
-for each instance, even if the instance is an easy one.
+for each instance, even if the instance is an easy one. Furthermore, this
+selection process should not penalise easy instances, i.e.\ it should not take more time than
+the solution process time for these instances.
 
@@
 \IncompleteLAD: if unsatisfiability is detected, we do not need to process it further.
-Otherwise, we record the number of candidate pairs removed and as a percentage
-of all pairs and the minimum and maximum values on a per-variable basis. The
-recorded values are then used as features for algorithm selection.
 
@@
 
-If presolving does not give us a solution, we extract additional features. For both the pattern and
+If presolving does not give us a solution, we extract features. For both the pattern and
 the target graph, we consider some basic graph properties, which can be computed
-very quickly,
+very quickly:
+
 \begin{itemize}
     \item the number of vertices and edges;
-    \item the density---we expect that some kinds of filtering might be expensive and ineffective on
+    \item the density---we expect that some kinds of filtering (like those based upon locally
+        all-different constraints) might be expensive and ineffective on
         dense graphs;
@@
 \end{itemize}
+
+Alongside these basic features, we include information computed by \IncompleteLAD. To (try to) prove
+inconsistency, \IncompleteLAD removes candidate pairs. The number of successfully removed pairs
+gives information on the distribution of edges (the fewer removed pairs, the more uniform the
+distribution). As well as the number of removed pairs, we also record the percentage with respect to
+all possible pairs, and the minimum and maximum values on a per-variable basis.
 
@@
 Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
-baselines. The virtual best solver is the oracle predictor that, for each instance, chooses the best
-solver from our portfolio. This is the upper bound of what an algorithm selection approach can
-achieve. The single best solver is the one solver from the portfolio that has the overall best
-performance across the entire set of instances, at the CPU time limit of \SI{e8}{\ms},
-i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
-algorithm selection approach.
-We are able to solve 34 more instances than the single best solver within the timeout, with only an
-additional 13 to the virtual best. In terms of average performance, we are able to close almost
-\SI{70}{\percent} of the gap between the single best and the virtual best solver.
+baselines, on the reduced set of 2336 instances. The virtual best solver is the oracle predictor
+that, for each instance, chooses the best solver from our portfolio. This is the upper bound of what
+an algorithm selection approach can achieve. The single best solver is the one solver from the
+portfolio that has the overall best performance across the entire set of instances, at the CPU time
+limit of \SI{e8}{\ms}, i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
+algorithm selection approach.  We are able to solve 34 more instances than the single best solver
+within the timeout, with only an additional 13 to the virtual best. In terms of average performance,
+we are able to close almost \SI{70}{\percent} of the gap between the single best and the virtual
+best solver.
 
-Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time
-for the individual solvers, the virtual best solver, and the \LLAMA algorithm
-selection approach. The performance of \VFtwo is far behind all other solvers.
-The algorithm selection model does not perform well for instances that can be
-solved quickly because of the overhead incurred through feature computation. As
-the instances become more difficult to solve, its performance improves.
+Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time for the individual
+solvers, the virtual best solver, and the \LLAMA algorithm selection approach. The performance of
+\VFtwo is far behind all other solvers.  The algorithm selection model does not perform well for
+instances that can be solved quickly because of the overhead incurred through feature computation.
+As the instances become more difficult to solve, its performance improves.
 
@@
 \end{tabular}
-\caption{Algorithm selection performance on the reduced set of 2336
-instances. MCP is the misclassification penalty; that is, the additional time
-required to solve an instance because of choosing solvers that perform
-worse than the best.}\label{tab:res}
+\caption{Algorithm selection performance on the reduced set of 2336 instances. MCP is the
+misclassification penalty; that is, the additional time required to solve an instance because of
+choosing solvers that perform worse than the best. Mean performance is over all instances; when an
+instance is not solved, its performance is set to the time limit
+(\SI{e8}{\ms}).}\label{tab:res}
 \end{table}
@@
 
-Table~\ref{tab:res} shows the performance of the selection model on its own. The performance of the entire
-algorithm selection system including the preprocessing is shown in
-Table~\ref{tab:resfull}.
-Our system is able to close more than \SI{50}{\percent} of the gap between
-single and virtual best.
+Table~\ref{tab:res} shows the performance of the selection model on its own. The performance of the
+entire algorithm selection system, including the preprocessing, is shown in Table~\ref{tab:resfull}.
+Our system is able to close more than \SI{50}{\percent} of the gap between single and virtual best.
 
-Figure~\ref{fig:portfolio-ecdf-full} shows the number of solved instances over
-time for the algorithm selection system including
-\IncompleteLAD and \VFtwo presolving on the full set of instances. Note that for
-very easy instances, we are better than the virtual best solver because
-\IncompleteLAD is not included in our portfolio. The performance on small
-instances is much better than the \LLAMA selector alone (cf.\
-Figure~\ref{fig:portfolio-ecdf}) and the region where \LLAMA performs
-worse than the individual solvers is now limited to approximately $10^2$ to
-$10^5$ milliseconds.
+Figure~\ref{fig:portfolio-ecdf-full} shows the number of solved instances over time for the
+algorithm selection system including \IncompleteLAD and \VFtwo presolving on the full set of
+instances. Note that for very easy instances, we are better than the virtual best solver because
+\IncompleteLAD is not included in our portfolio. The performance on small instances is much better
+than the \LLAMA selector alone (cf.\ Figure~\ref{fig:portfolio-ecdf}) and the region where \LLAMA
+performs worse than the individual solvers is now limited to approximately $10^2$ to $10^5$
+milliseconds.
 
@@
     \input{graph-full}
-\caption{Number of solved instances over time for solvers, virtual best
-solvers, and \LLAMA on the full set of 5725 instances. The
-actual solvers are shown as dashed lines.}
+\caption{Number of solved instances over time for the virtual best
+solver, \LLAMA, and individual solvers (shown as dotted lines, in the same
+colours as Figure~\ref{expTimeGraph}) on the full set of 5725 instances.}
 \label{fig:portfolio-ecdf-full}
@@
 
-The problem of identifying subgraph isomorphisms is a hard computation problem
-that has many applications in diverse areas. In this paper, we presented a
-portfolio of six algorithms from the literature and a novel variant of the \LAD
-algorithm. We introduced a
-set of novel features to characterise subgraph isomorphism problems and
-leveraged them to select the most appropriate algorithm from the portfolio for
-each instance.
+The problem of identifying subgraph isomorphisms is a hard computation problem that has many
+applications in diverse areas. In this paper, we presented a portfolio of six algorithms from the
+literature and two new variants of the \LAD algorithm. We introduced a set of novel features to
+characterise subgraph isomorphism problems and leveraged them to select the most appropriate
+algorithm from the portfolio for each instance.
 
@@
 on our benchmark set. We showed that combining an algorithm selection approach
-with a novel incomplete variant of \LAD that is able to detect inconsistencies
+with a new incomplete variant of \LAD that is able to detect inconsistencies
 and a presolver boosts performance even further.
-
-Mention parallel, directed, labelled in here briefly.
 

### 2015-12-16 paper/paper.tex
Muck around with list formatting

@@
 \usepackage{gnuplot-lua-tikz}
-\setlist{nosep}
 
@@
 In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
-approach, and show that it is able to close almost 70\% of the gap between the
-single best and the virtual best solver.
+approach, and show that it is able to close more than 50\% of the gap between
+the single best and the virtual best solver.
 
@@
 
-Audemard et al.~\cite{Audemard:2014} do not limit the length of paths considered, and
+\noindent Audemard et al.~\cite{Audemard:2014} do not limit the length of paths considered, and
 iteratively increment the length until no more pairs are removed. Battiti and Mascia~\cite{battiti-mascia07}, and McCreesh and Prosser~\cite{McCreesh:2015} parameterise their algorithms
@@
 
-Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
+\noindent Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
 Prosser~\cite{McCreesh:2015}.  This set of instances is also much larger than that of Battiti and
@@
 
-Alongside these basic features, we include information computed by \IncompleteLAD. To (try to) prove
+\noindent Alongside these basic features, we include information computed by \IncompleteLAD. To (try to) prove
 inconsistency, \IncompleteLAD removes candidate pairs. The number of successfully removed pairs
@@
 limit of \SI{e8}{\ms}, i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
-algorithm selection approach.  We are able to solve 34 more instances than the single best solver
-within the timeout, with only an additional 13 to the virtual best. In terms of average performance,
-we are able to close almost \SI{70}{\percent} of the gap between the single best and the virtual
+algorithm selection approach.  We are able to solve 28 more instances than the single best solver
+within the timeout, with only an additional 18 to the virtual best. In terms of average performance,
+we are able to more than \SI{50}{\percent} of the gap between the single best and the virtual
 best solver.
@@
 virtual best & 0 & 2219 & 5822809\\
-\LLAMA & 610835 & 2206 & 6435301\\
-\GlasgowTwo & 1965158 & 2172 & 7787967\\
+\LLAMA & 902316 & 2201 & 6726782\\
+\GlasgowTwo & 1960992 & 2173 & 7783801\\
    \bottomrule \\
@@
 entire algorithm selection system, including the preprocessing, is shown in Table~\ref{tab:resfull}.
-Our system is able to close more than \SI{50}{\percent} of the gap between single and virtual best.
+Our system is able to close more than \SI{50}{\percent} of the gap between
+single and virtual best, similar to the results on the reduced set of instances.
 
@@
 virtual best & 0 & 5608 & 2375913\\
-\LLAMA & 249242 & 5595 & 2625831\\
+\LLAMA & 368176 & 5590 & 2744765\\
 \GlasgowTwo & 798660 & 5562 & 3174573\\

### 2015-12-16 paper/paper.tex
Weasel word

@@
 for each instance, even if the instance is an easy one. Furthermore, this
-selection process should not penalise easy instances, i.e.\ it should not take more time than
+selection process should not unduly penalise easy instances, i.e.\ it should not take more time than
 the solution process time for these instances.

### 2015-12-16 portfolio/as-graphs.Rnw
some changes to the portfolio stuff

--- 2015-12-09 portfolio/as-graphs.Rnw
@@
 library(tidyr)
+library(ggthemes)
+library(grid)
+library(scales)
 
@@
 wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
-ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+write.csv(wide, file = "all.csv", row.names = FALSE)
+
+p.full = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
     scale_linetype_manual(values=c(3,1), guide = FALSE) +
-    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
-    ylab("fraction of instances solved") +
-    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    guides(col = guide_legend(ncol = 2, keyheight = .8)) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.5), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+
+p.zoom = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 3),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(asscenario$desc$algorithm_cutoff_time/10, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(.95,1)) +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    theme(legend.position="none",
+        axis.title.x=element_blank(), axis.title.y=element_blank(),
+        panel.background = element_rect(fill='white', colour = "white"),
+        axis.line = element_line(colour="black"),
+        panel.grid = element_line(),
+        panel.grid.major = element_line(colour="lightgray"))
+vp = viewport(width = 0.5, height = 0.3, x = 0.7, y = 0.42)
+print(p.full)
+print(p.zoom, vp = vp)
 @
@@
 wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
-ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+write.csv(wide, file = "no-presolved.csv", row.names = FALSE)
+p.full = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
     scale_linetype_manual(values=c(3,1), guide = FALSE) +
-    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
-    ylab("fraction of instances solved") +
-    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    guides(col = guide_legend(ncol = 2, keyheight = .8)) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.5), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+
+p.zoom = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 3),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(asscenario$desc$algorithm_cutoff_time/10, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(.95,1)) +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    theme(legend.position="none",
+        axis.title.x=element_blank(), axis.title.y=element_blank(),
+        panel.background = element_rect(fill='white', colour = "white"),
+        axis.line = element_line(colour="black"),
+        panel.grid = element_line(),
+        panel.grid.major = element_line(colour="lightgray"))
+vp = viewport(width = 0.5, height = 0.3, x = 0.7, y = 0.42)
+print(p.full)
+print(p.zoom, vp = vp)
 @
@@
 wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
-ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+write.csv(wide, file = "no-presolved-allinstances.csv", row.names = FALSE)
+p.full = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
     scale_linetype_manual(values=c(3,1), guide = FALSE) +
-    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
-    ylab("fraction of instances solved") +
-    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    guides(col = guide_legend(ncol = 2, keyheight = .8)) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.5), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+
+p.zoom = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 3),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(asscenario$desc$algorithm_cutoff_time/10, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(.95,1)) +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    theme(legend.position="none",
+        axis.title.x=element_blank(), axis.title.y=element_blank(),
+        panel.background = element_rect(fill='white', colour = "white"),
+        axis.line = element_line(colour="black"),
+        panel.grid = element_line(),
+        panel.grid.major = element_line(colour="lightgray"))
+vp = viewport(width = 0.5, height = 0.3, x = 0.7, y = 0.42)
+print(p.full)
+print(p.zoom, vp = vp)
 @
@@
 wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
-ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+write.csv(wide, file = "no-presolved-presolver.csv", row.names = FALSE)
+p.full = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
     scale_linetype_manual(values=c(3,1), guide = FALSE) +
-    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
-    ylab("fraction of instances solved") +
-    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    guides(col = guide_legend(ncol = 2, keyheight = .8)) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.5), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+
+p.zoom = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 3),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(asscenario$desc$algorithm_cutoff_time/10, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(.95,1)) +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    theme(legend.position="none",
+        axis.title.x=element_blank(), axis.title.y=element_blank(),
+        panel.background = element_rect(fill='white', colour = "white"),
+        axis.line = element_line(colour="black"),
+        panel.grid = element_line(),
+        panel.grid.major = element_line(colour="lightgray"))
+vp = viewport(width = 0.5, height = 0.3, x = 0.7, y = 0.42)
+print(p.full)
+print(p.zoom, vp = vp)
 @
@@
 wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
-ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+write.csv(wide, file = "no-presolved-presolver-allinstances.csv", row.names = FALSE)
+p.full = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
     scale_linetype_manual(values=c(3,1), guide = FALSE) +
-    scale_x_log10(limits = c(10e-5, asscenario$desc$algorithm_cutoff_time)) +
-    ylab("fraction of instances solved") +
-    geom_vline(xintercept = asscenario$desc$algorithm_cutoff_time, col = "red")
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(0,1)) +
+    ylab("fraction of instances solved") + xlab("time [ms]") +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    guides(col = guide_legend(ncol = 2, keyheight = .8)) +
+    theme(legend.justification=c(1,0), legend.position=c(1,0.5), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+
+p.zoom = ggplot(wide, aes(x = time, col = solver, linetype = type)) +
+    stat_ecdf() +
+    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 3),
+                  labels = trans_format("log10", math_format(10^.x)),
+                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+    coord_cartesian(xlim = c(asscenario$desc$algorithm_cutoff_time/10, asscenario$desc$algorithm_cutoff_time-1),
+        ylim = c(.95,1)) +
+    annotation_logticks(sides = "b") +
+    theme_tufte(base_family='Times', base_size = 14) +
+    theme(legend.position="none",
+        axis.title.x=element_blank(), axis.title.y=element_blank(),
+        panel.background = element_rect(fill='white', colour = "white"),
+        axis.line = element_line(colour="black"),
+        panel.grid = element_line(),
+        panel.grid.major = element_line(colour="lightgray"))
+vp = viewport(width = 0.5, height = 0.3, x = 0.7, y = 0.42)
+print(p.full)
+print(p.zoom, vp = vp)
 @
+
+\clearpage
+
+%\subsection*{Only Glasgow variants}
+%
+%<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
+%ldf.glasgow = ldf2
+%ldf.glasgow$performance = c("glasgow1", "glasgow2", "glasgow3", "glasgow4")
+%ldf.glasgow$success = c("glasgow1_success", "glasgow2_success", "glasgow3_success", "glasgow4_success")
+%
+%pnames = ldf.glasgow$performance
+%snames = ldf.glasgow$success
+%optfun = min
+%best = apply(ldf.glasgow$data, 1,
+%    function(x) {
+%        tosel = pnames
+%        nosuccs = sapply(snames[which(as.logical(x[snames]) == FALSE)], function(y) { unlist(strsplit(y, "_"))[1] })
+%        tosel = setdiff(pnames, nosuccs)
+%        if(length(tosel) == 0) {
+%            # nothing was able to solve this instance
+%            return(NA)
+%        } else {
+%            perfs = as.numeric(x[tosel])
+%            return(tosel[which(perfs == optfun(perfs))])
+%        }
+%    })
+%# simplify...
+%names(best) = NULL
+%ldf.glasgow$best = best
+%
+%resvbs = data.frame(model = "virtual best",
+%    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, vbs)),
+%    solved = sum(successes(ldf.glasgow, vbs)),
+%    mean.performance = mean(parscores(ldf.glasgow, vbs, factor = 1)),
+%    median.performance = median(parscores(ldf.glasgow, vbs, factor = 1)))
+%ressb = data.frame(model = "single best",
+%    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, singleBest)),
+%    solved = sum(successes(ldf.glasgow, singleBest)),
+%    mean.performance = mean(parscores(ldf.glasgow, singleBest, factor = 1)),
+%    median.performance = median(parscores(ldf.glasgow, singleBest, factor = 1)))
+%
+%modelc = classify(rflearner, ldf.glasgow)
+%resc = data.frame(model = "classification",
+%    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, modelc)),
+%    solved = sum(successes(ldf.glasgow, modelc)),
+%    mean.performance = mean(parscores(ldf.glasgow, modelc, factor = 1)),
+%    median.performance = median(parscores(ldf.glasgow, modelc, factor = 1)))
+%times = sapply(modelc$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
+%print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classification."), booktabs = TRUE)
+%
+%modelr = regression(makeLearner("regr.randomForest"), ldf.glasgow)
+%resr = data.frame(model = "regression",
+%    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, modelr)),
+%    solved = sum(successes(ldf.glasgow, modelr)),
+%    mean.performance = mean(parscores(ldf.glasgow, modelr, factor = 1)),
+%    median.performance = median(parscores(ldf.glasgow, modelr, factor = 1)))
+%times = sapply(modelr$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
+%print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regression."), booktabs = TRUE)
+%
+%modelcp = classifyPairs(rflearner, ldf.glasgow)
+%rescp = data.frame(model = "classifyPairs",
+%    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, modelcp)),
+%    solved = sum(successes(ldf.glasgow, modelcp)),
+%    mean.performance = mean(parscores(ldf.glasgow, modelcp, factor = 1)),
+%    median.performance = median(parscores(ldf.glasgow, modelcp, factor = 1)))
+%times = sapply(modelcp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "MeanDecreaseGini", asc = FALSE))[1])
+%print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for classificationPairs."), booktabs = TRUE)
+%
+%modelrp = regressionPairs(makeLearner("regr.randomForest"), ldf.glasgow)
+%resrp = data.frame(model = "regressionPairs",
+%    mean.misclassification.penalty = mean(misclassificationPenalties(ldf.glasgow, modelrp)),
+%    solved = sum(successes(ldf.glasgow, modelrp)),
+%    mean.performance = mean(parscores(ldf.glasgow, modelrp, factor = 1)),
+%    median.performance = median(parscores(ldf.glasgow, modelrp, factor = 1)))
+%times = sapply(modelrp$models, function(x) row.names(sortByCol(as.data.frame(x$learner.model$importance), "IncNodePurity", asc = FALSE))[1])
+%print(xtable(as.table(sort(table(times), decreasing = TRUE)), caption = "Most important features for regressionPairs."), booktabs = TRUE)
+%
+%tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
+%print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results with presolver for Glasgow solvers."), booktabs = TRUE, include.rownames = FALSE)
+%@
+%
+%<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio (best model).'>>=
+%vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf.glasgow), function(ss) { ss[1] })
+%vbs.agg$virtual.best = parscores(ldf.glasgow, vbs, factor = 1)
+%vbs.agg$iteration = NULL
+%vbs.agg$score = NULL
+%
+%pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf.glasgow, mod, factor = 1))
+%
+%perfs = subset(ldf.glasgow$data, TRUE, c("instance_id", ldf.glasgow$performance))
+%perfs = merge(perfs, vbs.agg, by = "instance_id")
+%perfs = merge(perfs, pmod, by = "instance_id")
+%wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+%wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+%ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+%    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+%    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+%                  labels = trans_format("log10", math_format(10^.x)),
+%                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+%    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+%        ylim = c(0,1)) +
+%    ylab("fraction of instances solved") + xlab("time [ms]") +
+%    annotation_logticks(sides = "b") +
+%    theme_tufte(base_family='sans', base_size = 12) +
+%    theme(legend.justification=c(1,0), legend.position=c(1,0.1), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+%@
+%
+%<<echo=FALSE,fig.cap='ECDF for algorithms and portfolio for all instances (best model).'>>=
+%vbs.agg = aggregate(as.formula(paste("score~", paste(c("instance_id", "iteration"), sep="+", collapse="+"))), vbs(ldf), function(ss) { ss[1] })
+%vbs.agg$virtual.best = parscores(ldf, vbs, factor = 1)
+%vbs.agg$iteration = NULL
+%vbs.agg$score = NULL
+%
+%pmod = data.frame(instance_id = unique(mod$predictions$instance_id), portfolio = parscores(ldf.glasgow, mod, factor = 1))
+%pmod = rbind(pmod, data.frame(instance_id = factor(presolved), portfolio = subset(ldf$data, ldf$data$instance_id %in% presolved)$lad_features))
+%vf2pre = setdiff(ldf3$data$instance_id, ldf.glasgow$data$instance_id)
+%pmod = rbind(pmod, data.frame(instance_id = factor(vf2pre), portfolio = subset(ldf$data, ldf$data$instance_id %in% vf2pre)$vf2))
+%
+%perfs = subset(ldf$data, TRUE, c("instance_id", ldf$performance))
+%perfs = merge(perfs, vbs.agg, by = "instance_id")
+%perfs = merge(perfs, pmod, by = "instance_id")
+%wide = gather_(perfs, "solver", "time", names(perfs)[-1])
+%wide$type = ifelse(wide$solver %in% c("virtual.best", "portfolio"), "pf", "alg")
+%ggplot(wide, aes(x = time, col = solver, linetype = type)) + stat_ecdf() +
+%    scale_linetype_manual(values=c(3,1), guide = FALSE) +
+%    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x, n = 10),
+%                  labels = trans_format("log10", math_format(10^.x)),
+%                  limits = c(1, asscenario$desc$algorithm_cutoff_time-1)) +
+%    coord_cartesian(xlim = c(1, asscenario$desc$algorithm_cutoff_time-1),
+%        ylim = c(0,1)) +
+%    ylab("fraction of instances solved") + xlab("time [ms]") +
+%    annotation_logticks(sides = "b") +
+%    theme_tufte(base_family='sans', base_size = 12) +
+%    theme(legend.justification=c(1,0), legend.position=c(1,0.1), aspect.ratio = 0.6, axis.line = element_line(colour="black"), panel.grid = element_line(), panel.grid.major = element_line(colour="lightgray"))
+%@
+%
+%\clearpage
+
+%\section*{Different timeouts}
+%
+%Consider timeouts $10^2$\ldots$10^8$ (after filtering and presolving).
+%
+%<<echo=FALSE, results='asis', warning=FALSE, message=FALSE>>=
+%timeoutExps = 2:8
+%
+%tab = rbind.fill(lapply(timeoutExps, function(to) {
+%    ldf.tmp = ldf2
+%    if(10^to < asscenario$desc$algorithm_cutoff_time) {
+%        for(p in ldf.tmp$performance) {
+%            ldf.tmp$data[paste0(p, "_success")] = sapply(ldf.tmp$data[,p], function(x) x <= 10^to)
+%            ldf.tmp$data[p] = sapply(ldf.tmp$data[,p], function(x) min(x, 10^to))
+%        }
+%        pnames = ldf.tmp$performance
+%        snames = ldf.tmp$success
+%        optfun = min
+%        best = apply(ldf.tmp$data, 1,
+%            function(x) {
+%                tosel = pnames
+%                nosuccs = sapply(snames[which(as.logical(x[snames]) == FALSE)], function(y) { unlist(strsplit(y, "_"))[1] })
+%                tosel = setdiff(pnames, nosuccs)
+%                if(length(tosel) == 0) {
+%                    # nothing was able to solve this instance
+%                    return(NA)
+%                } else {
+%                    perfs = as.numeric(x[tosel])
+%                    return(tosel[which(perfs == optfun(perfs))])
+%                }
+%            })
+%        names(best) = NULL
+%        ldf.tmp$best = best
+%    }
+%    ldf.tmp$train = NULL
+%    ldf.tmp$test = NULL
+%    ldf.tmp = cvFolds(ldf.tmp)
+%
+%    resvbs = data.frame(model = "virtual best",
+%        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, vbs)),
+%        solved = sum(successes(ldf.tmp, vbs)),
+%        mean.performance = mean(parscores(ldf.tmp, vbs, factor = 1)),
+%        median.performance = median(parscores(ldf.tmp, vbs, factor = 1)))
+%    ressb = data.frame(model = "single best",
+%        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, singleBest)),
+%        solved = sum(successes(ldf.tmp, singleBest)),
+%        mean.performance = mean(parscores(ldf.tmp, singleBest, factor = 1)),
+%        median.performance = median(parscores(ldf.tmp, singleBest, factor = 1)))
+%
+%    modelc = classify(rflearner, ldf.tmp)
+%    resc = data.frame(model = "classification",
+%        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, modelc)),
+%        solved = sum(successes(ldf.tmp, modelc)),
+%        mean.performance = mean(parscores(ldf.tmp, modelc, factor = 1)),
+%        median.performance = median(parscores(ldf.tmp, modelc, factor = 1)))
+%
+%    modelr = regression(makeLearner("regr.randomForest"), ldf.tmp)
+%    resr = data.frame(model = "regression",
+%        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, modelr)),
+%        solved = sum(successes(ldf.tmp, modelr)),
+%        mean.performance = mean(parscores(ldf.tmp, modelr, factor = 1)),
+%        median.performance = median(parscores(ldf.tmp, modelr, factor = 1)))
+%
+%    modelcp = classifyPairs(rflearner, ldf.tmp)
+%    rescp = data.frame(model = "classifyPairs",
+%        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, modelcp)),
+%        solved = sum(successes(ldf.tmp, modelcp)),
+%        mean.performance = mean(parscores(ldf.tmp, modelcp, factor = 1)),
+%        median.performance = median(parscores(ldf.tmp, modelcp, factor = 1)))
+%
+%    modelrp = regressionPairs(makeLearner("regr.randomForest"), ldf.tmp)
+%    resrp = data.frame(model = "regressionPairs",
+%        mean.misclassification.penalty = mean(misclassificationPenalties(ldf.tmp, modelrp)),
+%        solved = sum(successes(ldf.tmp, modelrp)),
+%        mean.performance = mean(parscores(ldf.tmp, modelrp, factor = 1)),
+%        median.performance = median(parscores(ldf.tmp, modelrp, factor = 1)))
+%
+%    sb = levels(singleBest(ldf.tmp)$algorithm)
+%    cbind(data.frame(timeout = 10^to, single.best = sb), rbind(resvbs, ressb, resc, resr, rescp, resrp))
+%}))
+%
+%print(xtable(tab, digits = 0, caption = "Preliminary algorithm selection results for different timeouts with presolver."), booktabs = TRUE, include.rownames = FALSE)
+%@
 

### 2015-12-17 paper/paper.tex
nosep just on one list
update results (which are slightly worse this time)

--- 2015-12-16 paper/paper.tex
@@
 \usepackage{gnuplot-lua-tikz}
+\setlist{nosep}
 
@@
 
-\noindent Audemard et al.~\cite{Audemard:2014} do not limit the length of paths considered, and
+Audemard et al.~\cite{Audemard:2014} do not limit the length of paths considered, and
 iteratively increment the length until no more pairs are removed. Battiti and Mascia~\cite{battiti-mascia07}, and McCreesh and Prosser~\cite{McCreesh:2015} parameterise their algorithms
@@
 
-\noindent Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
+Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
 Prosser~\cite{McCreesh:2015}.  This set of instances is also much larger than that of Battiti and
@@
 
-\noindent Alongside these basic features, we include information computed by \IncompleteLAD. To (try to) prove
+Alongside these basic features, we include information computed by \IncompleteLAD. To (try to) prove
 inconsistency, \IncompleteLAD removes candidate pairs. The number of successfully removed pairs

### 2015-12-17 paper/paper.tex
Fix blatant lies

--- 2015-12-16 paper/paper.tex
@@
 the number of paths of length $k$ between the target nodes with which they may be matched.
+These invariant properties are used, for example,
 
-These invariant properties are used, for example,
-\begin{itemize}
+\begin{itemize}[nosep]
 \item by Battiti and Mascia~\cite{battiti-mascia07}, to remove candidate pairs $(u,v)\in N_p\times
@@
 \GlasgowTwo becomes the best solver for time limits larger than \SI{3000}{\ms}.
-As we
-increase the CPU time limit, variants of \Glasgow with longer paths (\GlasgowThree and \GlasgowFour)
-become better. This is what we expect, as more reasoning is expensive, but
-increases the potential
-reduction of the search space. In particular, the relative performance of \Glasgow variants is
-monotone in the time it takes to solve the instance---\GlasgowOne is best at first, but gets worse as
-the time to solve the instance increases until \GlasgowTwo is better. After that point, \GlasgowOne is
-never better than \GlasgowTwo again. The same is true for path lengths 3 and 4.
+As we increase the CPU time limit, the performance of variants of \Glasgow with
+longer paths (\GlasgowThree and \GlasgowFour) improves. This is what we expect,
+as more reasoning is expensive, but increases the potential reduction of the
+search space. Eventually, \GlasgowTwo and \GlasgowThree become very closely
+matched, and with runtimes very close to the limit, \GlasgowFour nearly catches
+up. (This behaviour is class-dependent: over class 2, for example, the behavior
+is roughly monotone, with \GlasgowOne dominating for low runtimes, then
+\GlasgowTwo, then \GlasgowThree, then \GlasgowFour each becoming best as the runtimes
+increase.)
 
-The figure illustrates the potential for portfolios and algorithm selection we have. There is
+The figure illustrates the potential for portfolios and algorithm selection we have: there is
 clearly no single solver that dominates throughout.

### 2015-12-17 paper/paper.tex
Minor editing

@@
 
-In Section~\ref{sec:algs}, we experimentally compare eight state-of-the-art
+In Section~\ref{sec:exps}, we experimentally compare eight state-of-the-art
 algorithms. We introduce a large benchmark set composed of 5725 instances
-grouped into twelve classes. Ten of these classes have been introduced in the
-experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}. We
-evaluate the algorithms on this benchmark set, and show that they have very
-complementary performance. In particular, we show that depending on the CPU
-time limit, different algorithms achieve the best performance on the entire
-benchmark set.
+grouped into twelve classes. Ten of these classes were considered in the
+experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}; two
+are new. We evaluate the algorithms on this benchmark set, and show that they
+have very complementary performance. In particular, we show that depending on
+the CPU time limit, different algorithms achieve the best performance on the
+entire benchmark set.
 
-In Section~\ref{sec:algsel}, we describe the features that are used to describe
+In Section~\ref{sec:algsel}, we discuss the features that are used to describe
 instances, and we describe our algorithm selection approach. It combines a
-presolving step, which allows us to very quickly solve easy instances, with the
+presolving step, which allows us to easy instances very quickly, with the
 algorithm selection approach \LLAMA~\cite{kotthoff_llama_2013}.
-
 In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
-approach, and show that it is able to close more than 50\% of the gap between
-the single best and the virtual best solver.
-
+approach, and show that it is able to close more than 50\% of the gap between the
+single best and the virtual best solver.
 We conclude and give directions for future work in Section~\ref{sec:concs}.
@@
 
-Subgraph isomorphism problems may be solved by a systematic exploration of the search space composed
+Subgraph isomorphism problems may be solved by a systematic exploration of the search space consisting
 of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one
@@
 filtering which ensures that for each pattern edge $(u,u')\in E_p$ and each candidate pair $(u,v)\in
-N_p\times N_t$, there exists a candidate couple $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a
-target edge. Candidate couples $(u,v)$ that do not satisfy this property are iteratively removed
+N_p\times N_t$, there exists a candidate pair $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a
+target edge. Candidate pairs $(u,v)$ that do not satisfy this property are iteratively removed
 until a fixed point is reached. This filtering (called \emph{Arc Consistency}) removes more
@@
     to invariant properties;
-\item by McCreesh and Prosser~\cite{McCreesh:2015} to filter the set of candidate couples before
+\item by McCreesh and Prosser~\cite{McCreesh:2015} to filter the set of candidate pairs before
     starting the search, and to generate additional implied adjacency-like constraints which are
@@
 
-\subsection{Novel Algorithms}
+\subsection{New Algorithms}
 
@@
 within the timeout, with only an additional 18 to the virtual best. In terms of average performance,
-we are able to more than \SI{50}{\percent} of the gap between the single best and the virtual
+we are able to close more than \SI{50}{\percent} of the gap between the single best and the virtual
 best solver.
 
-Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time for the individual
-solvers, the virtual best solver, and the \LLAMA algorithm selection approach. The performance of
-\VFtwo is far behind all other solvers.  The algorithm selection model does not perform well for
-instances that can be solved quickly because of the overhead incurred through feature computation.
-As the instances become more difficult to solve, its performance improves.
+Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time
+for the individual solvers, the virtual best solver, and the \LLAMA algorithm
+selection approach. The algorithm selection model does not perform well for
+instances that can be solved quickly because of the overhead incurred through
+feature computation.  As the instances become more difficult to solve, its
+performance improves.
 

### 2015-12-17 paper/paper.tex
Model analysis and more conclusion

@@
 
+\subsection{Analysis of Features Used by the Model}
+
+Analysing the final model, we saw that the most important features were, in order:
+
+\begin{itemize}
+    \item The proportion of target vertices which are at least distance 3
+        apart. We introduced this feature expecting it to be helpful in
+        distinguishing between \Glasgow variants: if few vertices are far
+        apart, longer paths are unlikely to be useful. However, in practice
+        this feature also gives a rough indication of how sparse the graph
+        is---locally all-different filtering is weak and expensive on dense
+        graphs, and the feature turned out to be helpful for selecting between
+        \Glasgow and \LAD variants too;
+
+    \item The mean degree of the target graph;
+    \item The maximum degree of the pattern graph;
+    \item The number of values removed during \IncompleteLAD presolving;
+    \item And the mean degree of the pattern graph, the number of vertices in
+        the pattern graph, and the number of vertices in the target graph,
+        which were used once each.
+\end{itemize}
+
+\noindent This confirms that our decision to consider both the pattern graph
+and the target graph for feature extraction was correct, and shows that basic
+graph properties do affect the performance of sophisticated filtering
+algorithms in reasonably simple ways.
+
+\subsection{Analysis of \PathLAD versus \GlasgowTwo}
+
+In an attempt to gain insight into the behaviour of the algorithms, we also
+investigated a simpler setting comparing just \PathLAD and \GlasgowTwo. This
+pair is of particular interest because they are the best ``medium-case''
+algorithms that use strong and weak filtering during search respectively. We
+used machine learning techniques to train a simple, human-understandable model
+which is able to distinguish these cases with reasonable accuracy. The model
+produced uses four rules:
+
+\begin{enumerate}
+    \item If \IncompleteLAD removes at least \SI{28}{\percent} of the pairs,
+        and at least \SI{94}{\percent} of the values from at least one domain,
+        then pick \PathLAD.
+    \item If the target has at least $610$ vertices, and if the maximum
+        distance between any two pattern vertices is at most $8$, and if the
+        pattern is not regular, and if the time taken to compute the
+        distance-based features on the target graph is no more than
+        \SI{1277}{\ms}, then pick \PathLAD.
+    \item If \IncompleteLAD filtering removed at least \SI{6}{\percent} of
+        possible pairs, and if less than \SI{85}{\percent} of the pattern
+        vertices are within distance 2 of each other, then pick \PathLAD.
+    \item Otherwise, pick \GlasgowTwo.
+\end{enumerate}
+
+\noindent
+The first and third rules here intuitively make sense: if \IncompleteLAD
+filtering does well, it is likely that continuing with this kind of filtering
+during search will be successful. The third rule also excludes using \PathLAD
+on very dense pattern graphs, where locally all-different filtering is
+expensive and weak. The second rule is less obvious: \PathLAD filtering is weak
+on regular graphs, so this part of the rule is unsurprising, but the remainder
+of the rule appears to be a round-about way of excluding large, dense target
+graphs. Still, this model suggests that it could be worth exploring
+\emph{dynamically} enabling or disabling locally all-different filtering during
+search, based upon very simple features which could be recomputed as search
+progresses and conditions change.
+
 \section{Conclusion and Future Work}\label{sec:concs}
@@
 with a new incomplete variant of \LAD that is able to detect inconsistencies
-and a presolver boosts performance even further.
+and a presolver boosts performance even further. Finally, we discussed how
+insights from machine learning could feed back into algorithm development.
+
+We have yet to investigate parallel portfolios: as well as the possibility of
+running multiple single-threaded solvers simultaneously, the \Glasgow
+algorithms have a multi-core parallel implementation available, which can use a
+configurable number of threads. Nor have we investigated variants of the
+subgraph isomorphism problem---it would be interesting to see, for example,
+whether similar features are important when dealing with induced or directed
+isomorphisms.
 

### 2015-12-17 paper/paper.tex
update results again, this time slightly better

@@
 In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
-approach, and show that it is able to close more than 50\% of the gap between the
-single best and the virtual best solver.
-We conclude and give directions for future work in Section~\ref{sec:concs}.
+approach, and show that it is able to close more than \SI{60}{\percent} of the
+gap between the single best and the virtual best solver. We conclude and give
+directions for future work in Section~\ref{sec:concs}.
 
@@
 limit of \SI{e8}{\ms}, i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
-algorithm selection approach.  We are able to solve 28 more instances than the single best solver
-within the timeout, with only an additional 18 to the virtual best. In terms of average performance,
-we are able to close more than \SI{50}{\percent} of the gap between the single best and the virtual
+algorithm selection approach.  We are able to solve 30 more instances than the single best solver
+within the timeout, with only an additional 16 to the virtual best. In terms of average performance,
+we are able to close \SI{64}{\percent} of the gap between the single best and the virtual
 best solver.
@@
 virtual best & 0 & 2219 & 5822809\\
-\LLAMA & 902316 & 2201 & 6726782\\
-\GlasgowTwo & 1960992 & 2173 & 7783801\\
+\LLAMA & 705097 & 2203 & 6529563\\
+\GlasgowTwo & 1960683 & 2173 & 7783492\\
    \bottomrule \\
@@
 entire algorithm selection system, including the preprocessing, is shown in Table~\ref{tab:resfull}.
-Our system is able to close more than \SI{50}{\percent} of the gap between
+Our system is able to close more than \SI{60}{\percent} of the gap between
 single and virtual best, similar to the results on the reduced set of instances.
@@
 virtual best & 0 & 5608 & 2375913\\
-\LLAMA & 368176 & 5590 & 2744765\\
+\LLAMA & 287704 & 5592 & 2664293\\
 \GlasgowTwo & 798660 & 5562 & 3174573\\

### 2015-12-17 paper/paper.tex
update order of important features
tweak section on feature analysis

@@
 \begin{itemize}
-    \item The proportion of target vertices which are at least distance 3
-        apart. We introduced this feature expecting it to be helpful in
-        distinguishing between \Glasgow variants: if few vertices are far
-        apart, longer paths are unlikely to be useful. However, in practice
-        this feature also gives a rough indication of how sparse the graph
-        is---locally all-different filtering is weak and expensive on dense
-        graphs, and the feature turned out to be helpful for selecting between
-        \Glasgow and \LAD variants too;
-
-    \item The mean degree of the target graph;
-    \item The maximum degree of the pattern graph;
-    \item The number of values removed during \IncompleteLAD presolving;
-    \item And the mean degree of the pattern graph, the number of vertices in
-        the pattern graph, and the number of vertices in the target graph,
-        which were used once each.
+    \item the maximum degree of the pattern graph;
+    \item the mean degree of the target graph;
+    \item the proportion of target vertices that are at least distance 3
+        apart;
+    \item the number of values removed during \IncompleteLAD presolving.
 \end{itemize}
 
-\noindent This confirms that our decision to consider both the pattern graph
-and the target graph for feature extraction was correct, and shows that basic
-graph properties do affect the performance of sophisticated filtering
-algorithms in reasonably simple ways.
+We introduced the proportion of target vertices that are at least 3 apart as a
+feature expecting it to be helpful in distinguishing between
+\Glasgow variants---if few vertices are far apart, longer paths are unlikely to
+be useful. However, in practice this feature also gives a rough indication of
+how sparse the graph is---locally all-different filtering is weak and expensive
+on dense graphs, and the feature turned out to be helpful for selecting between
+\Glasgow and \LAD variants too.
+
+As expected, both the pattern graph and the target graph provide important
+features. We also conclude that even basic graph properties are predictive of
+sophisticated algorithms' performance.
 

### 2015-12-17 paper/paper.tex
put in exact numbers for rules

@@
 \begin{enumerate}
-    \item If \IncompleteLAD removes at least \SI{28}{\percent} of the pairs,
-        and at least \SI{94}{\percent} of the values from at least one domain,
+    \item If \IncompleteLAD removes at least \SI{28.014755}{\percent} of the pairs,
+        and at least \SI{94.117645}{\percent} of the values from at least one domain,
         then pick \PathLAD.
@@
         \SI{1277}{\ms}, then pick \PathLAD.
-    \item If \IncompleteLAD filtering removed at least \SI{6}{\percent} of
-        possible pairs, and if less than \SI{85}{\percent} of the pattern
+    \item If \IncompleteLAD filtering removed at least \SI{5.904644}{\percent} of
+        possible pairs, and if less than \SI{84.6565}{\percent} of the pattern
         vertices are within distance 2 of each other, then pick \PathLAD.

### 2015-12-17 paper/paper.tex
tweak 6.2

@@
 
-In an attempt to gain insight into the behaviour of the algorithms, we also
-investigated a simpler setting comparing just \PathLAD and \GlasgowTwo. This
-pair is of particular interest because they are the best ``medium-case''
+To gain further insight into the behaviour of the algorithms, we
+investigated what affects the relative performance of \PathLAD and \GlasgowTwo.
+This pair is of particular interest because they are the best ``medium-case''
 algorithms that use strong and weak filtering during search respectively. We
 used machine learning techniques to train a simple, human-understandable model
-which is able to distinguish these cases with reasonable accuracy. The model
-produced uses four rules:
+which is able to distinguish these solvers and gives performance better than
+always choosing one of them. The model uses four rules:
 
@@
 \noindent
-The first and third rules here intuitively make sense: if \IncompleteLAD
+The first and third rules intuitively make sense: if \IncompleteLAD
 filtering does well, it is likely that continuing with this kind of filtering
@@
 on very dense pattern graphs, where locally all-different filtering is
-expensive and weak. The second rule is less obvious: \PathLAD filtering is weak
-on regular graphs, so this part of the rule is unsurprising, but the remainder
-of the rule appears to be a round-about way of excluding large, dense target
-graphs. Still, this model suggests that it could be worth exploring
-\emph{dynamically} enabling or disabling locally all-different filtering during
-search, based upon very simple features which could be recomputed as search
-progresses and conditions change.
+expensive and weak. The second rule is less obvious: while \PathLAD filtering is
+weak on regular graphs and it makes sense to exclude this case, the other
+components appear to exclude large and dense target graphs. The model suggests
+that it would be worth exploring \emph{dynamically} enabling or disabling
+locally all-different filtering during search, based upon very simple features
+which could be recomputed as search progresses and conditions change.
+
+This provides an interesting insight into the behaviour of our algorithms, as
+well as giving indications for future work.
 

### 2015-12-17 paper/paper.tex
tweak conclusions and future work

@@
 with a new incomplete variant of \LAD that is able to detect inconsistencies
-and a presolver boosts performance even further. Finally, we discussed how
-insights from machine learning could feed back into algorithm development.
+and a presolver boosts performance even further. Finally, we showed how
+insights from machine learning can guide algorithm development.
 
-We have yet to investigate parallel portfolios: as well as the possibility of
-running multiple single-threaded solvers simultaneously, the \Glasgow
-algorithms have a multi-core parallel implementation available, which can use a
-configurable number of threads. Nor have we investigated variants of the
-subgraph isomorphism problem---it would be interesting to see, for example,
-whether similar features are important when dealing with induced or directed
-isomorphisms.
+Directions for future work include scheduling multiple solvers to run instead of
+a single one; in particular the \Glasgow algorithms provide a multi-core
+parallel implementation, which can use a configurable number of threads. It
+would also be interesting to investigate other variants of the subgraph
+isomorphism problem.
 

### 2015-12-17 paper/paper.tex
Christine's changes

@@
     graphs commonly used for benchmarking subgraph isomorphism
-    algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for
+    algorithms~\cite{GraphDatabase1}: bounded-degree graphs for
     classes 4 and 5, regular meshes for classes 6 and 7, and random graphs with
@@
 
-\noindent Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
-Prosser~\cite{McCreesh:2015}.  This set of instances is also much larger than that of Battiti and
-Mascia~\cite{battiti-mascia07}, who were the first to propose algorithm portfolios for subgraph
-isomorphism problems.  Battiti and Mascia only considered a pure parallel portfolio consisting of
-two solvers, and without a selection mechanism. Their problem set consisted entirely of easy
+\noindent Note that Classes 3 and 12 were not considered in the previous
+experimental study by McCreesh and Prosser~\cite{McCreesh:2015}.  This set of
+instances is also much larger than that of Battiti and
+Mascia~\cite{battiti-mascia07}, who were the first to propose algorithm
+portfolios for subgraph isomorphism problems.  Battiti and Mascia only
+considered a pure parallel portfolio consisting of two randomised solvers, and
+without a selection mechanism. Their problem set consisted entirely of
 satisfiable instances.
@@
 Scientific Linux 6.5. We used the C++ implementation of the \Glasgow algorithm~\cite{McCreesh:2015},
-the C implementation of \LAD{}~\cite{Solnon:2010}, and the VFLib C
+the C implementation of \LAD{}~\cite{Solnon:2010}, and the VFLib C++
 implementation of \VFtwo{}~\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
@@
 distribution). As well as the number of removed pairs, we also record the percentage with respect to
-all possible pairs, and the minimum and maximum values on a per-variable basis.
+all possible pairs, and the minimum and maximum percentages of removed values on
+a per-variable basis.
 
@@
 misclassification penalty; that is, the additional time required to solve an instance because of
-choosing solvers that perform worse than the best. Mean performance is over all instances; when an
+choosing solvers that perform worse than the best. Mean MCP and performance are over all 2336 instances; when an
 instance is not solved, its performance is set to the time limit
@@
 milliseconds. In particular, we are interested in minimising the performance
-difference to the virtual best. Problem instances that take longer to solver
+difference to the virtual best. Problem instances that take longer to solve
 contribute more to this difference than easy instances and therefore carry more
@@
 used machine learning techniques to train a simple, human-understandable model
-which is able to distinguish these solvers and gives performance better than
-always choosing one of them. The model uses four rules:
+which is able to distinguish these solvers for the 2336 hard instances and gives
+performance better than always choosing one of them. The model uses four rules:
 

### 2015-12-17 paper/paper.tex
Little formatting tweaks

@@
 
+\noindent
 We introduced the proportion of target vertices that are at least 3 apart as a
@@
 \begin{enumerate}
-    \item If \IncompleteLAD removes at least \SI{28.014755}{\percent} of the pairs,
-        and at least \SI{94.117645}{\percent} of the values from at least one domain,
-        then pick \PathLAD.
+    \item If \IncompleteLAD presolving removes at least
+        \SI{28.014755}{\percent} of the pairs, and at least
+        \SI{94.117645}{\percent} of the values from at least one domain, then
+        pick \PathLAD.
     \item If the target has at least $610$ vertices, and if the maximum

### 2015-12-18 paper/paper.tex
This sentence no longer applies
slight reformatting
Table formatting
Prettier graphs

@@
 \begin{table}[p]
-\centering
-\begin{tabularx}{.8\textwidth}{XXrXrrrXrrrr}
+    \centering\setlength{\tabcolsep}{0.2em}
+\begin{tabularx}{.85\textwidth}{XXrXrrrXrrrr}
 \toprule
-Class && \VFtwo && \multicolumn{3}{c}{\LAD} && \multicolumn{4}{c}{\hspace*{2em}\Glasgow}\\
-&&&&incomplete&default&path&&1&2&3&4\\
+Class && \VFtwo && \multicolumn{3}{c}{\LAD} && \multicolumn{4}{c}{\Glasgow}\\
+      &&&&\textsc{Incomplete}&\textsc{Default}&\textsc{Path}&&\multicolumn{1}{c}{\textsc{1}}&\multicolumn{1}{c}{\textsc{2}}&\multicolumn{1}{c}{\textsc{3}}&\multicolumn{1}{c}{\textsc{4}}\\
 \midrule
@@
 
-Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
-baselines, on the reduced set of 2336 instances. The virtual best solver is the oracle predictor
-that, for each instance, chooses the best solver from our portfolio. This is the upper bound of what
-an algorithm selection approach can achieve. The single best solver is the one solver from the
-portfolio that has the overall best performance across the entire set of instances, at the CPU time
-limit of \SI{e8}{\ms}, i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
-algorithm selection approach.  We are able to solve 30 more instances than the single best solver
-within the timeout, with only an additional 16 to the virtual best. In terms of average performance,
-we are able to close \SI{64}{\percent} of the gap between the single best and the virtual
-best solver.
+Table~\ref{tab:res} shows the performance of our algorithm selection approach,
+compared to two baselines, on the reduced set of 2336 instances. The virtual
+best solver is the oracle predictor that, for each instance, chooses the best
+solver from our portfolio. This is the upper bound of what an algorithm
+selection approach can achieve. The single best solver is the one solver from
+the portfolio that has the overall best performance across the entire set of
+instances, at the CPU time limit of \SI{e8}{\ms}, i.e.\ \GlasgowTwo. We consider
+it a lower bound on the performance of the algorithm selection approach.  We are
+able to solve 30 more instances than the single best solver within the timeout,
+with only an additional 16 to the virtual best. In terms of average performance,
+we are able to close \SI{64}{\percent} of the gap between the single best and
+the virtual best solver.
 
@@
 
-\caption{Number of solved instances over time for the virtual best
-solver, \LLAMA, and individual solvers (shown as dotted lines, in the same
-colours as Figure~\ref{expTimeGraph}) on the reduced set of 2336 instances.}
+\caption{Number of solved instances over time for the virtual best solver,
+\LLAMA, and the single best solver \GlasgowTwo on the reduced set of 2336
+instances. Other individual solvers are shown as dotted lines, in the same
+colours as Figure~\ref{expTimeGraph}. }
 \label{fig:portfolio-ecdf}
@@
 
-Table~\ref{tab:res} shows the performance of the selection model on its own. The performance of the
-entire algorithm selection system, including the preprocessing, is shown in Table~\ref{tab:resfull}.
-Our system is able to close more than \SI{60}{\percent} of the gap between
-single and virtual best, similar to the results on the reduced set of instances.
+Table~\ref{tab:res} shows the performance of the selection model on its own. The
+performance of the entire algorithm selection system, including the
+preprocessing, is shown in Table~\ref{tab:resfull}. Our system is able to close
+more than \SI{60}{\percent} of the gap between single and virtual best, similar
+to the results on the reduced set of instances.
 
-Figure~\ref{fig:portfolio-ecdf-full} shows the number of solved instances over time for the
-algorithm selection system including \IncompleteLAD and \VFtwo presolving on the full set of
-instances. Note that for very easy instances, we are better than the virtual best solver because
-\IncompleteLAD is not included in our portfolio. The performance on small instances is much better
-than the \LLAMA selector alone (cf.\ Figure~\ref{fig:portfolio-ecdf}) and the region where \LLAMA
-performs worse than the individual solvers is now limited to approximately $10^2$ to $10^5$
+Figure~\ref{fig:portfolio-ecdf-full} shows the number of solved instances over
+time for the algorithm selection system including \IncompleteLAD and \VFtwo
+presolving on the full set of instances. The performance on small instances is
+much better than the \LLAMA selector alone (cf.\
+Figure~\ref{fig:portfolio-ecdf}) and the region where \LLAMA performs worse than
+the individual solvers is now limited to approximately $10^2$ to $10^5$
 milliseconds.
@@
     \input{graph-full}
-\caption{Number of solved instances over time for the virtual best
-solver, \LLAMA, and individual solvers (shown as dotted lines, in the same
-colours as Figure~\ref{expTimeGraph}) on the full set of 5725 instances.}
+\caption{Number of solved instances over time for the virtual best solver,
+\LLAMA, and the single best solver \GlasgowTwo on the full set of 5725
+instances. Other individual solvers are shown as dotted lines, in the same
+colours as Figure~\ref{expTimeGraph}. }
 \label{fig:portfolio-ecdf-full}
+
 \end{figure}

### 2015-12-18 paper/paper.tex
clarify ECDF

@@
 
-    \caption{Number of solved instances over CPU time for the eight algorithms we
+    \caption{Cumulative number of solved instances over CPU time for the eight algorithms we
         consider in this paper, and the virtual best solver (VBS) that shows the best
@@
 
-Figure~\ref{expTimeGraph} displays the evolution of the number of instances solved with respect to
+Figure~\ref{expTimeGraph} displays the evolution of the cumulative number of instances solved with respect to
 CPU time. It shows us that the best solver depends on the time limit considered. \IncompleteLAD is
@@
 
-Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time
+Figure~\ref{fig:portfolio-ecdf} shows the cumulative number of solved instances over time
 for the individual solvers, the virtual best solver, and the \LLAMA algorithm
@@
 
-\caption{Number of solved instances over time for the virtual best solver,
+\caption{Cumulative number of solved instances over time for the virtual best solver,
 \LLAMA, and the single best solver \GlasgowTwo on the reduced set of 2336
@@
 
-Figure~\ref{fig:portfolio-ecdf-full} shows the number of solved instances over
+Figure~\ref{fig:portfolio-ecdf-full} shows the cumulative number of solved instances over
 time for the algorithm selection system including \IncompleteLAD and \VFtwo
@@
     \input{graph-full}
-\caption{Number of solved instances over time for the virtual best solver,
+\caption{Cumulative number of solved instances over time for the virtual best solver,
 \LLAMA, and the single best solver \GlasgowTwo on the full set of 5725

### 2015-12-18 paper/paper.tex
small fixes

@@
 selection approach which leverages
-novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm
+novel features to characterise subgraph isomorphism problems and dynamically
+decides which algorithm
 to use on a per-instance basis. We demonstrate substantial performance improvements on a large set
@@
 instances, and we describe our algorithm selection approach. It combines a
-presolving step, which allows us to easy instances very quickly, with the
-algorithm selection approach \LLAMA~\cite{kotthoff_llama_2013}.
+presolving step, which allows us to easy instances very quickly, with an
+algorithm selection step that uses \LLAMA~\cite{kotthoff_llama_2013}.
 In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
-approach, and show that it is able to close more than \SI{60}{\percent} of the
+approach and show that it is able to close more than \SI{60}{\percent} of the
 gap between the single best and the virtual best solver. We conclude and give
@@
 
-In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of 
-subgraph isomorphism.  We  define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\#
-E_t$, and $d_p$ and $d_t$ to be the maximum degrees of the graphs $G_p$ and $G_t$.
+We define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\# E_t$, and $d_p$
+and $d_t$ to be the maximum degrees of the graphs $G_p$ and $G_t$.
 
@@
 
-We have selected a set of algorithms with complementary performance. These
-algorithms will compose our portfolio in the per-instance selection approach
-described in Section~\ref{sec:algsel}.
+We selected the following algorithms from the literature, based on their
+performance.
 \begin{itemize}
@@
 filtering which is applied once, without performing a backtracking search, and very quickly
-detects inconsistencies on many instances: for each pattern node $u$, we check that there exists at
+detects inconsistencies on many instances: for each pattern node $u$, we check
+if there exists at
 least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor
@@
 \noindent Note that Classes 3 and 12 were not considered in the previous
-experimental study by McCreesh and Prosser~\cite{McCreesh:2015}.  This set of
-instances is also much larger than that of Battiti and
+experimental study by McCreesh and Prosser~\cite{McCreesh:2015}.  Our set of
+instances is much larger than that of Battiti and
 Mascia~\cite{battiti-mascia07}, who were the first to propose algorithm
 portfolios for subgraph isomorphism problems.  Battiti and Mascia only
-considered a pure parallel portfolio consisting of two randomised solvers, and
+considered a pure parallel portfolio consisting of two randomised solvers
 without a selection mechanism. Their problem set consisted entirely of
@@
 CPU time. It shows us that the best solver depends on the time limit considered. \IncompleteLAD is
-able to solve easy unsatisfiable instances very quickly, in a few milliseconds. Hence, for time
-limits lower than \SI{5}{\ms}, the best solver is \IncompleteLAD. However, it is not able to solve harder
+able to solve easy unsatisfiable instances very quickly, in a few milliseconds.
+For time
+limits less than \SI{5}{\ms}, it is the best solver. However, it is not able to solve harder
 unsatisfiable instances, nor can it solve satisfiable instances.
@@
 \PathLAD and \GlasgowOne outperform \IncompleteLAD for longer time limits: \PathLAD is the
-best solver for time limits larger than \SI{5}{\ms} and lower than \SI{40}{\ms}, and \GlasgowOne is the best solver
-for time limits larger than \SI{40}{\ms} and lower than \SI{3000}{\ms}.
+best solver for time limits greater than \SI{5}{\ms} and less than \SI{40}{\ms}, and \GlasgowOne is the best solver
+for time limits greater than \SI{40}{\ms} and less than \SI{3000}{\ms}.
 
-\GlasgowTwo becomes the best solver for time limits larger than \SI{3000}{\ms}.
+\GlasgowTwo becomes the best solver for time limits greater than \SI{3000}{\ms}.
 As we increase the CPU time limit, the performance of variants of \Glasgow with
@@
 matched, and with runtimes very close to the limit, \GlasgowFour nearly catches
-up. (This behaviour is class-dependent: over class 2, for example, the behavior
+up. This behavior is class-dependent: for class 2, for example, the behavior
 is roughly monotone, with \GlasgowOne dominating for low runtimes, then
 \GlasgowTwo, then \GlasgowThree, then \GlasgowFour each becoming best as the runtimes
-increase.)
+increase.
 
@@
 
-Our approach is composed of three steps. First, we run two presolvers in a static way to quickly
-solve easy instances; second, we extract features from instances which are not solved by the first
-step; and third, we select an algorithm and run it.
+Our approach is composed of three steps. First, we run two presolvers in a
+static way to quickly solve easy instances. This ensures that we achieve good
+performance on instances that can be solved in a small amount of time. Second,
+we extract features from instances which are not solved by the first step.
+Finally, we run algorithm selection to choose the algorithm to solve the
+instance with.
 
@@
 
-Experimental results reported in Section~\ref{expComp} have shown us that \IncompleteLAD is very fast
-(\SI{7}{\ms} on average) and is able to solve 1919 instances from our benchmark
+Experimental results reported in Section~\ref{expComp} show that \IncompleteLAD is very fast
+(\SI{7}{\ms} on average) and able to solve 1919 instances from our benchmark
 set very quickly. Therefore, we first run
@@
 
-\VFtwo is also able to solve many easy instances very quickly: among the 3806 instances which are not
+\VFtwo is also able to solve many easy instances very quickly: from the 3806
+instances that are not
 solved by \IncompleteLAD, 1470 are solved by \VFtwo in less than 50 milliseconds. Therefore, after
@@
 
-We include \VFtwo in the portfolio, as it may solve an instance given more time,
-but not \IncompleteLAD.
+We also include \VFtwo in the portfolio, as it may solve an instance given more
+time, but not \IncompleteLAD, as it is an incomplete solver that cannot solve
+satisfiable instances.
 
@@
 
-If presolving does not give us a solution, we extract features. For both the pattern and
-the target graph, we consider some basic graph properties, which can be computed
+If presolving does not give us a solution, we extract features that characterize
+the instances. For both the pattern and
+the target graph, we consider some basic graph properties that can be computed
 very quickly:
@@
         dense graphs;
-    \item how many loops (self-adjacent vertices) the graph contains---since loops must be mapped to
+    \item how many loops (self-adjacent vertices) the graph contains---as loops must be mapped to
         loops, this could have a strong effect on how easy an instance is;
@@
 all possible pairs, and the minimum and maximum percentages of removed values on
-a per-variable basis.
+a per-variable basis. Finally, we include the CPU time required to compute these
+features as features.
 
@@
 Table~\ref{tab:res} shows the performance of our algorithm selection approach,
-compared to two baselines, on the reduced set of 2336 instances. The virtual
+compared to two baselines, on the set of 2336 hard instances. The virtual
 best solver is the oracle predictor that, for each instance, chooses the best
@@
 \end{tabular}
-\caption{Algorithm selection performance on the reduced set of 2336 instances. MCP is the
+\caption{Algorithm selection performance on the set of 2336 hard instances. MCP is the
 misclassification penalty; that is, the additional time required to solve an instance because of
@@
 \caption{Cumulative number of solved instances over time for the virtual best solver,
-\LLAMA, and the single best solver \GlasgowTwo on the reduced set of 2336
+\LLAMA, and the single best solver \GlasgowTwo on the set of 2336 hard
 instances. Other individual solvers are shown as dotted lines, in the same
-colours as Figure~\ref{expTimeGraph}. }
+colors as Figure~\ref{expTimeGraph}. }
 \label{fig:portfolio-ecdf}
@@
 more than \SI{60}{\percent} of the gap between single and virtual best, similar
-to the results on the reduced set of instances.
+to the results on the set of hard instances.
 
@@
 instances. Other individual solvers are shown as dotted lines, in the same
-colours as Figure~\ref{expTimeGraph}. }
+colors as Figure~\ref{expTimeGraph}. }
 \label{fig:portfolio-ecdf-full}
@@
 
-Even though Figures~\ref{fig:portfolio-ecdf} and~\ref{fig:portfolio-ecdf-full}
-show that on easy instances, the model does not perform well, it does over
-the entire set of instances (cf.\ Tables~\ref{tab:res} and~\ref{tab:resfull}).
+Figures~\ref{fig:portfolio-ecdf} and~\ref{fig:portfolio-ecdf-full}
+show that for the easy instances from the set of hard instances, the performance
+improvement through algorithm selection is negated by the cost of computing the
+features. The presolving steps improve performance dramatically over the full
+set of instances (cf.\ Tables~\ref{tab:res} and~\ref{tab:resfull}).
 
@@
 As expected, both the pattern graph and the target graph provide important
-features. We also conclude that even basic graph properties are predictive of
+features. We conclude that even basic graph properties are predictive of
 sophisticated algorithms' performance.
@@
 
-To gain further insight into the behaviour of the algorithms, we
+To gain further insight into the behavior of the algorithms, we
 investigated what affects the relative performance of \PathLAD and \GlasgowTwo.
 This pair is of particular interest because they are the best ``medium-case''
-algorithms that use strong and weak filtering during search respectively. We
-used machine learning techniques to train a simple, human-understandable model
-which is able to distinguish these solvers for the 2336 hard instances and gives
-performance better than always choosing one of them. The model uses four rules:
+algorithms that use strong and weak filtering during search, respectively. We
+used machine learning techniques (JRip~\cite{Cohen1995}) to train a simple,
+human-understandable model which is able to distinguish these solvers for the
+2336 hard instances and gives performance better than always choosing one of
+them. The model uses four rules:
 
@@
 
-This provides an interesting insight into the behaviour of our algorithms, as
+This provides an interesting insight into the behavior of our algorithms, as
 well as giving indications for future work.
@@
 
-The problem of identifying subgraph isomorphisms is a hard computation problem that has many
+The problem of identifying subgraph isomorphisms is a hard computational problem that has many
 applications in diverse areas. In this paper, we presented a portfolio of six algorithms from the

### 2015-12-18 portfolio/as-graphs.Rnw
random changes

@@
 
+print(xtable(table(singleBest(ldf2)$algorithm)))
+
 tab = rbind(resvbs, ressb, resc, resr, rescp, resrp)
@@
     median.performance = median(c(parscores(ldf2, modelrp, factor = 1), addtime)))
+
+print(xtable(table(singleBest(ldf)$algorithm)))
 
@@
 write.csv(wide, file = "no-presolved-presolver-allinstances.csv", row.names = FALSE)
+
+#rs = asscenario$feature.runstatus[c("instance_id", "lad_features")]
+#fc = asscenario$feature.costs[c("instance_id", "lad_features")]
+#sc = merge(rs, fc, by = "instance_id")
+#sc$time = sc$lad_features.y
+#sc$time[sc$lad_features.x != "presolved"] = 10^8
+#foo = rbind.fill(lapply(unique(wide$instance_id), function(id) { data.frame(instance_id = id, solver = "virtual.best.lad", time = min(sc$time[sc$instance_id == id], wide$time[wide$solver == "virtual.best"&wide$instance_id == id]), type = "pf") }))
+#newwide = rbind(wide, foo)
+
 p.full = ggplot(wide, aes(x = time, col = solver, linetype = type)) +

### 2016-04-12 paper/paper.tex
Christine's changes

@@
 vision~\cite{cviu11,pr15}, biochemistry~\cite{Giugno:2013}, and model checking~\cite{Sevegnani:2015}. There
-exist various exact algorithms, which have been compared on a large suite of problem instances by
+exist various exact algorithms, which have been compared on a large suite of instances by
 McCreesh and Prosser~\cite{McCreesh:2015}. These experiments indicated that the single best
-algorithm depends on the CPU time limit considered: for very small CPU time
+algorithm depends on the CPU time limit considered: for very small time
 limits, \VFtwo{}~\cite{Cordella:2004} is the best choice, whereas the \Glasgow algorithm~\cite{McCreesh:2015} has
@@
 
-Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
+% MODIFICATIONS
+In our subgraph isomorphism context, algorithm performance is highly constrained by memory bandwidth (as pointed out by Sabharwal and Samulowitz \cite{Sabharwal2014} for SAT solvers).
+Therefore, we cannot simply run different algorithms in parallel, and we consider the case where exactly one algorithm is selected for solving the problem. 
+%
+One of
 the most prominent and successful systems that employs this approach is
@@
 Solnon's \LAD algorithm \cite{Solnon:2010}.
-
 In Section~\ref{sec:exps}, we experimentally compare eight state-of-the-art
@@
 entire benchmark set.
-
 In Section~\ref{sec:algsel}, we discuss the features that are used to describe
@@
 an edge may be mapped to two target nodes which are linked by an edge.
-
 We define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\# E_t$, and $d_p$
@@
 \end{itemize}
+% MODIFICATIONS
+We have not considered the algorithm introduced in \cite{constraints10} because it is outperformed by \LAD{}.
+Also, we have not considered MIP nor SAT solvers because they are not competitive with the selected algorithms \cite{UpcomingIJCAIPaper}.
+%
 
@@
 
+% MODIFICATIONS: removed decimals (just keep to digits after the point)
 \begin{enumerate}
     \item If \IncompleteLAD presolving removes at least
-        \SI{28.014755}{\percent} of the pairs, and at least
-        \SI{94.117645}{\percent} of the values from at least one domain, then
+        \SI{28.01}{\percent} of the pairs, and at least
+        \SI{94.12}{\percent} of the values from at least one domain, then
         pick \PathLAD.
@@
         \SI{1277}{\ms}, then pick \PathLAD.
-    \item If \IncompleteLAD filtering removed at least \SI{5.904644}{\percent} of
-        possible pairs, and if less than \SI{84.6565}{\percent} of the pattern
+    \item If \IncompleteLAD filtering removed at least \SI{5.90}{\percent} of
+        possible pairs, and if less than \SI{84.66}{\percent} of the pattern
         vertices are within distance 2 of each other, then pick \PathLAD.

### 2016-04-14 paper/paper.tex
further small changes
submitted final version

@@
 
+\usepackage{enumitem}
+\setlist{nolistsep}
+
 \title{Portfolios of Subgraph Isomorphism Algorithms}
@@
     \and Ciaran McCreesh\thanks{This work was supported by the Engineering
-        and Physical Sciences Research Council [grant number EP/K503058/1]}\inst{2}
+        and Physical Sciences Research Council (grant number EP/K503058/1).}\inst{2}
     \and Christine Solnon\thanks{This work has been supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01).}\inst{3}}
@@
 a per-variable basis. Finally, we include the CPU time required to compute these
-features as features.
+features as features. However, those features were not more informative than the
+other ones.
 
